---
layout: page
author: Derek H. Ogle
title: MTH107 FAQ
subtitle: How do I install NCStats?
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/FAQs")
source("../../rhelpers/rhelpers.R")
fnm <- "install-NCStats"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message='FALSE'}
source("../../rhelpers/knitr_setup.R")
```

----

The `NCStats` package is a set of functions that I have written to make aspects of using R for statistics classes at Northland easier. Descriptions for installing `NCStats` are below. `NCStats` is likely loaded on all Northland College computers, though it may not be the most recent version at all times.

## Installing
As long as your computer is connected to the internet you can install `NCStats` directly from GitHub by running the following command in R.

```{r eval=FALSE}
if (!require('manipulate')) utils::install.packages('manipulate')
if (!require('remotes')) install.packages('remotes'); require('remotes')
remotes::install_github('droglenc/NCStats',type='binary')
```

## Checking the Installation
You can test whether the `NCStats` package installed correctly by running the line below (again, copy the line, paste it into the CONSOLE in RStudio, and press Enter/Return).

```{r}
library(NCStats)
```

If the installation is successful you should get about 20 lines of results in the console. As long as none of those lines says “Error” then the install should be correct. You can further test the install by running the first line below to see if you get the results on the two lines below that.

```{r}
Summarize(1:3)
```

If your installation failed then please contact me for help.
