---
layout: page
author: Derek H. Ogle
title: MTH107 FAQ
subtitle: How can I reorder the levels of my categorical variable?
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/FAQs")
source("../../../rhelpers/rhelpers.R")
fnm <- "reorder-levels"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message='FALSE'}
source("../../../rhelpers/knitr_setup.R")
```

----

By default, R will alphabetically order the levels of a factor (i.e., categorical) variable.  There may be times when you want an order other than alphabtical.  Reordering the levels may be accomplished by using the `levels=` argument with the `factor()` function.  The first argument to `factor()` is the name of the variable for which you wish to reorder the levels.  The `levels=` argument can then be set equal to a vector of level names that are spelled **exactly** as they appear in the original variable, but are placed in the order you wish.

For example, suppose that you used the code below to create a data.frame of information about three species of iris.
```{r}
df <- read.csv("data/Iris.csv")
str(df)
```

From this, it is seen that the `species` variable has three levels, but it is not clear what the three levels are.  The complete list of levels may be seen by including the variable name in `levels()` as shown below.
```{r}
levels(df$species)
```

Now suppose that, for some reason, you would like the levels of `species` to be reordered such that `versicolor` is listed fist, followed by `setosa` and then `viriginica`.  The code below creates a new variable called `species2` in which the species will follow this new odering
```{r}
df$species2 <- factor(df$species,levels=c("versicolor","setosa","virginica"))
levels(df$species2)
```

The new order is seen more clearly in the following two frequency tables.
```{r}
xtabs(~species,data=df)
xtabs(~species2,data=df)
```
