---
layout: page
author: Derek H. Ogle
title: MTH107 FAQ
subtitle: How can I reorder the levels of my categorical variable?
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/FAQs")
source("../../rhelpers/rhelpers.R")
fnm <- "reorder-levels"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message='FALSE'}
source("../../rhelpers/knitr_setup.R")
```

----

By default, R will order character vectors in the order that the words appear or alphabetically order the levels of a factor variable.For example, suppose that you used the code below to create a data.frame of information about three species of iris.

```{r}
df <- read.csv("data/Iris.csv")
str(df)
```

In this case suppose that we want to consider the species in a specific order. First, we need to see what names are present in the `species` variable.

```{r}
unique(df$species)
```

Second, we need to change the data type of `species` from character to a factor. This is done with `factor()`. However, at the same time we can set the order of the groups by setting the `levels=` argument to a vector of level names that are spelled **exactly** as they appear in the original variable, but are placed in the order you wish. For example, suppose that you would like `versicolor` to be listed fist, followed by `setosa` and then `viriginica`.

```{r}
df$species <- factor(df$species,levels=c("versicolor","setosa","virginica"))
```

You can then use `levels()` to see the ordering of the new factor variable.

```{r}
levels(df$species)
```

The new order is also seen in the following two frequency tables.
```{r}
xtabs(~species,data=df)
```

