---
layout: page
author: Derek H. Ogle
title: MTH107 FAQ
subtitle: How do I select one group of a factor variable?
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/FAQ/FAQs")
source("../../../rhelpers/rhelpers.R")
fnm <- "subset-1group"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message='FALSE'}
source("../../../rhelpers/knitr_setup.R")
```

----

In some instances you may want to create a smaller data frame that contains only those individuals that belong in one group. The smaller data frame can be constructed using `filterD()` from `FSA` (which is loaded with `NCStats`) with the `==` operator. For example, the following data frame contains measurements on petals and sepals from three species of iris.
```{r message=FALSE}
library(NCStats)
df <- read.csv("data/Iris.csv")
str(df)
levels(df$species)
```

One can select just one group with `filterD()` where the first argument is the original data frame and the second argument is a condition statement that uses the grouping (or factor) variable followed by `==` (note the two equals signs), and the name of single group contained in quotes. For example, one can select the `versicolor` species from the `species` variable in the `df` data frame with the following code.
```{r}
df2 <- filterD(df,species=="versicolor")
levels(df2$species)
```

Similar, one could select just the `virginica` species with the following code.
```{r}
df3 <- filterD(df,species=="virginica")
levels(df3$species)
```

One would use whichever data frame contains the specific information in all ensuing analyses. For example, one could do the following to compute summary statistics of petal length for the `versicolor` irises.
```{r}
Summarize(~petlen,data=df2,digits=1)
```

However, the following is the summary statistics of petal length for ALL three species (i.e., uses the original data frame).
```{r}
Summarize(~petlen,data=df,digits=1)
```

And for comparisons purposes.
```{r}
Summarize(petlen~species,data=df,digits=1)
```


This type of filtering or subsetting is also described in [this R Tutorial video](https://vimeo.com/user45324800/filterd).
