---
layout: page
author: Derek H. Ogle
title: MTH107 FAQ
subtitle: Why does the df for my two-sample t-test contain decimals?
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/FAQs")
source("../../../rhelpers/rhelpers.R")
fnm <- "t2test-df-decimals"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message='FALSE'}
source("../../../rhelpers/knitr_setup.R")
```

----

There are two types of two-sample t-tests -- (1) variances are equal and (2) variances are unequal.  In our introductory statistics course we only discussed the equal variances case.  In the equal variances case the degrees-of-freedom (df) are $n_{1}+n_{2}-2$ which alwasy results in a whole number.  However, the df formula for the unequal variances case is much more complicated and often results in decimal degrees-of-freedom.  Thus, if you expected the two-sample t-test df to be $n_{1}+n_{2}-2$, but they were a decimal, then you forgot to set `var.equal=TRUE` in `t.test()`.  Of course, before doing this you should verify with `levenesTest()` from `car` that the variances are statistically equal (see [this FAQ](cant-find-Levenes.html)).

As an example, the following is the results from using `t.test()` without setting `var.equal=TRUE` (*note that the unequal variances version is also called the Welch two-sample t-test*).
```{r}
t.test(extra~group,data=sleep)
```

In contrast, these are the results from using `t.test()` with `var.equal=TRUE` for the same data.
```{r}
t.test(extra~group,data=sleep,var.equal=TRUE)
```
