---
title: "Answer Key"
subtitle: "Data Production CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/")
source("../rhelpers/rhelpers.R")
modHTML("DataProduction_CE_Keys")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../rhelpers/knitr_setup.R")
```

----

<div class="alert alert-warning">
  <strong>Note:</strong> The key distinction is to determine if a treatment was imparted by the research or not. If it was, then it is an experiment.
</div>

## Study Types I

<div class="alert alert-warning">
  <strong>Note:</strong> The key distinction is to determine if a treatment was imparted by the research or not. If it was, then it is an experiment.
</div>

1. Experiment ... the people were given a treatment (placebo or fish oil).
1. Observational ... the women were "followed" over time, no treatment was imparted (the B12 deficiency or not was naturally induced).
1. Experimental ... a treatment (animations or not) was imparted.
1. Observational ... no treatment was imparted (observation of church records)
1. Experiment ... a treatment (type of trap) was imparted.

----

## Types of Observational Studies I
1. Convenience sample ... the individuals (a car) were not randomly selected nor did the individuals voluntarily choose to be part of the study. This data collection was simply convenient for the researcher.
1. Simple random sample (i.e., probability-based sample) ... individuals (a registration) were randomly selected and could not opt-out of the study.
1. Convenience sample ... individuals (a deer) were not randomly selected nor didt he individuals voluntarily choose to be part of the study. This data collection was simply convenient for the researcher.

----

## Blood Pressure Study
1. Experimental ... treatments (see below) were imposed.
1. The response variable is blood pressure.
1. Blood pressure is a quantitative, continuous variable because a blood pressure value can be found between any two blood pressure values.
1. There are two factor variables -- calcium and exercise.
1. There are two levels of calcium (pill or not) and two levels of exercise (do or don't).
1. There are four treatments (see diagram below, but is from 2 levels times 2 levels).
1. A "replicate" (i.e., an "individual") is a male subject.
1. There are 8 replicates (male subjects) per treatment (see diagram below, but note that this 32 male subjects total divided by 4 treatments).
1. A diagram of this experiment (rows are levels of one factor, columns are levels of other factor, each cell represents a treatment, and the numbers in each cell represent the randomly selected males allocated to that treatment).
```{r CAalloc, echo=FALSE, results="hide"}
alloc <- sample(32)
alloc[1:8]    # first treatment
alloc[9:16]   # second treatment
alloc[17:24]  # third treatment
alloc[25:32]  # fourth treatment
```

|             | No Pill | CA Pill |
|-------------|---------|---------|
| No Exercise | `r alloc[1:8]` | `r alloc[9:16]` |
| Exercise    | `r alloc[17:24]` | `r alloc[25:32]` |

#### R Appendix
```{r eval=FALSE, prompt=FALSE}
<<CAalloc>>
```

----

## Identify Response and Explanatory Variables I
1. The response variable is "success rate" because the distance of the putt effects the success rate (i.e., longer putts likely result in a lower success rate). And success rate (on putts) does not effect the length of a putt.
1. The response variable is "response time" because the amount of time awake effects the response time (i.e., the longer one is awake, to a point, they will likely show more dexterity). And dexterity (on the test) does not effect how long the subject was awake.
1. The response variable is growth because growth is effected by the amount of fertilizer.
1. It is not obvious which is the response variable. Thus, the response variable would depend on the researcher's question (i.e., if they were trying to predict shoe sales from hat sales, then shoe sales would be the response variable).

----

## Salt, Sand, and Ice
1. This was an experiment because the researchers controlled the levels of some variables at specific values (salt percentage and temperature) and attempted to control others at constant values.
1. The two factors are salt percentage and temperature.
1. There were two levels of salt percentage and three levels of temperature.
1. There were six (2*3) treatments in this experiment
1. The response variable was the melting rate of ice.
1. Melting rate is a continuous quantitative variable.
1. A replicate or individual in this experiment is a chamber with a small piece of highway in it.
1. A diagram of this experiment (rows are levels of one factor, columns are levels of other factor, each cell represents a treatment, and the numbers in each cell represent the randomly selected males allocated to that treatment).
```{r SSTalloc, echo=FALSE, results="hide"}
alloc2 <- sample(36)
alloc2[1:6]    # first treatment
alloc2[7:12]   # second treatment
alloc2[13:18]  # third treatment
alloc2[19:24]  # fourth treatment
alloc2[25:30]  # fifth treatment
alloc2[31:36]  # sixth treatment
```

|          | 5F      | 15F     | 25F     |
|----------|---------|---------|---------|
| 10% salt | `r alloc[1:6]` | `r alloc[7:12]` | `r alloc[13:18]` |
| 15% salt | `r alloc[19:24]` | `r alloc[25:30]` | `r alloc[31:36]` |

#### R Appendix
```{r eval=FALSE, prompt=FALSE}
<<SSTalloc>>
```


----

## Study Types II
1. Experiment ... the men were given a treatment (training program) with before/after results recorded
1. Observational ... no treatment was imparted (the brain cancer factor was naturally induced).
1. Observational ... no treatment was imparted (alcohol use was a choice of the women).
1. Observational ... no treatment was imparted (work at night was a "choice" of the women)
1. Observational ... no treatment was imposed on the subjects in the study and the subjects chose to participate in the study or not.

----

## Types of Observational Studies II
1. Voluntary response because even though car owners were randomly called they were given the opportunity to opt-out of the sample.
1. Voluntary response because the responses were not collected randomly; rather, the customer voluntarily chose to leave a response or not.

----

## Softball Bats and Balls
1. Experimental ... imparting treatments (ball types and bats were chosen specifically)
1. Individual .. a "hit"
1. Below
    1. Response ... Distance the ball was hit.
    1. Factors ... Brand of ball and type of bat
    1. Levels ... 2 brands of ball and 3 types of bat
    1. Treatments ... 6
    1. Replicates ... 4 (balls of each brand by each type of bat)
1. I would first randomly order the bats and then randomly order the two brands for each bat.
```{r}
sample(3)  # random order of bats
sample(8)  # first order of balls ... 1-4 is first brand, 5-8 second
sample(8)  # second order of balls
sample(8)  # third order of balls
```

----

## Identify Response and Explanatory Variables II
1. The response variable is sales of umbrellas because umbrella sales are effected by the amount of rainfall. And the amount of rainfall does not depend on umbrella sales.
1. The response variable is weight as the child does not get older as it weights more (but it could weigh more as it got older).

----
