---
layout: page
title: Goodness-of-Fit Tests
subtitle: R Handout
author: Derek H. Ogle
css: /css/modules.css
output:
  html_document:
    fig_width: 3.5
    fig_height: 3.5
    self_contained: no
  pdf_document:
    fig_width: 3.5
    fig_height: 3.5
  word_document:
    fig_width: 3.5
    fig_height: 3.5
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/")
source("../rhelpers/rhelpers.R")
modHTML("GOFTest_RHO")
```
```{r echo=FALSE, results="hide", message=FALSE, warnings=FALSE}
source("../rhelpers/knitr_setup.R")
purl2("GOFTest_RHO.Rmd")
```

----

## First Commands
```{r message=FALSE, warning=FALSE}
library(NCStats)
```

## Goodness-of-Fit Test
[Four-o'clocks (*Mirabilis jalapa*)](https://en.wikipedia.org/wiki/Mirabilis_jalapa) are plants native to tropical America.  Individual four-o'clocks can have red, white, or pink flowers.  Flower color in this species is thought to be controlled by a single gene locus with two alleles experssing incomplete dominance, so that heterozygotes are pink-flowered, while homozygotes for one allele are white-flowered and homozygotes for the other allele are red-flowered [(see this)](https://en.wikipedia.org/wiki/Mirabilis_jalapa#Genetic_studies).  According to [Mendelian genetic principles](https://en.wikipedia.org/wiki/Mendelian_inheritance), self-pollination of pink-flowered plants should produce progeny that have red, pink, and white flowers in a 1:2:1 ratio.  A horticulturist allowed several pink-flowered plants to self-pollinate and produce 240 progeny with 55 that were red-flowered, 132 that were pink-flowered, and 53 that were white-flowered.  Use the results to determine, at the 5% level, if the theoretical 1:2:1 ratio is upheld with these data.

```{r par1a=TRUE, fig.show='asis', warning=FALSE}
obs <- c(red=55,pink=132,white=53)
exp.p <- c(red=1/4,pink=2/4,white=1/4)
chi1 <- chisq.test(obs,p=exp.p,rescale.p=TRUE,correct=FALSE)
chi1$expected
chi1$observed
chi1
gofCI(chi1,digits=3)
```
