---
layout: page
title: Review Exercise
subtitle: Normal Distributions
author: Derek H. Ogle
css: "/css/modules.css"
output:
  html_document:
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/")
source("../rhelpers/rhelpers.R")
modHTML("NormalDist_RevEx_Simple")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../rhelpers/knitr_setup.R")
```

----

## Normal Distribution Characteristics I

Identify the mean and standard deviation for each population on the graph below (HINT: ``eyeball'' integers).
```{r NormDistGuess1, echo=FALSE}
par(mar=c(3.5,0.5,0.5,0.5),mgp=c(1.7,0.4,0),las=1,tcl=-0.2,yaxs="i")
x <- seq(-5,7,by=0.001)
B <- dnorm(x,-1,1)
A <- dnorm(x,1,2)
plot(B~x,type="l",lwd=3,xlab="Quantitative Variable (X)",
     ylab="",yaxt="n",xlim=c(-5,7),ylim=c(0,1.01*max(c(A,B))),col="red")
lines(B~x,lwd=3,col="red")
lines(A~x,lwd=3,col="blue")
abline(v=seq(-5,7,1),lwd=1,lty=3,col="lightgray")
text(-2.65,0.2,"B",col="red",cex=2)
text(3.5,0.13,"A",col="blue",cex=2)
```

[See answer here](zRevExAns/NormalDist_Simple.html#normal-distribution-characteristics-i)

----

## Normal Distribution Characteristics II

Identify the mean and standard deviation for each population on the graph below (HINT: ``eyeball'' integers).
```{r NormDistGuess2, echo=FALSE}
par(mar=c(3.5,0.5,0.5,0.5),mgp=c(1.7,0.4,0),las=1,tcl=-0.2,yaxs="i")
x <- seq(-20,60,by=0.001)
B <- dnorm(x,20,10)
A <- dnorm(x,30,5)
plot(B~x,type="l",lwd=3,xlab="Quantitative Variable (X)",
     ylab="",yaxt="n",xlim=c(-15,55),ylim=c(0,1.01*max(c(A,B))),col="red")
lines(B~x,lwd=3,col="red")
lines(A~x,lwd=3,col="blue")
grid(ny=NA)
text(5,0.02,"B",col="red",cex=2)
text(38,0.06,"A",col="blue",cex=2)
```

[See answer here](zRevExAns/NormalDist_Simple.html#normal-distribution-characteristics-ii)

----

## 68-95-99.7 (Empirical) Rule

You can click on each question to see an answer. Click on it again to close the answer.

<div class="panel-group">

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#EmpRule1">On any normal distribution, what percentage of individuals is within &plusmn;1&sigma; of &mu;?</a>
</div>
</div>
<div id="EmpRule1" class="panel-collapse collapse">
<div class="panel-body">Left-skewed</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#Shape2">What is a distribution with a long right-tail called?</a>
</div>
</div>
<div id="Shape2" class="panel-collapse collapse">
<div class="panel-body">Right-skewed</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#MeanMedian1">Is the mean divided by the median less than 1, equal to 1, or greater than 1 for a symmetric distribution?</a>
</div>
</div>
<div id="MeanMedian1" class="panel-collapse collapse">
<div class="panel-body">Equal to 1</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#MeanMedian2">Is the mean divided by the median less than 1, equal to 1, or greater than 1 for a strongly right-skewed distribution?</a>
</div>
</div>
<div id="MeanMedian2" class="panel-collapse collapse">
<div class="panel-body">Greater than 1</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#ShapeQuarts1">Is Q3-Q2 less than, approximately equal to, or greater than Q2-Q1 if the data are left-skewed?</a>
</div>
</div>
<div id="ShapeQuarts1" class="panel-collapse collapse">
<div class="panel-body">Less than</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#ShapeQuarts2">What is the shape of the distribution if Q3-Q2 is greater than Q2-Q1?</a>
</div>
</div>
<div id="ShapeQuarts2" class="panel-collapse collapse">
<div class="panel-body">Right-skewed</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#Which1">If the distribution is obviuosly skewed left, which measures should you generally use for center and dispersion?</a>
</div>
</div>
<div id="Which1" class="panel-collapse collapse">
<div class="panel-body">Median and IQR</div>
</div>
</div>
  
<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#Which2">Which measures of center and dispersion should you generally be used for an obviously right-skewed distribution?</a>
</div>
</div>
<div id="Which2" class="panel-collapse collapse">
<div class="panel-body">Median and IQR.</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#Which3">Which measures of center and dispersion should you generally be used for a symmetric distribution?</a>
</div>
</div>
<div id="Which3" class="panel-collapse collapse">
<div class="panel-body">Mean and standard deviation.</div>
</div>
</div>

</div>

----

## Shape from Histograms

Answer the questions below using the histograms below.
```{r HistQuest, echo=FALSE}
par(mfrow=c(1,3), mar=c(0.25,0.5,1,0.5))
set.seed(12489)
hist(rbeta(100,6,1),main="A",xlab="",ylab="",xaxt="n",yaxt="n",col="gray90")
hist(rbeta(100,1,6),main="B",xlab="",ylab="",xaxt="n",yaxt="n",col="gray90")
hist(rbeta(100,6,5),main="C",xlab="",ylab="",xaxt="n",yaxt="n",col="gray90")
```

1. What is the shape of each histogram?
1. For each histogram describe whether the mean is greater than, less than, or approximately equal to the median.

[See answer here](zRevExAns/UEDAQuant2_Defns.html#shape-from-histograms)

----

## Shape from Boxplots

Answer the questions below using the boxplots below.
```{r BoxQuest, echo=FALSE}
par(mar=c(2,0.5,1,0.5),mgp=c(1.9,0.5,0))
set.seed(12487)
d1 <- rbeta(100,16,15)
d2 <- rbeta(100,6,1)
d3 <- rbeta(100,1,6)
d1 <- d1-mean(d1)
d2 <- d2-mean(d2)
d3 <- d3-mean(d3)
bp <- data.frame(d=c(d1,d2,d3),g=factor(rep(c("A","B","C"),each=100)))
boxplot(d~g,data=bp,main="",xlab="",ylab="",yaxt="n",range=0)
```

1. What is the shape of each boxplot?

[See answer here](zRevExAns/UEDAQuant2_Defns.html#shape-from-boxplots)

----
