---
title: "Answer Key"
subtitle: "Normal Distributions CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/")
source("../rhelpers/rhelpers.R")
modHTML("NormalDist_CE_Keys")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../rhelpers/knitr_setup.R")
```

----

## Normal Distribution Characteristics

1. Using `snorm()`
    1. The whole distributions shifts to the right.
    1. The distribution becomes wider (more dispersed).
1. Below
    1. $\mu$=75, $\sigma$=10
    1. $\mu$=5500, $\sigma$=600
    1. $\mu$=0, $\sigma$=1

----

## Hand Calculations I
For X~N(75,10) 

1. 97.5%
1. 0.15%
1. 68%
1. 81.5%
1. Less than 55.
1. Between X=55 and 95.
1. Greater than X=95.

----

## Carpenter Ants
#### Top questions
1. A nest
1. Number of Carpenter Ants in a nest
1. Discrete quantitative
1. $\mu$=1400
1. $\sigma$=300

#### Bottom questions
1. `r formatC(distrib(1900,mean=1400,sd=300,lower.tail=FALSE)*100,format="f",digits=2)`%.
1. `r formatC(distrib(0.15,mean=1400,sd=300,type="q",lower.tail=FALSE),format="f",digits=0)` ants.
1. `r formatC(distrib(0.33,mean=1400,sd=300,type="q"),format="f",digits=0)` ants.
1. `r formatC((distrib(1900,mean=1400,sd=300)-distrib(700,mean=1400,sd=300))*100,format="f",digits=2)`%.
1. `r formatC(distrib(300,mean=1400,sd=300)*100,format="f",digits=3)`%.
1. `r formatC(distrib(0.05,mean=1400,sd=300,type="q",lower.tail=FALSE),format="f",digits=0)` ants.
1. Between `r formatC(distrib(0.10,mean=1400,sd=300,type="q"),format="f",digits=0)` and  `r formatC(distrib(0.10,mean=1400,sd=300,type="q",lower.tail=FALSE),format="f",digits=0)` ants.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(1900,mean=1400,sd=300,lower.tail=FALSE)*100
distrib(0.15,mean=1400,sd=300,type="q",lower.tail=FALSE)
distrib(0.33,mean=1400,sd=300,type="q")
ab <- distrib(1900,mean=1400,sd=300)
a <- distrib(700,mean=1400,sd=300)
(ab-a)*100
distrib(300,mean=1400,sd=300)*100
distrib(0.05,mean=1400,sd=300,type="q",lower.tail=FALSE)
distrib(0.10,mean=1400,sd=300,type="q")
distrib(0.10,mean=1400,sd=300,type="q",lower.tail=FALSE)
```

----

## Pollen Counts
#### Top questions
1. A day in LaCrosse, WI n early September
1. Pollen count (spores per cubic meter)
1. Discrete quantitative
1. $\mu$=40
1. $\sigma$=8

#### Bottom questions
1. `r formatC(distrib(0.1,mean=40,sd=8,type="q",lower.tail=FALSE),format="f",digits=1)` spores per cubic meter.
1. `r formatC(distrib(50,mean=40,sd=8),format="f",digits=3)`.
1. `r formatC(distrib(55,mean=40,sd=8)-distrib(32,mean=40,sd=8),format="f",digits=3)`.
1. `r formatC(distrib(0.3,mean=40,sd=8,type="q"),format="f",digits=1)` spores per cubic meter.
1. Between `r formatC(distrib(0.15,mean=40,sd=8,type="q"),format="f",digits=1)` and  `r formatC(distrib(0.15,mean=40,sd=8,type="q",lower.tail=FALSE),format="f",digits=1)` spores per cubic meter.
1. `r formatC(distrib(35,mean=40,sd=8,lower.tail=FALSE),format="f",digits=3)`.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(0.1,mean=40,sd=8,type="q",lower.tail=FALSE)
distrib(50,mean=40,sd=8)
ab <- distrib(55,mean=40,sd=8)
a <- distrib(32,mean=40,sd=8)
ab-a
distrib(0.3,mean=40,sd=8,type="q")
distrib(0.15,mean=40,sd=8,type="q")
distrib(0.15,mean=40,sd=8,type="q",lower.tail=FALSE)
distrib(35,mean=40,sd=8,lower.tail=FALSE)
```

----

## Driving Speed
#### Top questions
1. A car traveling between the two points
1. Time to travel between the two points
1. Continuous quantitative
1. $\mu$=2.5
1. $\sigma$=0.75

#### Bottom questions
1. `r formatC(distrib(0.15,mean=2.5,sd=0.75,type="q",lower.tail=FALSE),format="f",digits=2)` s
1. `r formatC(distrib(1,mean=2.5,sd=0.75),format="f",digits=3)`.
1. `r formatC(distrib(4.5,mean=2.5,sd=0.75)-distrib(1.5,mean=2.5,sd=0.75),format="f",digits=3)`.
1. The IQR is from a Q1 of `r formatC(distrib(0.25,mean=2.5,sd=0.75,type="q"),format="f",digits=2)` s to a Q3 of `r formatC(distrib(0.25,mean=2.5,sd=0.75,type="q",lower.tail=FALSE),format="f",digits=2)` s.
1. `r formatC(distrib(7,mean=2.5,sd=0.75,lower.tail=FALSE),format="f",digits=12)`.
1. 2.5 s.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(0.15,mean=2.5,sd=0.75,type="q",lower.tail=FALSE)
distrib(1,mean=2.5,sd=0.75)
ab <- distrib(4.5,mean=2.5,sd=0.75)
a <- distrib(1.5,mean=2.5,sd=0.75)
ab-a
distrib(0.25,mean=2.5,sd=0.75,type="q")
distrib(0.25,mean=2.5,sd=0.75,type="q",lower.tail=FALSE)
distrib(7,mean=2.5,sd=0.75,lower.tail=FALSE)
```

----

## Hand Calculations II
For Z~N(0,1)

1. Z=1
1. 2.5%
1. 99.85%
1. Between Z=-3 and 3.

----
