---
title: "Answer Key"
subtitle: "Probability CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/")
source("../rhelpers/rhelpers.R")
modHTML("Probability_CE_Keys")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../rhelpers/knitr_setup.R")
```

----

## Banking

1. `r formatC(distrib(800,mean=550,sd=150,lower.tail=FALSE),format="f",digits=4)` 
1. `r formatC(distrib(200,mean=550,sd=150),format="f",digits=4)` 
1. `r formatC(distrib(600,mean=550,sd=150),format="f",digits=4)` 
1. `r formatC(distrib(850,mean=550,sd=150)-distrib(700,mean=550,sd=150),format="f",digits=4)` 

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(800,mean=550,sd=150,lower.tail=FALSE)
distrib(200,mean=550,sd=150)
distrib(600,mean=550,sd=150)
ab <- distrib(850,mean=550,sd=150)
a <- distrib(700,mean=550,sd=150)
ab-a
```


----

## Acorn Production

1. `r formatC(distrib(1000,mean=875,sd=220,lower.tail=FALSE),format="f",digits=4)` 
1. `r formatC(distrib(1000,mean=875,sd=220)-distrib(700,mean=875,sd=220),format="f",digits=4)` 
1. `r formatC(distrib(5000/4,mean=875,sd=220),format="f",digits=4)` 
1. An Oak tree must be randomly selected.
1. `r formatC(distrib(0.20,mean=875,sd=220,type="q"),format="f",digits=0)` acorns.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(1000,mean=875,sd=220,lower.tail=FALSE)
ab <- distrib(1000,mean=875,sd=220)
a <- distrib(700,mean=875,sd=220)
ab-a
distrib(5000/4,mean=875,sd=220)
distrib(0.20,mean=875,sd=220,type="q")
```

----
## Melatonin and Sleep

1. About a mean from n=4 ...
    1. Sampling.
    1. Can be answered.
    1. N(15,`r formatC(5/sqrt(4),format="f",digits=3)`)
    1. `r formatC(distrib(12,mean=15,sd=5/sqrt(4),lower.tail=FALSE),format="f",digits=4)`.
1. About an individual ...
    1. Population.
    1. Can be answered.
    1. N(15,5)
    1. `r formatC(distrib(12,mean=15,sd=5,lower.tail=FALSE),format="f",digits=4)`.
1. About a mean from n=25 ...
    1. Sampling.
    1. Can be answered.
    1. N(15,`r formatC(5/sqrt(25),format="f",digits=3)`)
    1. `r formatC(distrib(15.5,mean=15,sd=5/sqrt(25),lower.tail=FALSE),format="f",digits=4)`.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(12,mean=15,sd=5/sqrt(4),lower.tail=FALSE)
distrib(12,mean=15,sd=5,lower.tail=FALSE)
distrib(15.5,mean=15,sd=5/sqrt(25),lower.tail=FALSE)
```

----

## Pollutants in Auto Exhaust

1. About an individual ...
    1. Population.
    1. Can NOT be answered.
    1. Right-skewed (not normal).
    1. Can't answer.
1. About a mean from n=15 ...
    1. Sampling.
    1. Can NOT be answered.
    1. Not normal.
    1. Can't answer.
1. About a mean from n=50 ...
    1. Sampling.
    1. Can be answered.
    1. N(0.9,`r formatC(0.15/sqrt(50),format="f",digits=4)`)
    1. `r formatC(distrib(0.92,mean=0.9,sd=0.15/sqrt(50),lower.tail=FALSE),format="f",digits=4)`.
1. About a mean from n=50 ...
    1. Sampling.
    1. Can be answered.
    1. N(0.9,`r formatC(0.15/sqrt(50),format="f",digits=4)`)
    1. `r formatC(distrib(0.85,mean=0.9,sd=0.15/sqrt(50)),format="f",digits=4)`.
1. About a mean from n=50 ...
    1. Sampling.
    1. Can be answered.
    1. N(0.9,`r formatC(0.15/sqrt(50),format="f",digits=4)`)
    1. `r formatC(distrib(0.01,mean=0.9,sd=0.15/sqrt(50),lower.tail=FALSE,type="q"),format="f",digits=3)` g/mi.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(0.92,mean=0.9,sd=0.15/sqrt(50),lower.tail=FALSE)
distrib(0.85,mean=0.9,sd=0.15/sqrt(50))
distrib(0.01,mean=0.9,sd=0.15/sqrt(50),lower.tail=FALSE,type="q")
```

----

## Acorn Length

1. About an individual ...
    1. Population.
    1. Can NOT be answered.
    1. Right-skewed (not normal).
    1. Can't answer.
1. About a mean from n=20 ...
    1. Sampling.
    1. Can be answered.
    1. N(21,`r formatC(6/sqrt(20),format="f",digits=4)`)
    1. `r formatC(distrib(25,mean=21,sd=6/sqrt(20),lower.tail=FALSE),format="f",digits=4)`.
1. About a mean from n=50 ...
    1. Sampling.
    1. Can be answered.
    1. N(21,`r formatC(6/sqrt(50),format="f",digits=4)`)
    1. `r formatC(distrib(25,mean=21,sd=6/sqrt(50))-distrib(20,mean=21,sd=6/sqrt(50)),format="f",digits=4)`.
1. About a mean from n=6 ...
    1. Sampling.
    1. Can NOT be answered.
    1. Not normal.
    1. Can't answer.
1. About an individual ...
    1. Population.
    1. Can NOT be answered.
    1. Right-skewed (not normal).
    1. Can't answer.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(25,mean=21,sd=6/sqrt(20),lower.tail=FALSE)
ab <- distrib(25,mean=21,sd=6/sqrt(50))
a <- distrib(20,mean=21,sd=6/sqrt(50))
ab-a
```

----

## People in Car

1. No.  Left side will be truncated at 0 (i.e., more than 2 SD away from mean becomes negative, which is not possible).  Discrete data with lots of ones and twos with few higher than that.
1. About a mean from n=40 ...
    1. Sampling.
    1. Can be answered.
    1. N(1.5,`r formatC(0.75/sqrt(40),format="f",digits=4)`)
    1. `r formatC(distrib(1.2,mean=1.5,sd=0.75/sqrt(40)),format="f",digits=4)`.
1. About a mean from n=50 ...
    1. Sampling.
    1. Can be answered.
    1. N(1.5,`r formatC(0.75/sqrt(50),format="f",digits=4)`)
    1. `r formatC(distrib(2,mean=1.5,sd=0.75/sqrt(50),lower.tail=FALSE),format="f",digits=7)`.
1. About a mean from n=5 ...
    1. Sampling.
    1. Can NOT be answered.
    1. Not normal.
    1. Can't answer.
1. About a mean from n=700 (through a total) ...
    1. Sampling.
    1. Can be answered.
    1. N(1.5,`r formatC(0.75/sqrt(700),format="f",digits=4)`)
    1. `r formatC(distrib(1075/700,mean=1.5,sd=0.75/sqrt(700),lower.tail=FALSE),format="f",digits=4)`.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(1.2,mean=1.5,sd=0.75/sqrt(40))
distrib(2,mean=1.5,sd=0.75/sqrt(50),lower.tail=FALSE)
distrib(1075/700,mean=1.5,sd=0.75/sqrt(700),lower.tail=FALSE)
```

----

## Plowing Snow

1. About a mean from n=5 ...
    1. Sampling.
    1. Can NOT be answered.
    1. Not normal.
    1. Can't answer.
1. About an individual ...
    1. Population.
    1. Can NOT be answered.
    1. Right-skewed (not normal).
    1. Can't answer.
1. About a mean from n=40 ...
    1. Sampling.
    1. Can be answered.
    1. N(45,`r formatC(28/sqrt(40),format="f",digits=4)`)
    1. `r formatC(distrib(50,mean=45,sd=28/sqrt(40))-distrib(40,mean=45,sd=28/sqrt(40)),format="f",digits=4)`.
1. About a mean from n=40 ...
    1. Sampling.
    1. Can be answered.
    1. N(45,`r formatC(28/sqrt(40),format="f",digits=4)`)
    1. `r formatC(distrib(0.25,mean=45,sd=28/sqrt(40),type="q"),format="f",digits=1)` secs.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
ab <- distrib(50,mean=45,sd=28/sqrt(40))
a <- distrib(40,mean=45,sd=28/sqrt(40))
distrib(0.25,mean=45,sd=28/sqrt(40),type="q")
```

----
