---
title: "Answer Key"
subtitle: "Why Statistics CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/")
source("../rhelpers/rhelpers.R")
modHTML("WhyStats_CE_Keys")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../rhelpers/knitr_setup.R")
set.seed(388293)
library(captioner)
tbls <- captioner(prefix="Table")
tbls(name="Frogs1","Visual and call counts of frogs on the first sample of 10 lakes from northern Wisconsin.")
tbls(name="Frogs2","Visual and call counts of frogs on the second sample of 10 lakes from northern Wisconsin.")
```

----

## Realities
1. The three "realities" that, if they did not exist, would eliminate the need for the field of statistics are that (i) you cannot "see" the entire population and must, therefore, take a sample, (ii) variability exists among individuals and among statistics, and (iii) conclusions about the population are uncertain.

<br>

## Variability
1. Natural variability is the realization that no two individuals are exactly identical. Sampling variability is the realization that the results (statistics) from any two samples are not exactly identical.
1. Consider a situation where a biologist is sampling Bluegill from a particular lake. Natural variability is the realization that no two Bluegill will be exactly identical; i.e., their lengths, weights, number of scales, etc. will be different. Sampling variability is the realization that the results from any two samples of Bluegill will not be identical. For example, the mean lengths from two samples may differ, or the mean weights, or the mean number of scales, or the mean percentage of sexually mature fish.

<br>

----

## Wisconsin Lakes
```{r Lakes, echo=FALSE, results='hide'}
named <- filterD(ABDLakes,named)
smpl1 <- srsdf(named,n=50,vars=c("county","area"))
smpl2 <- srsdf(named,n=50,vars=c("county","area"))
```

1. The sample of 50 lakes is now shown (to save space). The area of the first three lakes were all different, which is an example of natural variability.
1. The proportion of lakes in my first sample that were from Bayfield County was `r formatC(prop.table(xtabs(~county,data=smpl1))[["Bayfield"]],format="f",digits=3)`.
1. The proportion of lakes in my second sample that were from Bayfield County was `r formatC(prop.table(xtabs(~county,data=smpl2))[["Bayfield"]],format="f",digits=3)`. These sample proportions differ, which is an example of sampling variability.
1. My two sample proportions did not equal the population proportions. I was not surprized because samples do not perfectly represent the entire population. I would expect the statistics from the sample to be near (on average) the population parameter, but not always equal to it.

----

## Frog Survey
```{r Frogs, echo=FALSE,results="hide",message=FALSE}
( sampA <- srsdf(Frogs,n=10) )
Summarize(~visual,data=sampA,digits=1)
( sampB <- srsdf(Frogs,n=10) )
Summarize(~visual,data=sampB,digits=1)
```

1. An individual is a lake in northern Wisconsin (north of Highway 8) in 2010.
1. No, it is not reasonable to count frogs on every lake in northern Wisconsin because there is no reasonable way to get to every lake in the northern Wisconsin in a week.
1. My first sample of ten lakes is in `r tbls("Frogs1",display="cite")`.

`r tbls("Frogs1")`

```{r echo=FALSE}
print(sampA,row.names=FALSE)
```

4. The number of frogs differs between the first three lakes `r tbls("Frogs1",display="cite")`. This is an example of natural variability.
1. The average number of frogs in my first sample is `r formatC(mean(sampA$visual),format="f",digits=1)`.
1. My second sample of ten lakes is in `r tbls("Frogs2",display="cite")`.

`r tbls("Frogs2")`

```{r echo=FALSE}
print(sampB,row.names=FALSE)
```

1. The average number of frogs in my first sample is `r formatC(mean(sampB$visual),format="f",digits=1)`.
1. The average number of frogs per lake differed between the two samples. This is an example of sampling variability.
1. Neither average number of frogs per lake equalled the known population average number of frogs per lake. I was not surprised by this result because a sample is representative but not perfectly representative of the entire population.

----

R Appendix
```{r eval=FALSE, prompt=FALSE}
library(NCStats)

## Lakes Exercise
<<Lakes>>

## Frog Exercise
<<Frogs>>
```

----
