---
layout: page
title: Class Exercise 1
subtitle: Univariate EDA - Quantitative
author: Derek H. Ogle
css: "/css/lecture.css"
output:
  html_document:
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/UnivEDA_Quantitative/")
source("../../rhelpers/rhelpers.R")
fnm <- "CE1"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
fnm <- "CE1"
source("../../rhelpers/knitr_setup.R")
```

----

## Histograms

1. The [U.S. Census](http://www.indexmundi.com/facts/united-states/quick-facts/all-states/average-commute-time#map) asked individuals 16 years old and older "How long in minutes is your one-way commute to work each day?".  The mean for individuals from each state was computed and is plotted in the histogram below.  Use this information to answer the questions further below.

```{r echo=FALSE, fig.cap="Histogram of mean commute time for U.S. states.", fig.width=3.5, fig.height=3.5}
df <- data.frame(time=c(24.2,18.8,24.6,21.3,27.2,24.5,24.8,24.8,29.7,25.9,27,
                        26,20,28,23.2,18.8,19,22.8,24.9,23.3,32,28,24,22.9,
                        23.9,23.1,18,18.1,23.8,26.3,30.4,21.6,31.6,23.6,16.9,
                        23,21,22.5,25.9,23.6,23.5,16.9,24.3,25,21.4,22.2,27.7,
                        25.7,25.5,21.7,18.3))
hist(~time,data=df,breaks=seq(16,32,2),xaxt="n",xlab="Mean Commute Time (mins)")
axis(1,seq(16,32,2))
```

    1. What is an individual in this example?
    1. What type of variable is recorded?
    1. How many individuals are represented in this histogram?
    1. How many individuals had a mean commute time between 16 and 18 minutes?
    1. How many states had a mean commute time between 18 and 24 minutes?
    1. What was the range of mean commute times for the bin with exactly five states?
    1. What is the shape of this distribution.
    1. Are there any outliers represented in this distribution?
    
1. Describe the shape for each histogram shown below.

```{r echo=FALSE, fig.height=3, fig.width=3, fig.show='hold'}
set.seed(125689)
n <- 200
xlbl <- "Quantitative Variable"
par(mar=c(2,2,2,2),mgp=c(0.5,0,0),tcl=-0.2)
hist(~rbeta(n,1,1),xlab=xlbl,main="#1",xaxt="n",yaxt="n")
hist(~rbeta(n,1,6),xlab=xlbl,main="#2",xaxt="n",yaxt="n")
hist(~rbeta(n,6,3),xlab=xlbl,main="#3",xaxt="n",yaxt="n")
hist(~rbeta(n,2,7),xlab=xlbl,main="#4",xaxt="n",yaxt="n")
hist(~rbeta(n,5,5),xlab=xlbl,main="#5",xaxt="n",yaxt="n")
hist(~rbeta(n,9,1),xlab=xlbl,main="#6",xaxt="n",yaxt="n")
hist(~rbeta(n,8,4),xlab=xlbl,main="#7",xaxt="n",yaxt="n")
hist(~c(rbeta(n/2,7,2),rbeta(n/2,2,7)),xlab=xlbl,main="#8",xaxt="n",yaxt="n")
```


----

## Hand Calculations

Consider the following three data sets for the questions below.  Show your work for all questions.

<pre>
Data Set 1
18 28 25 21 16 24
<pre>

<pre>
Data Set 2
4 54 16 85 52 29 24 71 61 60 2
<pre>

<pre>
Data Set 3
93 81 34 5 54 84 54 13 1 35 79 63 97 71
<pre>

1. Hand calculate the mean of Data Set 1.
1. Hand calculate the standard deviation of Data Set 1.

1. Hand calculate the median of Data Set 2.
1. Hand calculate the IQR of Data Set 2.

1. Hand calculate the median of Dat Set 3.
1. Hand calculate the IQR of Dat Set 3.
1. Hand calculate the range of Data Set 3.

You can check your work with `sdCalc()` and `iqrCalc`.  For example,
```{r}
d <- c(16,23,47,13,18,12)
sdCalc(d)
iqrCalc(d)
```


## Calculate Measures of Dispersion

1. Calculate the standard deviation of 
1. 
