---
layout: page
title: Univariate EDA - Quantitative
subtitle: Class Exercise 1
author: Derek H. Ogle
css: "/css/modules.css"
output:
  html_document:
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/UnivEDA_Quantitative/")
source("../../rhelpers/rhelpers.R")
fnm <- "CE1"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
fnm <- "CE1"
source("../../rhelpers/knitr_setup.R")
```

----

## Histograms I

The [U.S. Census](http://www.indexmundi.com/facts/united-states/quick-facts/all-states/average-commute-time#map) asked individuals 16 years old and older "How long in minutes is your one-way commute to work each day?".  The mean for individuals from each state was computed and is plotted in the histogram below.

```{r echo=FALSE, fig.width=4, fig.height=4, par1=TRUE}
df <- data.frame(time=c(24.2,18.8,24.6,21.3,27.2,24.5,24.8,24.8,29.7,25.9,27,
                        26,20,28,23.2,18.8,19,22.8,24.9,23.3,32,28,24,22.9,
                        23.9,23.1,18,18.1,23.8,26.3,30.4,21.6,31.6,23.6,16.9,
                        23,21,22.5,25.9,23.6,23.5,16.9,24.3,25,21.4,22.2,27.7,
                        25.7,25.5,21.7,18.3))
hist(~time,data=df,breaks=seq(16,32,2),xaxt="n",xlab="Mean Commute Time (mins)")
axis(1,seq(16,32,2))
```

Use this histogram to answer these questions.

1. What is an individual in this example?
1. What type of variable is recorded?
1. How many individuals are represented in this histogram?
1. How many individuals had a mean commute time between 16 and 18 minutes?
1. How many states had a mean commute time between 18 and 24 minutes?
1. What was the range of mean commute times for the bin with exactly five states?
1. What is the shape of this distribution.
1. Are there any outliers represented in this distribution?

----

## Histograms II

For each histogram below, describe the shape and presence of outliers.

```{r echo=FALSE, fig.height=3, fig.width=3, fig.show='hold'}
set.seed(125689)
n <- 200
xlbl <- "Quantitative Variable"
par(mar=c(2,2,2,2),mgp=c(0.5,0,0),tcl=-0.2)
hist(~rbeta(n,1,6),xlab=xlbl,main="#1",xaxt="n",yaxt="n")
hist(~rbeta(n,6,3),xlab=xlbl,main="#2",xaxt="n",yaxt="n")
hist(~c(rbeta(n,2,7),0.91),xlab=xlbl,main="#3",xaxt="n",yaxt="n")
hist(~rbeta(n,5,5),xlab=xlbl,main="#4",xaxt="n",yaxt="n")
hist(~rbeta(n,1,1),xlab=xlbl,main="#5",xaxt="n",yaxt="n")
hist(~rbeta(n,9,1),xlab=xlbl,main="#6",xaxt="n",yaxt="n")
hist(~c(rbeta(n,8,4),0),xlab=xlbl,main="#7",xaxt="n",yaxt="n")
hist(~c(rbeta(n/2,7,2),rbeta(n/2,2,7)),xlab=xlbl,main="#8",xaxt="n",yaxt="n")
```

----

## Hand Calculations

Consider the following three data sets for the questions below.  Show your work for all questions.

<pre>
Data Set 1
18 28 25 21 16 24

Data Set 2
4 54 16 85 52 29 24 71 61 60 2

Data Set 3
93 81 34 5 54 84 54 13 1 35 79 63 97 71
</pre>

1. Hand calculate the mean of Data Set 1.
1. Hand calculate the standard deviation of Data Set 1.
1. Hand calculate the median of Data Set 2.
1. Hand calculate the IQR of Data Set 2.
1. Hand calculate the median of Dat Set 3.
1. Hand calculate the IQR of Dat Set 3.
1. Hand calculate the range of Data Set 3.

You can check your work with `sdCalc()` and `iqrCalc()`.  For example,
```{r}
d <- c(16,23,47,13,18,12)
sdCalc(d)
iqrCalc(d)
```

----

## Mean Versus Median

Using the `meanMedian()` function with no arguments will open a plot constructed from random data that looks like that below.

```{r echo=FALSE, fig.height=4, fig.width=5}
meanMedian()
```

The top plot is a histogram.  The lower plot shows individuals above the median (blue vertical line) with red dots and individuals below the median with black dots.  In addition, the dashed horizontal lines represent the differences of each individual from the mean (red vertical line) with negative values in black and positive values in red.

In RStudio, this plot will also have a small "gear" icon in the upper-left corner.  Clicking the gear opens a dialog box where you can control the sample size, two shape parameters, and whether an outlier is modeled or not.  In the questions below, leave the sample size at 30 and press the "reradnomize" button several (~10) times for each question to make sure that your conclusions are not based on a single random sample.

1. For each scenario below identify (i) the shape of the distribution, (ii) whether a substantive outlier is present, and (iii) whether the value of the mean is more or less than the value of the median.
    1. `shape1=1`, `shape2=1`, and no outliers.
    1. `shape1=8`, `shape2=1`, and no outliers.
    1. `shape1=1`, `shape2=8`, and no outliers.
    1. `shape1=3`, `shape2=3`, and no outliers.
    1. `shape1=3`, `shape2=3`, and an outlier at the `min`.
    1. `shape1=3`, `shape2=3`, and an outlier at the `max`.
1. Can you make a conclusion about the effect of the shape of the distribution on the relative values of the mean and median?  [*Compare your observations from scenarios 1-4.*] 
1. Can you make a conclusion about the effect of outliers on the relative values of the mean and median?  [*Compare your observations from scenarios 4-6.*]
1. Return to the [Histograms II](#histograms-ii) question and, for each histogram, identify whether the mean would be less than, greater than, or approximately equal to the median.
1. Why do you think the mean and median behave this way?

----

<div class="text-center">
<ul class="pagination pagination-lg">
  <li><a href="index.html">^</a></li>
  <li class="active"><a href="#">1</a></li>
  <li><a href="CE2.html">2</a></li>
</ul>
