---
layout: page
author: Derek H. Ogle
title: MTH107 FAQ
subtitle: Why does R say the alternative hypothesis is true, even though the p-value>alpha?
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/RFAQ/FAQs")
source("../../../rhelpers/rhelpers.R")
fnm <- "alt-hyp-misinterp"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message='FALSE'}
source("../../../rhelpers/knitr_setup.R")
```

----

Typical results from a `t.test()` (two-sample t-test in this exapmle) look like this.
```{r}
t.test(extra~group,data=sleep,var.equal=TRUE)
```

We would not reject the null hypothesis with these results if $\alpha=0.05$ was being used (note `kPvalue(t.test(extra~group,data=sleep,var.equal=TRUE)$p.value)`$\alpha$).  However, some students mistakingly interpret the third line of output as saying that the alternative hypothesis is true (and thus the null hypothesis is rejected).  R does NOT make the rejection decision for you.  The third line of output is simply a reminder of what you set the alternative hypothesis to in `alt=` of `t.test()` (note that the default is `alt="two.sided"` or "not equals").  You must decide to reject or not to reject the null hypothesis by comparing the p-value reported in the second line of output to the $\alpha$ that you have chosen to use.
