---
layout: page
author: Derek H. Ogle
title: MTH107 FAQ
subtitle: Why can't R find the XXX function?
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/RFAQ/FAQs")
source("../../../rhelpers/rhelpers.R")
fnm <- "cant-find-function"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message='FALSE'}
source("../../../rhelpers/knitr_setup.R")
```

----

This error usually occurs for one of three reasons described in the sections below.

## Typo

Most of the time this error occurs because you typed the name of the function incorrectly (notice mispelling of `library()`).
```{r error=TRUE}
libary(NCStats)
```


## NCStats is not Loaded

If the problem is not a typo, then this error typically occurs because you do not have a required package loaded.  More often than not in this class that means that you did not load `NCStats` with ...
```{r eval=FALSE}
library(NCStats)
```

If `NCStats` is not installed on your computer then see [this FAQ.](install-NCStats.html)

## Other Package is not Loaded

Some functions used in this course are contained in other packages, most of which are loaded when `NCStats` is loaded.  However, some of these functions will required that the other package be installed and loaded.  Installation of these packages [in this FAQ](install-other-packages.html).  Packages are loaded with `library()` as demonstrated above for `NCStats`.
