---
layout: page
author: Derek H. Ogle
title: MTH107 FAQ
subtitle: read.csv() does not find my data file
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/RFAQ/FAQs")
source("../../../rhelpers/rhelpers.R")
fnm <- "cant-find-file"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message='FALSE'}
source("../../../rhelpers/knitr_setup.R")
```

----

This error typically looks like this ...
```{r error=TRUE}
df <- read.csv("no-header.csv")
```

The most common reasons for this error are as follows

* You have misspelled the actual name of the file.
* You have not changed the working directory to where the data file is saved.  See [this FAQ]() on how to change the working directory.  Make sure you change the working directory to where your file is actually saved.
* You have not saved the file as a CSV file.  Most commonly you saved it as an Excel file.  Saving an Excel file as a CSV file is described in [this R Tutorial video](https://vimeo.com/user45324800/ncstats-preparedataexcel).
