---
layout: page
author: Derek H. Ogle
title: MTH107 FAQ
subtitle: What does the warning from chisq.test() mean?
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/RFAQ/FAQs")
source("../../../rhelpers/rhelpers.R")
fnm <- "chisq-warning"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message='FALSE'}
source("../../../rhelpers/knitr_setup.R")
```

----

```{r echo=FALSE}
freq <- matrix(c(8,7,3,1),nrow=1)
colnames(freq) <- LETTERS[1:4]
```

In some instances, `chisq.test()` will return a warning such as that shown below.
```{r}
( chi1 <- chisq.test(freq) )
```

This warning occurs when one or more expected values is less than five.  This is confirmed by viewing the expected table.
```{r}
chi1$expected
```

In this example, each cell in the expected table is less than five (**note that this example is for a goodness-of-fit test but the concept is the same for a two-variable chi-square test**).

This problem may be corrected in two ways.  First, you may collect a larger sample size.  Second, some categories may be combined to reduce the number of cells in the table and, thus, increase the expected values.
