---
layout: page
author: Derek H. Ogle
title: MTH107 FAQ
subtitle: How do I install the NCStats package?
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/RFAQ/FAQs")
source("../../../rhelpers/rhelpers.R")
fnm <- "install_NCStats"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message='FALSE'}
source("../../../rhelpers/knitr_setup.R")
```

----

The `NCStats` package is a set of functions that I have written to make aspects of using R for statistics classes at Northland easier.  Descriptions for installing `NCStats` are below.  `NCStats` is likely loaded on all Northland College computers, though it may not be the most recent version at all times.

## Installing

As long as your computer is connected to the internet you can install `NCStats` directly from the RForge.net repository by running the following command in R.

```{r eval=FALSE}
source("http://www.rforge.net/NCStats/InstallNCStats.R")
```

## Checking the Installation

After installing the `NCStats`, it is loaded for use with ...
```{r}
library(NCStats)
```

If this command produces an error where R says that it cannot find another package, then see [this FAQ](install_other_packages.html).

The following command should open the help file and also indicates that the download and installation were successful.
```{r eval=FALSE}
?NCStats
```
