---
title: "Answer Key"
subtitle: "Hypothesis Testing CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/Hypothesis_Testing/")
source("../../rhelpers/rhelpers.R")
modHTML("CE_Keys")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../rhelpers/knitr_setup.R")
```

----

## Identifying Hypotheses

<ol>
<li>$H_{A}: \mu < 90$ vs. $H_{0}: \mu = 90$</li>
<li>$H_{A}: \mu > 15$ vs. $H_{0}: \mu = 15$</li>
<li>$H_{A}: \mu \neq 22500$ vs. $H_{0}: \mu = 22500$</li>
<li>$H_{A}: p > 0.61$ vs. $H_{0}: p = 0.61$</li>
</ol>

----

## P-Value Calculations and Decisions

<ol>
<li>The p-value is "the probabiity of observing a sample mean of 79.5 or greater if the population mean is 75." Reject $H_{0}$ because p-value=`r formatC(distrib(79.5,mean=75,sd=12/sqrt(50),lower.tail=FALSE),format="f",digits=4)`<&alpha;.</li>
```{r echo=FALSE}
distrib(79.5,mean=75,sd=12/sqrt(50),lower.tail=FALSE)
```

<li>The p-value is "the probabiity of observing a sample mean of 11.2 or 'different' if the population mean is 14." Reject $H_{0}$ because p-value=`r formatC(2*distrib(11.2,mean=14,sd=6/sqrt(25)),format="f",digits=4)`<&alpha;.</li>
```{r echo=FALSE, results='hide'}
2*distrib(11.2,mean=14,sd=6/sqrt(25))
```

<li>The p-value is "the probabiity of observing a sample mean of 918 or greater if the population mean is 880." Reject $H_{0}$ because p-value=`r formatC(distrib(918,mean=880,sd=90/sqrt(80),lower.tail=FALSE),format="f",digits=4)`<&alpha;.</li>

```{r echo=FALSE}
distrib(918,mean=880,sd=90/sqrt(80),lower.tail=FALSE)
```

<li>The p-value is "the probabiity of observing a sample mean of 13700 or less if the population mean is 15000." Do not reject $H_{0}$ because p-value=`r formatC(distrib(13700,mean=15000,sd=8000/sqrt(50)),format="f",digits=4)`>&alpha;.</li>

```{r echo=FALSE}
distrib(13700,mean=15000,sd=8000/sqrt(50))
```

<li>The p-value is "the probabiity of observing a sample mean of 824 or 'different' if the population mean is 800." Do not reject $H_{0}$ because p-value=`r formatC(2*distrib(824,mean=800,sd=60/sqrt(25),lower.tail=FALSE),format="f",digits=4)`>&alpha;.</li>

```{r echo=FALSE, results='hide'}
2*distrib(824,mean=800,sd=60/sqrt(25),lower.tail=FALSE)
```

<li>The p-value is "the probabiity of observing a sample mean of 73 or less if the population mean is 80." Reject $H_{0}$ because p-value=`r formatC(distrib(73,mean=80,sd=20/sqrt(30)),format="f",digits=4)`<&alpha;.</li>

```{r echo=FALSE}
distrib(73,mean=80,sd=20/sqrt(30))
```

<li>The p-value is "the probabiity of observing a sample mean of 1370 or 'different' if the population mean is 1500." Do not reject $H_{0}$ because p-value=`r formatC(2*distrib(1370,mean=1500,sd=800/sqrt(40)),format="f",digits=4)`>alpha;.</li>

```{r echo=FALSE, results='hide'}
2*distrib(1370,mean=1500,sd=800/sqrt(40))
```
</ol>

----

## Body Temperature
<ol>
<li>$H_{A}: \mu \neq 37$ vs. $H_{0}: \mu = 37$</li>
```{r echo=FALSE}
d <- read.csv("https://github.com/droglenc/NCData/raw/master/NormTemp.csv")
mn <- mean(d$bodytemp)
n <- nrow(d)
```
<li>x&#772;=`r formatC(mn,format="f",digits=3)`</li>
<li>p-value=`r formatC(2*distrib(round(mn,3),mean=37,sd=0.41/sqrt(n)),format="f",digits=9)`</li>
<li>Reject $H_{0}$ because p-value < alpha;.</li>
<li>Average body temperature is somewhat less (but significantly) less than 37C.</li>
</ol>

----

## Identifying Type I and II Errors

<ol>
<li>The $H_{0}: \mu=15.4$ and $H_{A}: \mu>15.4$.  A Type I error would be saying that the chicken is contaminated when it really is not contaminated.  A Type II error would be saying that the chicken is not contaminated when it really is contaminated.  In my mind, a Type I error is more egregious as people would be eating contaminated chicken that they have been told is not contaminated.</li>
<li>The $H_{0}$ is that the drug does not exceed specification whereas the $H_{A}$ is that the drug does exceed the specification.  A Type I error occurs if the drug does not exceed the specifications but it concluded that it does.  A Type II error occurs if the the drug does exceed the specifications but it is concluded that it does not.  In my mind, a Type II error is more egregious as people would be using a "bad" drug but would have been told that it was OK.</li>
</ol>
----

## Understanding Power Relationships

<ol>
<li>See below.</li>
```{r}
distrib(0.05,mean=100,sd=10/sqrt(30),type="q")
```

<li>See below.</li>
```{r}
distrib(96.997,mean=95,sd=10/sqrt(30))
```

<li>Effects on &beta;</li>
<ol>
<li>Decrease &beta;</li>
<li>Decrease &beta;</li>
<li>Increase &beta;</li>
<li>Decrease &beta;</li>
</ol>
<li>Effects on power</li>
<ol>
<li>Increase power</li>
<li>Increase power</li>
<li>Decrease power</li>
<li>Increase power</li>
</ol>
<li>Increase n, increase &alpha;</li>
<li>Increase n, increase &alpha;</li>
</ol>

----
