---
title: "Answer Key"
subtitle: "1-Sample Z Class Example"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_1SampleZ_CExmpl")
```
```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
```

----

## Sales Transactions
```{r GerminationCalc, echo=FALSE, results='hide', fig.show='hide'}
xbar <- 130.26; sigma <- 15; n <- 48; se <- sigma/sqrt(n)
mu0 <- 120; z <- round((xbar-mu0)/se,3)
pv <- distrib(z,lower.tail=FALSE)
zstar <- distrib(0.9,type="q",lower.tail=FALSE)
lcb <- xbar+zstar*se
```

<ol start="1">
<li>&alpha; = 0.10.</li>
<li>H<sub>A</sub>: &mu; > `r mu0` and H<sub>0</sub>: &mu; = `r mu0`, where &mu; is the mean sales transaction for this company following the training.</li>
</ol>

<div class="alert alert-warning">
<ul>
  <li>The H<sub>A</sub> is "greater than" because they are trying to see if the mean sales transaction amount increased.</li>
  <li>Remember to define the parameter.</li>
</ul>
</div>

<ol start="3">
  <li>A 1-sample Z-test will be used because</li>
  <ol type="i">
    <li>one group/population is considered (this company after the training),</li>
    <li>a quantitative variable (sales transaction) was recorded, and</li>
    <li>&sigma; is known (=15).</li>
  </ol>
</ol>

<div class="alert alert-warning">
<ul>
  <li>These are the three characteristics that separate a 1-sample Z-test from other tests we will see this semester.</li>
  <li>Please include some sort of description that shows that the characteristics are met and that you are not just listing the characteristics -- in other words, include the parts in parentheses.</li>
  <li>The &sigma; will be given in the background information NOT in the table of results from the sample. The standard deviation in the table of results from the sample is s, not &sigma;. If you know &sigma; (as here) then you should use it as it is "better" than s.</li>
</ul>
</div>

<ol start="4">
  <li>This study is observational as the researcher did not create groups or impose a treatment on any individuals. It is not clear or obvious whether the individuals were randomly selected or not.</li>
</ol>

<div class="alert alert-warning">
<ul>
  <li>The very vast majority of 1-sample tests will be observational, as experiments usually compare among treatments which requires two or more groups.</li>
  <li>Do not say that the collection was random if it is not obviously so. Ultimately, this is a problem because the p-value requires randomization. However, we will continue with the analysis in this class, but we will be aware of the potential problems that this poses.</li>
</ul>
</div>

<ol start="5">
  <li>The assumptions are met because (i) &sigma; is known (=15) and  (ii) n = 48 &ge; 30.</li>
</ol>  
  
<div class="alert alert-warning">
<ul>
  <li>If there had been less than 30 then you would need to know the shape (strongly skewed, normal, etc.) of the population distribution.</li>
  <li>Make sure to address whether &sigma; is known or not, even though it feels redundant with Step 3.</li>
</ul>
</div>

<ol start="6">
  <li>x&#772; = $`r xbar`.</li>
  <li>Z = $\frac{`r xbar` = `r mu0`}{\frac{`r sigma`}{\sqrt{`r n`}}}$ = $\frac{`r xbar-mu0`}{`r se`}$ = `r z`.</li>
  <li>p-value=`r kPvalue(pv,include.p=FALSE,latex=FALSE,digits=6)` as computed with `distrib(130.26,mean=120,sd=15/sqrt(48),lower.tail=FALSE)` or `distrib(4.739,lower.tail=FALSE)`.</li>
<li>Reject H<sub>0</sub> because the p-value &lt; &alpha;.</li>
<li>It appears that the mean sales transaction is greater than $120. Thus, it appears that the training did result in an increase in the mean sales transaction amount.</li>
</ol>

<div class="alert alert-warning">
  <ul>
  <li>Note how the conclusion is about the MEAN sales transaction amount and not the sales transaction amount. The test is about the summary (i.e., the MEAN) not about individuals.</li>
  <li>Also note the use of "appears" in the summary. This language acknowledges the fact that an error (Type I or Type II) could have been made. No conclusion from a statistical hypothesis test will ever be 100% definitive.</li>
  <li>For the sake of learning, if we had not rejected the H<sub>0</sub> then we would have concluded that the "mean transaction amount does not appear to be greater than $120". Note that we would not have said that the mean transaction amount was $120, because it is possible that it was less than $120.
  </ul>
</div>

<br>

<div class="alert alert-info">
<p>I performed the following steps or calculations in preparation for the next step.</p>
  <ul>
    <li>90% confidence (because &alpha;=0.1)</li>
    <li>Lower confidence bound because H<sub>A</sub> was "greater than."
    <li>Z<sup>*</sup>=`r formatC(zstar,format="f",digits=3)` as computed with `distrib(0.90,type="q",lower.tail=FALSE)`.
    <li>`r xbar` `r formatC(zstar,format="f",digits=3)`*`r formatC(se,format="f",digits=3)` (where `r formatC(se,format="f",digits=3)` is the SE computed for Step 7).
  </ul>
</div>

<ol start="11">
<li>I am 90% confident that the mean sales transaction amount for all sales after the training is greater than $`r formatC(lcb,format="f",digits=2)`. Thus, the mean transacgtion amount appears to have increased by at least $`r formatC(lcb-mu0,format="f",digits=2)`.
</ol>

<div class="alert alert-warning">
<ul>
  <li>This confidence region is a lower bound, so the conclusion is that the mean is greater than the calculated amount.</li>
  <li>The second sentence is a nice addition as it more directly answers the researcher's research hypothesis.</li>
</ul>
</div>
