---
title: "Answer Key"
subtitle: "Foundational Definitions CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_FoundationalDefns_CE")
```
```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
```

----

<div class="alert alert-warning">
<strong>Note:</strong> 
<ul>
<li>In the next two sections, the variable type must be one of continuous, discrete, nominal, or ordinal (saying quantitative or categorical is not specific enough).</li>
</ul>
</div>

## Types of Variables I

1. Miles driven per week is quantitative, continuous because another value CAN be found between any two miles driven values (e.g., a student can drive between 10 and 10.1 miles per week, between 10 and 10.01 miles per week, and so on).
1. Type of vehicle is categorical, nominal because the types represent groups but there is no order among the groups.
1. Signs of disease is categorical, nominal because "yes" and "no" are groups and a categorical variables with two levels is nominal by definition.
1. Number of fleas is quantitative, discrete because another value can NOT be found between all pairs of numbers of fleas (e.g., there cannot be between 10 and 11 fleas).
1. Relative health of a rabbit is categorical, ordinal because the health was organized by groups (e.g., excellent) and those groupings were ordered.
1. Species of rabbit is categorical, nominal because species is a type of group, but there is no order among those groups.
1. Number of correct answers is quantitative, discrete because a value can NOT be found between all pairs of correct answers (e.g., there cannot be between 19 and 20 correct answers).
1. Concentration of lead is quantitative, continuous because a value CAN be found between any pair of lead concentrations.
1. Risk potential is categorical, ordinal because a grouping was created (e.g., "high risk") and those groupings have an order.
1. Information medium is categorical, nominal because they are types, but those types do not have an ordering.
1. The Koppen scheme is definitely categorical and I think it is nominal because I don't see an ordering in the given categories (i.e., they don't appear to range from wet to dry, cold to hot, north to south).

----

## Types of Variables II

1. Answers will vary by student. See instructor if you have questions about your example.

----

<br>

<div class="alert alert-warning">
<strong>Note:</strong> 
<ul>
<li>The specificity in the description for the individual (usually specific about "when" and "where").</li>
<li>The use of "ALL" when describing the population and parameter.</li>
<li>The list of a specific number of individuals when describing the sample and statistic.</li>
<li>The use of a specific type of summary (almost always the mean/average or percentage/proportion) when describing the parameter and statistic. Further note that the same summary word is used for both the parameter and statistic.</li>
<li>A summary word usually does not appear in the description of the variable.</li>
<li>The variable type must be one of continuous, discrete, nominal, or ordinal (saying quantitative or categorical is not specific enough).</li>
</ul>
</div>

## Sustainability Survey

* I - A Northland College Student in Fall, 2013
* V - Number of miles driven per week
* Po - **ALL** Northland College Students in Fall, 2013
* Pa - *Average* number of miles driven per week by **ALL** Northland College Students in Fall, 2013
* Sa - 147 sampled Northland College students
* St - *Average* number of miles driven per week by 147 sampled Northland College Students in Fall, 2013 
* Variable Type - Miles driven per week is quantitative, continuous because another value can be found been any two miles driven.

----

## Snow in Duluth

* I - A December in Duluth, MN in recorded history
* V - Total snowfall amount
* Po - **ALL** Decembers in Duluth, MN in recorded history
* Pa -- *Average* total snowfall amount from **ALL** Decembers in Duluth, MN in recorded history
* Sa - 35 randomly sampled Decembers in Duluth, MN
* St - *Average* total snowfall amount for the 35 sampled Decembers
* Variable Type - Total snowfall is quantitative, continuous because a value can be found between any two snowfall amounts.

----

## Crayfish and Bass

* I - a Rusty Crayfish in the lake with Smallmouth Bass in 2015
* V - carapace length of Rusty Crayfish
* Po -**ALL** Rusty Crayfish in the lake with Smallmouth Bass in 2015
* Pa - *mean* carapace length of **ALL** Rusty Crayfish in the lake with Smallmouth Bass in 2015
* Sa - 235 Rusty Crayfish in the lake with Smallmouth Bass in 2015 that were actually examined
* St - *mean* carapace length of 235 Rusty Crayfish in the lake with Smallmouth Bass in 2015
* Variable Type - Carapace length is a continuous quantitative variable because a carapace length can be found between any two carapace lengths.
  
----

## Transfer of Invasive Species

* I - a ballast tank on an ocean-going vessel in 2001.
* V - whether or not the tank contained any living organisms.
* Po - **ALL** ballast tanks on ocean-going vessels in 2001.
* Pa - *proportion* of **ALL** ballast tanks that contained living organisms.
* Sa - 43 ballast tanks in 2001.
* St - *proportion* of 43 ballast tanks that contained living organisms.
* Variable Type - Whether or not the tank contains any living organisms is a nominal categorical variable.
  
----

## Playing Time Among NHL Rookies

* I - A NHL rookie drafted straight from college (... from professional league)
* V - Playing time (mins) per game
* Po - **ALL** NHL rookies drafted straight from college (... from professional league)
* Pa -- *Average* playing time per game by **ALL** NHL rookies drafted straight from college (... from professional leagues)
* Sa - 25 NHL rookies drafted straight from college (... from professional league)
* St - *Average* playing time per game by 25 NHL rookies drafted straight from college (... from professional leagues)
* Variable Type - Playing time is quantitative, continuous because an amount of playing time can be found between any two amounts of playing time (we could define time in smaller and smaller increments, we just choose not to).

----

## Public Attitudes Towards Wolves in Wisconsin

* I - A Wisconsin resident that had participated in the three previous surveys
* V - Response ("Strongly Agree", "Agree", "Neutral", "Disagree", or "Strongly Disagree") to "In general, the DNR is responsive to the concerns of most Wisconsin citizens."
* Po - **ALL** Wisconsin resident that had participated in the three previous surveys
* Pa - *Percent* of **ALL** Wisconsin residents that had participated in the three previous surveys who either agreed or strongly agreed with the question
* Sa - 773 sampled Wisconsin residents that had participated in the three previous surveys
* St - *Percent* of 773 sampled Wisconsin residents that had participated in the three previous surveys who either agreed or strongly agreed with the question
* Variable Type - The response to the question is categorical, ordinal because a ranking is asked for and those rankings have an order.

----
