---
title: "Answer Key"
subtitle: "Hypothesis Testing - Introduction CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_HypTesting1_CE1")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
library(captioner)
figs <- captioner(prefix="Figure")
figs(name="Pgreater","Normal distribution representing the p-value calculation for the first example.")
figs(name="Pless","Normal distribution representing the p-value calculation for the second example.")
figs(name="Pnot","Normal distribution representing the p-value calculation for the third example.")
```

----

## Identifying Hypotheses

<div class="alert alert-warning">
<strong>When identifying the hypotheses note:</strong> 
<ul>
  <li>Each statistical hypothesis must contain a parameter -- either &mu; or p. That parameter must be defined (note the "where ..." constructions below).</li>
  <li>Each statistical hypothesis must contain a specific value that the parameter is compared to.</li>
  <li>The null hypothesis (i.e., H<sub>0</sub>) must contain the equals sign. No other mathematical sign is allowed.</li>
  <li>The alternative hypothesis (i.e., H<sub>A</sub> must contain either the less than (<), greater than (>), or not equals (&ne;) sign. No other mathematical sign is allowed.</li>
</ul>
</div>

<ol>
<li>H<sub>A</sub> &mu; > 15 vs. H<sub>0</sub> &mu; = 15, where &mu; is the mean length of all oil changes at the Toyota dealership.</li>
<li>H<sub>A</sub> &mu; &ne; 22500 vs. H<sub>0</sub> &mu; = 22500, where &mu; is the mean cost of all bariatric surgeries.</li>
<li>H<sub>A</sub> p > 0.61 vs. H<sub>0</sub> p = 0.61, where p is the proportion of all Washburn High School students that are involved in at least one extracurricular activity.</li>
<li>H<sub>A</sub> &mu; < 90 vs. H<sub>0</sub> &mu; = 90, where &mu; is the mean number of murders for all burroughs in 2010.</li>
</ol>

----

## P-Value Calculations and Decisions I

<div class="alert alert-warning">
<strong>Note:</strong> 
<ul>
<li>When writing the definition of the p-value in the context of the problem you are replacing "the observed statistic" with the value of x&#772;; "more extreme" with "less", "greater", or "different" depending on the sign in H<sub>A</sub>; and the "null hypothesis is true" with the value of &mu; in "the probability of the observed statistic or a value more extreme assuming the null hypothesis is true."</li>
</ul>
</div>

<ol>
<li>The p-value is "the probability of observing a sample mean of 79.5 or greater if the population mean is 75." Reject H<sub>0</sub> because p-value=`r formatC(distrib(79.5,mean=75,sd=12/sqrt(50),lower.tail=FALSE),format="f",digits=4)`<&alpha; (`r figs("Pgreater",display="cite")`).</li>
```{r pgreater, echo=FALSE}
distrib(79.5,mean=75,sd=12/sqrt(50),lower.tail=FALSE)
```

`r figs("Pgreater")`

<li>The p-value is "the probability of observing a sample mean of 13700 or less if the population mean is 15000." Do not reject H<sub>0</sub> because p-value=`r formatC(distrib(13700,mean=15000,sd=8000/sqrt(50)),format="f",digits=4)`>&alpha; (`r figs("Pless",display="cite")`).</li>

```{r pless, echo=FALSE}
distrib(13700,mean=15000,sd=8000/sqrt(50))
```

`r figs("Pless")`

<li>The p-value is "the probability of observing a sample mean of 11.2 or 'different' if the population mean is 14." Reject H<sub>0</sub> because p-value=`r formatC(2*distrib(11.2,mean=14,sd=6/sqrt(25)),format="f",digits=4)`<&alpha; (`r figs("Pnot",display="cite")`).</li>
```{r pnot, echo=FALSE, results='hide'}
2*distrib(11.2,mean=14,sd=6/sqrt(25))
```

`r figs("Pnot")`

</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
<<pgreater>>
<<pless>>
<<pnot>>
```
