---
title: "Answer Key"
subtitle: "Hypothesis Testing Errors CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_HypTestingErrs_CE")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
library(captioner)
figs <- captioner(prefix="Figure")
figs(name="Pwr1","NULL normal distribution representing the rejection region for the power calculation example.")
figs(name="Pwr2","ACTUAL normal distribution representing power in the power calculation example.")
```

----

## Identifying Type I and II Errors

<ol>
<li>The H<sub>0</sub>: &mu;=15.4 and H<sub>A</sub>: &mu;>15.4. A Type I error would be saying that the chicken is contaminated when it really is not contaminated. A Type II error would be saying that the chicken is not contaminated when it really is contaminated. In my mind, a Type II error is more egregious as people would be eating contaminated chicken that they have been told is not contaminated.</li>
<li>The H<sub>0</sub> is that the drug does not exceed specification whereas the H<sub>A</sub> is that the drug does exceed the specification. A Type I error occurs if the drug does not exceed the specifications but it is concluded that it does. A Type II error occurs if the the drug does exceed the specifications but it is concluded that it does not. In my mind, a Type II error is more egregious as people would be using a "bad" drug but would have been told that it was OK.</li>
</ol>

----

## Understanding Power Relationships

<ol>
<li>See `r figs("Pwr1",display="cite")`.</li>
```{r Pwr1,echo=FALSE}
distrib(0.05,mean=100,sd=10/sqrt(30),type="q")
```

`r figs("Pwr1")`

<li>See `r figs("Pwr2",display="cite")`.</li>
```{r Pwr2,echo=FALSE}
distrib(96.997,mean=95,sd=10/sqrt(30))
```

`r figs("Pwr2")`

<li>Effects on power</li>
<ol>
<li>Increasing &alpha; will INcrease power</li>
<li>Increasing n will INcrease power</li>
<li>Increasing &sigma; will DEcrease power</li>
<li>Increasing how wrong the H<sub>0</sub> is will INcrease power</li>
</ol>
<li>Effects on &beta;</li>
<ol>
<li>Increasing &alpha; will DEecrease &beta;</li>
<li>Increasing n will DEcrease &beta;</li>
<li>Increasing &sigma; will INcrease &beta;</li>
<li>Increasing how wrong the H<sub>0</sub> is will DEcrease &beta;</li>
</ol>
<li>Power may be INcreased by INcreasing n or INcreasing &alpha;.</li>
<li>&beta; my be DEcreased by INcreasing n or INcreasing &alpha;.</li>
</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
<<Pwr1>>
<<Pwr2>>
```

----
