---
title: "Answer Key"
subtitle: "Summaries for One Quantitative Variable CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_RUnivEDA_CE1")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
library(ggplot2)
source("../../../rhelpers/knitr_setup.R")
```

<br>

# Coarse Woody Debris

1. The vast majority of sites (82%) were exposed to medium levels of wind (as compared to low levels of wind).
1. The diameter of coarse woody debris in the north basin of Allequash Lake is strongly right-skewed with four potential outliers above 30 cm, a median (center) of 19.5 cm, and an IQR from Q1 of 16.2 to a Q3 of 20.8 cm. The median and IQR were used because of the strongly skew and potential outliers.

#### R Code and Results
```{r echo=FALSE}
cwdobj <- read.csv("https://raw.githubusercontent.com/droglenc/NCMTH107/gh-pages/modules/CE/1_CE_KEYS/cwd.csv")
```
```{r eval=FALSE}
cwdobj <- read.csv("cwd.csv")
```
```{r}
str(cwdobj)
```
```{r}
( freq1 <- xtabs(~exposure,data=cwdobj) )
percTable(freq1)
ggplot(data=cwdobj,mapping=aes(x=exposure)) +
  geom_bar(color="black",fill="lightgray") +
  labs(x="Wind Expsure Category",y="Frequency of Sites") +
  scale_y_continuous(expand=expansion(mult=c(0,0.05))) +
  theme_NCStats()
```
```{r}
Summarize(~diameter,data=cwdobj,digits=1)
ggplot(data=cwdobj,mapping=aes(x=diameter)) +
  geom_histogram(binwidth=1,boundary=0,color="black",fill="lightgray") +
  labs(x="CWD Diameter (cm)",y="Frequency of Sites") +
  scale_y_continuous(expand=expansion(mult=c(0,0.05))) +
  theme_NCStats()
```

<br>

----

# Brain Weights

1. The distribution of head sizes for the 20-46 age-group is very slightly right-skewed with no obvious outliers, a mean of 3675, and a standard deviation of  364. The mean and standard deviation were used because the shape was not strongly skewed and there were not obvious outliers.
1. The distribution of head sizes for the 46+ age-group is very slightly right-skewed with no obvious outliers, a mean of 3598, and a standard deviation of  364. The mean and standard deviation were used because the shape was not strongly skewed and there were not obvious outliers.

#### R Code and Results
```{r echo=FALSE}
bh <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/BrainHead.csv")
```
```{r eval=FALSE}
bh <- read.csv("Brainhead.csv")
```
```{r}
str(bh)
```
```{r}
Summarize(head.size~age.group,data=bh,digits=1)
ggplot(data=bh,mapping=aes(x=head.size)) +
  geom_histogram(binwidth=100,boundary=0,color="black",fill="lightgray") +
  labs(x="Head Size (cm^3)",y="Frequency of Subjects") +
  scale_y_continuous(expand=expansion(mult=c(0,0.05))) +
  theme_NCStats() +
  facet_wrap(vars(age.group))

```

<br>

----

# Water Usage
Most students either let the water run for 6-10 (37.7%) or 11-15 minutes (40.0%).

#### R Code and Results
```{r echo=FALSE}
wtr <- read.csv("https://raw.githubusercontent.com/droglenc/NCMTH107/gh-pages/modules/CE/1_CE_KEYS/shower.csv")
```
```{r eval=FALSE}
wtr <- read.csv("shower.csv")
```
```{r}
str(wtr)
```
```{r}
( freq1 <- xtabs(~water.use,data=wtr) )
percTable(freq1)
ggplot(data=wtr,mapping=aes(x=water.use)) +
  geom_bar(color="black",fill="lightgray") +
  labs(x="Water Usage Category",y="Frequency of Students") +
  scale_y_continuous(expand=expansion(mult=c(0,0.05))) +
  theme_NCStats()
```
