---
title: "Answer Key"
subtitle: "1-Sample t CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE, warning=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_1Samplet_CE")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
library(captioner)
tbls <- captioner(prefix="Table")
figs <- captioner(prefix="Figure")
tbls(name="pH","Results from the one-sample t-test for testing that the pH is less than 5.6.")
tbls(name="Ozone","Results from the one-sample t-test for testing that the mean thickness of the ozone layer is different than 300 Dobson units.")
tbls(name="FC","Results from the one-sample t-test for testing that the mean CFU per 100 ml of lake water is greater than 150 CFU per 100 ml.")
figs(name="FChist","Histogram of the CFU per 100 ml in samples of lake water.")
```

----

## P-Value Calculations and Decisions I

<ol>
<li>Reject H<sub>0</sub> because p-value=`r formatC(2*distrib((11.2-14)/(6.1/sqrt(25)),df=24,distrib="t"),format="f",digits=4)`<&alpha;.</li>
```{r echo=FALSE, par1a=TRUE}
distrib((11.2-14)/(6.1/sqrt(25)),df=24,distrib="t")
```

<li>Reject H<sub>0</sub> because p-value=`r formatC(distrib((79.5-75)/(11.7/sqrt(50)),df=49,lower.tail=FALSE,distrib="t"),format="f",digits=4)`<&alpha;.</li>
```{r echo=FALSE, par1a=TRUE}
distrib((79.5-75)/(11.7/sqrt(50)),df=49,lower.tail=FALSE,distrib="t")
```
</ol>

----

## Confidence Interval Calculations I

<ol>
<li>See below.</li>
<ol type="a">
<li>C=0.95</li>
<li>Interval</li>
```{r echo=FALSE, results='hide', fig.show='hide'}
tstar <- distrib(0.975,df=24,type="q",distrib="t")
ci <- 11.2+c(-1,1)*tstar*6.1/sqrt(25)
```
<li>t*=&plusmn;`r formatC(tstar,format="f",digits=3)`.</li>
```{r echo=FALSE, par1a=TRUE}
distrib(0.975,df=24,type="q",distrib="t")
```
<li>(`r formatC(ci[1],format="f",digits=1)`,`r formatC(ci[2],format="f",digits=1)`).</li>
<li>I am 95% confident that the true population mean is between `r formatC(ci[1],format="f",digits=1)` and `r formatC(ci[2],format="f",digits=1)`.</li>
</ol>

<li>See below.</li>
<ol type="a">
<li>C=0.90</li>
<li>Lower bound</li>
```{r echo=FALSE, results='hide', fig.show='hide'}
tstar <- distrib(0.9,df=49,type="q",lower.tail=FALSE,distrib="t")
lcb <- 79.5+tstar*11.7/sqrt(50)
```
<li>t*=`r formatC(tstar,format="f",digits=3)`.</li>
```{r echo=FALSE, par1a=TRUE}
distrib(0.9,df=49,type="q",lower.tail=FALSE,distrib="t")
```
<li>`r formatC(lcb,format="f",digits=1)`.</li>
<li>I am 90% confident that the true population mean is greater than `r formatC(lcb,format="f",digits=1)`.</li>
</ol>
</ol>

----

<div class="alert alert-warning">
<strong>Note:</strong> 
<ul>
<li>Make sure to describe your evidence in Step 3 for why you are performing a 1-sample t-test. It is not adequate to simply say "because there is a quantitative response variable, only one group is being tested, and sigma is unknown." At least say what the response variable is and what the one groups is.</li>
<li>Also make sure to describe your evidence in Step 5 for why the assumptions are met. It is not adequate to simply say "n&ge;40" or "n&ge;15 and the sample distribution is not strongly skewed." You should state the actual n and direct the reader to how you know the distribution is not strongly skewed (i.e., likely a histogram).</li>
<li>In hand calculations, note that the test statistic (t in Step 7) is NOT the critical value (i.e., t<sup>*</sup>) needed for the confidence region (in Step 11). The t you calculate in Step 7 has nothing to do with the t<sup>*</sup> you will find in Step 11.</li>
<li>When computing the t in Step 7, the numerator is always x&#772; minus the specific value in H<sub>0</sub> from Step 2.</li>
<li>The p-value in this example is shaded to the left of the t from Step 7 and the result from `distrib()` is multiplied by 2 because the H<sub>A</sub> is a not equals. If the t had been positive you would have shaded to the right and then muliplied by 2. When the H<sub>A</sub> is a not equals you always shaded into the tail nearest t and multiply the `distrib()` result by 2.</li>
<li>In Steps 10 and 11 don't say &mu;. Rather replace it with &mu; means. See how those steps below say "mean CFU for ALL 100 ml of water" rather than just &mu;.</li>
<li>Also in Step 10 you can be more specific then saying that the mean CFL differs from 178. Because the x&#772; is less than what was hypothesized you can say that the mean CFL differs from 178 by being less than 178. So, even though you strictly only tested for a difference in this not equal situation, you can be more specific with your statement in Step 10.</li>
<li>You values may differ slightly from mine as I calculate everything in R that holds many decimal places on intermediate values.</li>
</ul>
</div>

## Fecal Coliform Counts

```{r CFU1Calc, echo=FALSE, results='hide', fig.show='hide'}
xbar <- 168.19; s <- 22.20; n <- 25
se <- s/sqrt(n)
( t <- (xbar-178)/se )
( df <- n-1 )
( pv <- 2*distrib(t,distrib="t",df=df) )
( tstar <- distrib(0.975,type="q",distrib="t",df=df) )
( tstar <- c(-1,1)*tstar )
( ci <- xbar+tstar*se )
```

<ol>
<li>&alpha;=0.05.</li>
<li>H<sub>A</sub>: &mu;&ne;178 and H<sub>0</sub>: &mu;=178, where &mu; is the mean number of CFU per 100 ml in ALL samples from this lake on this date.</li>
<li>1-sample t-test because (i) one group or population (this lake) was sampled, (ii) a quantitative variable (number of CFU per 100 ml) was recorded, and (iii) &sigma; is not known.</li>
<li>An observational study without clear randomization of 25 water samples.</li>
<li>The assumptions are met because (i) &sigma; is not known and (ii) n=25&ge;15 and the histogram (given in question) is not strongly skewed.</li>
<li>x&#772;=`r xbar` CFU per 100 ml.</li>
<li>t=$\frac{168.19-178}{\frac{22.20}{\sqrt{25}}}$=`r formatC(t,format="f",digits=3)` with `r df` df.</li>
<li>p-value=`r kPvalue(pv,include.p=FALSE,latex=FALSE)`.</li>
<li>Reject H<sub>0</sub> because the p-value<&alpha;.</li>
<li>It appears that the mean CFU per ALL 100 ml of water sampled from this lake on this date is less than what the computer model predicted.</li>
<li>One can be 95% confident that the mean CFU per ALL 100 ml of water sampled from this lake on this date is between 168.19-2.064$\frac{22.20}{\sqrt{25}}$=`r formatC(ci[1],format="f",digits=0)` and 168.19+2.064$\frac{22.20}{\sqrt{25}}$=`r formatC(ci[2],format="f",digits=0)`, which further indicates that the mean CFU per ALL 100 ml of water is less than 178.</li>
</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
<<CFU1Calc>>
```

<br>

----

## Increased Civility in Nation-States?
```{r KillingsCalc, echo=FALSE, results='hide', fig.show='hide'}
xbar <- 3292.8; s <- 4814.5; n <- 75
( se <- s/sqrt(n) )
( t <- (xbar-4104)/se )
( df <- n-1 )
( pv <- distrib(t,distrib="t",df=df) )
( tstar <- distrib(0.95,type="q",distrib="t",df=df) )
( ucb <- xbar+tstar*se )
```

<ol>
<li>&alpha;=0.05.</li>
<li>H<sub>A</sub>: &mu;<4104 and H<sub>0</sub>: &mu;=4104, where &mu; is the mean number of citizens killed by ALL nation-states.</li>
<li>1-sample t-test because (i) one group or population (all nation-states) was sampled, (ii) a quantitative variable (number of killings) was recorded, and (iii) &sigma; is not known.</li>
<li>An observational study with clear randomization of 75 countries.</li>
<li>The assumptions are met because (i) &sigma; is not known and (ii) n=75&ge;40.</li>
<li>x&#772;=`r xbar` killings.</li>
<li>t=$\frac{3292.8-4104}{\frac{4814.5}{\sqrt{75}}}$=`r formatC(t,format="f",digits=3)` with `r df` df.</li>
<li>p-value=`r kPvalue(pv,include.p=FALSE,latex=FALSE)`.</li>
<li>Do not reject H<sub>0</sub> because the p-value>&alpha;.</li>
<li>It appears that the mean number of killings for recent times is NOT significanlty lower than it was as determined from 19th century history books.</li>
<li>One is 95% confident that the number of killings in recent years for ALL nation-states is less than $3292.8+1.666\frac{4814.5}{\sqrt{75}}$=`r formatC(ucb,format="f",digits=0)`, which further indicates that the mean number of killings is not less than 4104.</li>
</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
<<KillingsCalc>>
```
