---
title: "Answer Key"
subtitle: "Normal Distribution Class Example"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_NormalDist2_CExmpl")
```
```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
```

----

## Wood Duck Weights

<ol start="1">
<li>The percentage of wood ducks that weigh more than 575 g is `r formatC(distrib(575,mean=450,sd=50,lower.tail=FALSE)*100,format="f",digits=1)`%.</li>
```{r}
distrib(575,mean=450,sd=50,lower.tail=FALSE)
```
</ol>


<ol start="2">
<li>The percentage of wood ducks that weigh less than 420 g is `r formatC(distrib(420,mean=450,sd=50)*100,format="f",digits=1)`%.</li>
```{r}
distrib(420,mean=450,sd=50)
```
</ol>


<ol start="3">
<li>The percentage of wood ducks that weigh betwen 360 and 480 g is `r formatC(distrib(480,mean=450,sd=50)*100-distrib(360,mean=450,sd=50)*100,format="f",digits=1)`%.</li>
```{r fig.show='hold'}
( ab <- distrib(480,mean=450,sd=50) )
( a <- distrib(360,mean=450,sd=50) )
ab-a
```
</ol>


<ol start="4">
<li>The lightest 10% of wood ducks weigh less than `r formatC(distrib(0.10,mean=450,sd=50,type="q"),format="f",digits=0)` g.</li>
```{r}
distrib(0.10,mean=450,sd=50,type="q")
```
</ol>


<ol start="5">
<li>The heaviest 25% of wood ducks weigh more than `r distrib(0.25,mean=450,sd=50,type="q",lower.tail=FALSE)` g</li>
```{r}
distrib(0.25,mean=450,sd=50,type="q",lower.tail=FALSE)
```
</ol>
