---
title: "Answer Key"
subtitle: "Foundational Definitions CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_FoundationalDefns_CE")
```
```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
```

----

<div class="alert alert-warning">
<strong>Note:</strong> 
<ul>
<li>The variable type must be one of continuous, discrete, nominal, or ordinal (saying quantitative or categorical is not specific enough).</li>
</ul>
</div>

## Types of Variables I

1. Miles driven per week is quantitative, **continuous** because another value CAN be found between any two miles driven values (e.g., a student can drive between 10 and 10.1 miles per week, between 10 and 10.01 miles per week, and so on).
1. Type of vehicle is categorical, **nominal** because the types represent groups but there is no order among the groups.
1. Signs of disease is categorical, **nominal** because "yes" and "no" are groups and a categorical variables with two levels is nominal by definition.
1. Number of fleas is quantitative, **discrete** because another value can NOT be found between all pairs of numbers of fleas (e.g., there cannot be between 10 and 11 fleas).
1. Relative health of a rabbit is categorical, **ordinal** because the health was organized by groups (e.g., excellent) and those groupings were ordered.
1. Species of rabbit is categorical, **nominal** because species is a type of group, but there is no order among those groups.
1. Number of correct answers is quantitative, **discrete** because a value can NOT be found between all pairs of correct answers (e.g., there cannot be between 19 and 20 correct answers).
1. Concentration of lead is quantitative, **continuous** because a value CAN be found between any pair of lead concentrations.
1. Risk potential is categorical, **ordinal** because a grouping was created (e.g., "high risk") and those groupings have an order.
1. Information medium is categorical, **nominal** because they are types, but those types do not have an ordering.
1. The Koppen scheme is definitely categorical and I think it is **nominal** because I don't see an ordering in the given categories (i.e., they don't appear to range from wet to dry, cold to hot, north to south).

----

<br>

<div class="alert alert-warning">
<strong>Note:</strong> 
<ul>
<li>The specificity in the description for the individual (usually specific about "when" and "where").</li>
<li>The use of "ALL" when describing the population and parameter.</li>
<li>The list of a specific number of individuals when describing the sample and statistic.</li>
<li>The use of a specific type of summary (almost always the mean/average or percentage/proportion) when describing the parameter and statistic. Further note that the same summary word is used for both the parameter and statistic.</li>
<li>A summary word usually does not appear in the description of the variable.</li>
<li>The variable type must be one of continuous, discrete, nominal, or ordinal (saying quantitative or categorical is not specific enough).</li>
</ul>
</div>

## Sustainability Survey

* I - A Northland College Student in Fall, 2013
* V - Number of miles driven per week
* Po - **ALL** Northland College Students in Fall, 2013
* Pa - *Average* number of miles driven per week by **ALL** Northland College Students in Fall, 2013
* Sa - 147 sampled Northland College students
* St - *Average* number of miles driven per week by 147 sampled Northland College Students in Fall, 2013 
* Variable Type - Miles driven per week is quantitative, continuous because another value can be found been any two miles driven.

----

## Crayfish and Bass

* I - a Rusty Crayfish in the lake with Smallmouth Bass in 2015
* V - carapace length of Rusty Crayfish
* Po -**ALL** Rusty Crayfish in the lake with Smallmouth Bass in 2015
* Pa - *mean* carapace length of **ALL** Rusty Crayfish in the lake with Smallmouth Bass in 2015
* Sa - 235 Rusty Crayfish in the lake with Smallmouth Bass in 2015 that were actually examined
* St - *mean* carapace length of 235 Rusty Crayfish in the lake with Smallmouth Bass in 2015
* Variable Type - Carapace length is a continuous quantitative variable because a carapace length can be found between any two carapace lengths.

----

## Types of Variables II

1. Answers will vary by student. See instructor if you have questions about your example.
 
