---
title: "Answer Key"
subtitle: "Getting Data Into R CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_GettingDataIntoR_CE")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
```

----

The following two commands were run before the code run below because I saved all of my files in the same directory.
```{r eval=FALSE}
library(NCStats)
setwd("T:/Stats")
```

<div class="alert alert-warning">
<strong>Note:</strong> The object name to the left of each assignment operator (i.e., `<-`) can largely be whatever you want. In the examples below, I tried to uses names that made sense for the data being imported.
</div>
----

## Loading Data from Webpage I
#### **BrainHead.csv**
```{r echo=FALSE}
bhobj <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/BrainHead.csv")
```

1. There are `r kCounts(ncol(bhobj))` variables in BrainHead.csv.
1. The variables in Brainhead.csv are: `r paste(names(bhobj),collapse=", ")`.
1. There are `r kCounts(nrow(bhobj))` individuals in BrainHead.csv.

```{r eval=FALSE}
bhobj <- read.csv("BrainHead.csv")
str(bhobj)
```

----

## Loading Your Own Data I

* Answers will vary by student.

----

## Loading Data from Webpage II

#### **PopularKids.csv**
```{r echo=FALSE}
pkobj <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/PopularKids.csv")
```

1. There are `r kCounts(ncol(pkobj))` variables in PopularKids.csv.
1. The variables in PopularKids.csv are: `r paste(names(pkobj),collapse=", ")`.
1. There are `r kCounts(nrow(pkobj))` individuals in PopularKids.csv.

```{r eval=FALSE}
pkobj <- read.csv("PopularKids.csv")
str(pkobj)
```

#### **Yahara.csv**
```{r echo=FALSE}
yobj <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/Yahara.csv")
```

1. There are `r kCounts(ncol(yobj))` variables in Yahara.csv.
1. The variables in Yahara.csv are: `r paste(names(yobj),collapse=", ")`.
1. There are `r kCounts(nrow(yobj))` individuals in Yahara.csv.

```{r eval=FALSE}
yobj <- read.csv("Yahara.csv")
str(yobj)
```

----

## Loading Your Own Data II

* Answers will vary by student.

----
