---
title: "Answer Key"
subtitle: "R Bivariate EDA Assignment"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
---

```{r echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_RBivEDA_CE")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
library(ggplot2)
source("../../../rhelpers/knitr_setup.R")
```

<br>

# Urban Runoff

1. The response variable is IBI.
1. The explanatory variable is percent impervious surface.
1. The relationship between IBI and the percent of impervious surface is **negative, nonlinear**, and **moderately strong** with **no obvious outliers.** I did not report the value of the correlation coefficient because of the nonlinear form.

#### R Code and Results
```{r echo=FALSE}
d <- data.frame(imp=c(60,43,43,34,27,25,21,18,11,8,8,8,7,7,5,4,2),
                IBI=c(9,11,13,23,31,31,21,23,27,37,39,29,31,43,33,35,37))
```
```{r eval=FALSE}
d <- read.csv("IBI.csv")
```
```{r}
ggplot(data=d,mapping=aes(x=imp,y=IBI)) +
  geom_point(pch=21,color="black",fill="lightgray") +
  labs(x="Percent Impervious Land",y="Benthic Index of Biotic Integrity") +
  theme_NCStats()
```

<br>

----

## Lights and Nearsightedness
```{r echo=FALSE}
d <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/nightlight.csv")
```
```{r LightTables, echo=FALSE, results='hide'}
tbl <- xtabs(~Light+Nearsightedness,data=d)
addmargins(tbl)
( row.tbl <- percTable(tbl,margin=1) )
( col.tbl <- percTable(tbl,margin=2) )
( perc.tbl <- percTable(tbl) )
```

1. The percentage of children that slept with a "night light" that did not develop nearsightedness is **`r formatC(row.tbl["night light","No"],format="f",digits=1)`%** (see row percent table).
1. The percentage of all children that slept with a "lamp" and developed nearsightedness is **`r formatC(perc.tbl["lamp","Yes"],format="f",digits=1)`%** (see total percent table).
1. The percentage of children that slept in "no light" conditions that then developed nearsightedness is **`r formatC(row.tbl["no light","Yes"],format="f",digits=1)`%** (see row percent table).
1. A total of **`r tbl["no light","Yes"]`** children slept in "no light" conditions and developed nearsightedness (see frequency table).
1. The percentage of children that developed nearsightedness is **`r formatC(perc.tbl["Sum","Yes"],format="f",digits=1)`%** (see total percent table).
1. The percentage of children that developed nearsightedness that slept with a "lamp" is **`r formatC(col.tbl["lamp","Yes"],format="f",digits=1)`%** (see column percent table).
1. It appears that the percentage of children that developed nearsightedness is greater when the child slept with some sort of light (either a lamp or a night light), with a somewhat greater prevalence of nearsightedness with the lamp (see row percent table).

#### R Code and Results
```{r eval=FALSE}
d <- read.csv("Nearsight.csv")
```
```{r}
<<LightTables>>
```

<br>

----

<!----
## Forest Composition
```{r echo=FALSE}
d <- read.csv("https://github.com/droglenc/NCData/raw/master/Fire.csv")
d$sttype <- factor(d$sttype,levels=c("c","mc","md","d"))
```
```{r ForestTables, echo=FALSE, results='hide'}
tbl <- xtabs(~sttype+tslf,data=d)
addmargins(tbl)
( row.tbl <- percTable(tbl,margin=1) )
( col.tbl <- percTable(tbl,margin=2) )
( perc.tbl <- percTable(tbl) )
```

1. A total of **`r sum(tbl[,"200-300"])`** stands last had a fire between 200 and 300 years ago.
1. The percentage of deciduous tree stands that had a fire within the last 100 years is **`r formatC(row.tbl["d","0-100"],format="f",digits=1)`%** (see row percent table).
1. The percentage of mixed-coniferous tree stands that had a fire between 200 and 300 years ago is **`r formatC(row.tbl["mc","200-300"],format="f",digits=1)`%** (see row percent table).
1. The percentage of all stands that were coniferous and had a fire within the last 100 years is **`r formatC(perc.tbl["c","200-300"],format="f",digits=1)`%** (see total percent table).
1. The percentage of stands that last had a fire 200 to 300 years ago that were coniferous is **`r formatC(col.tbl["c","200-300"],format="f",digits=1)`%** (see column percent table).
1. The percentage of all stands that were deciduous and had a fire between 100 and 200 years ago is **`r formatC(perc.tbl["d","100-200"],format="f",digits=1)`%** (see total percent table).
1. The percentage of mixed-deciduous tree stands that had a fire within the last 100 years is **`r formatC(row.tbl["md","0-100"],format="f",digits=1)`%** (see row percent table).
1. The recency of the last fire decreases as the "amount" of coniferous trees decreases (i.e., from coniferous to mixed-coniferous to mixed-deciduous to deciduous). In other words, stands with more deciduous trees last had a fire more recently.

#### R Code and Results
```{r eval=FALSE}
d <- read.csv("Fire.csv")
```
```{r}
<<ForestTables>>
```

<br>
---->
<!----

----

# Red-Imported Fire Ants and Deer
The relationship between the abundance of red-imported fire ants and fawn recruitment is **linear, negative, moderately strong** (r=-0.700), and **without any obvious outliers**. I assessed the strength of the relationship with the correlation coefficient (r) because the form is linear and there are no outliers.

#### R Code and Results
```{r echo=FALSE}
d <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/RIFA.csv")
```
```{r eval=FALSE}
d <- read.csv("RIFA.csv")
```
```{r}
ggplot(data=d,mapping=aes(x=rifa,y=fawnrec)) +
  geom_point(pch=21,color="black",fill="lightgray") +
  labs(x="Red-Imported Fire Ant Abundance",y="Fawn Recruitment Index") +
  theme_NCStats()
corr(~rifa+fawnrec,data=d,digits=4)
```
--->

----

# Wolves and Whitetail Deer
The relationship between the territory size of wolf pakces and the abundance of deer is **linear, negative, moderately strong** (r=-0.787), and **without any obvious outliers**. I assessed the strength of the relationship with the correlation coefficient (r) because the form is linear and there are no outliers.

#### R Code and Results
```{r echo=FALSE}
d <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/Wolves2.csv")
```
```{r eval=FALSE}
d <- read.csv("Wolves2.csv")
```
```{r}
ggplot(data=d,mapping=aes(x=deer,y=terr)) +
  geom_point(pch=21,color="black",fill="lightgray") +
  labs(x="Abundance of Deer (#/km^2)",y="Wolf Territory Size (km^2") +
  theme_NCStats()
corr(~deer+terr,data=d,digits=3)
```
