---
title: "Answer Key"
subtitle: "Why Statistics CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_WhyStats_CE")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
set.seed(388293)
library(captioner)
tbls <- captioner(prefix="Table")
tbls(name="Frogs1","Visual and call counts of frogs on the first sample of 10 lakes from northern Wisconsin.")
tbls(name="Frogs2","Visual and call counts of frogs on the second sample of 10 lakes from northern Wisconsin.")
```

----

## Realities

<div class="alert alert-warning">
<strong>Note:</strong> 
<ul>
<li>Provide some context for your answers. For example, don't just say "variability is everywhere." Rather say something like "Variability is everywhere, specifically among individuals and among summaries computed from different samples." Similarly, don't just say  "conclusions are uncertain." Rather say "conclusions about the population from information in the sample are uncertain."</li>
</ul>
</div>

1. The three "realities" that, if they did not exist, would eliminate the need for the field of statistics are that (i) you cannot "see" the entire population and must, therefore, take a sample; (ii) variability is "everywhere," both among individuals and among statistics; and (iii) conclusions about the population from information in the sample are uncertain.

<br>

----

## Variability

<div class="alert alert-warning">
<strong>Note:</strong> 
<ul>
<li>In the second question.</li>
<ul>
<li>Be specific to your example, don't just say the "individuals differ" or the "samples differ." For example, it is not adequate to say that each Bluegill will be different. Instead, be specific, and say (any one of these) each Bluegill will have a different length, different weight, or be of a different sex.</li>
<li>Your description of sampling variability should refer to a <strong>summary</strong> measure, usually the mean/average or percentage/proportion. For example, don't say that two samples of Bluegills will differ. Rather, say (any one of these) two samples of Bluegill will have different mean lengths, mean weights, or percentage of females.</li>
<li>When describing sampling variability make sure that you describe <strong>repeated</strong> samples from the <strong>same</strong> population. For example, it would be incorrect to say that sampling variability is illustrated by the fact that Bluegill sampled from Lake A had a different mean length than Bluegill sampled from Lake B because Lake A and Lake B are not repeated samples from the same population; rather they are two different populations.</li>
<li>The answer below was just one example. Your situation will likely differ.</li>
</ul>
</ul>
</div>

1. Natural variability is the realization that no two individuals are exactly identical. Sampling variability is the realization that the results (statistics) from any two samples are not exactly identical.
1. Consider a situation where a biologist is sampling Bluegill from a particular lake. Natural variability is the realization that no two Bluegill from this population will be exactly identical; i.e., their lengths, weights, number of scales, etc. will be different. Sampling variability is the realization that the results from any two samples of Bluegill from this population will not be identical. For example, the mean lengths from two samples of Bluegill from this population may differ, or the mean weights, or the mean number of scales, or the percentage of sexually mature fish.

<br>

----

## Wisconsin Lakes
```{r Lakes, echo=FALSE, results='hide'}
named <- filterD(ABDLakes,named)
smpl1 <- srsdf(named,n=50,vars=c("county","area"))
smpl2 <- srsdf(named,n=50,vars=c("county","area"))
```

1. The sample of 50 lakes is not shown (to save space). The area of the first three lakes were all different, which is an example of natural variability.
1. The proportion of lakes in my first sample that were from Bayfield County was `r formatC(prop.table(xtabs(~county,data=smpl1))[["Bayfield"]],format="f",digits=3)`.
1. The proportion of lakes in my second sample that were from Bayfield County was `r formatC(prop.table(xtabs(~county,data=smpl2))[["Bayfield"]],format="f",digits=3)`. These sample proportions differ, which is an example of sampling variability.
1. My two sample proportions did not equal the population proportions. I was not surprised because samples do not perfectly represent the entire population. I would expect the statistics from the sample to be near (on average) the population parameter, but not always equal to it.

<br>

----

## Frog Survey
```{r Frogs, echo=FALSE,results="hide",message=FALSE}
( sampA <- srsdf(Frogs,n=10) )
Summarize(~visual,data=sampA,digits=1)
( sampB <- srsdf(Frogs,n=10) )
Summarize(~visual,data=sampB,digits=1)
```

1. An individual is a lake in northern Wisconsin (north of Highway 8) in 2010.
1. No, it is not reasonable to count frogs on every lake in northern Wisconsin because there is no reasonable way to get to every lake in the northern Wisconsin in a week.
1. My first sample of ten lakes is in `r tbls("Frogs1",display="cite")`.

`r tbls("Frogs1")`

```{r echo=FALSE}
print(sampA,row.names=FALSE)
```

4. The number of frogs differs between the first three lakes `r tbls("Frogs1",display="cite")`. This is an example of natural variability.
1. The average number of frogs in my first sample is `r formatC(mean(sampA$visual),format="f",digits=1)`.
1. My second sample of ten lakes is in `r tbls("Frogs2",display="cite")`.

`r tbls("Frogs2")`

```{r echo=FALSE}
print(sampB,row.names=FALSE)
```

1. The average number of frogs in my first sample is `r formatC(mean(sampB$visual),format="f",digits=1)`.
1. The average number of frogs per lake differed between the two samples. This is an example of sampling variability.
1. Neither average number of frogs per lake equaled the known population average number of frogs per lake. I was not surprised by this result because a sample is representative but not perfectly representative of the entire population.

----

R Appendix
```{r eval=FALSE, prompt=FALSE}
library(NCStats)

## Lakes Exercise
<<Lakes>>

## Frog Exercise
<<Frogs>>
```

----
