---
title: "Answer Key"
subtitle: "Normal Distributions CE 1"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_NormalDist_CE1")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
```

----

## Normal Distribution Characteristics I

1. Using `snorm()`
    1. The whole distributions shifts to the right.
    1. The distribution becomes wider (more dispersed).
1. Below
    1. $\mu$=75, $\sigma$=10, hand-drawn distribution.
    1. $\mu$=5500, $\sigma$=600, hand-drawn distribution.

```{r echo=FALSE, results="hide", fig.width=7}
library(NCStats)
library(ggplot2)
library(patchwork)

mn <- 70
sd <- 10
brks <- mn+seq(-3,3,by=1)*sd

bX <- ggplot(data.frame(x=c(-Inf,Inf)),mapping=aes(x=x)) +
  stat_function(fun=dnorm,args=list(mean=mn,sd=sd),xlim=c(mn-4*sd,mn+4*sd),
                geom="line",color="black",size=1.1) +
  scale_y_continuous(expand=expansion(mult=c(0,0.04))) +
  scale_x_continuous(breaks=brks,expand=expansion(mult=c(0,0))) +
  labs(x="X") +
  theme_NCStats() +
  theme(axis.title.y=element_blank(),axis.title.x=element_text(size=16),
        axis.text.x=element_text(size=12),axis.text.y=element_blank())

mn <- 5500
sd <- 600
brks <- mn+seq(-3,3,by=1)*sd

bY <- ggplot(data.frame(x=c(-Inf,Inf)),mapping=aes(x=x)) +
  stat_function(fun=dnorm,args=list(mean=mn,sd=sd),xlim=c(mn-4*sd,mn+4*sd),
                geom="line",color="black",size=1.1) +
  scale_y_continuous(expand=expansion(mult=c(0,0.04))) +
  scale_x_continuous(breaks=brks,expand=expansion(mult=c(0,0))) +
  labs(x="Y") +
  theme_NCStats() +
  theme(axis.title.y=element_blank(),axis.title.x=element_text(size=16),
        axis.text.x=element_text(size=12,angle=45,vjust=0.9,hjust=0.9),
        axis.text.y=element_blank())

bX + bY
```

----

## Hand Calculations I
For X~N(75,10) 

1. 97.5%
1. 0.15%
1. 68%
1. 81.5%
1. Less than 55.
1. Between X=55 and 95.
1. Greater than X=95.

----

<div class="alert alert-warning">
<strong>Note:</strong>
<ul>
<li>Always stop to ask if it makes sense to record the variable on the individual. There is an urge to call the individual an "ant", but does it make sense to record the "number of ants" on an ant?</li>
<li>Another hint for identifying the individual is to look at the "forward" questions. The forward questions will always ask you to computer the percentage of individuals.</li>
</ul>
</div>

## Carpenter Ants
#### Top questions
1. A nest
1. Number of Carpenter Ants in a nest
1. Discrete quantitative
1. $\mu$=1400
1. $\sigma$=300

#### Bottom questions
1. forward, right-of.
1. reverse, right-of.
1. reverse, left-of.
1. forward, between.
1. forward, left-of.
1. reverse, right-of.
1. reverse, between.
