---
title: "Answer Key"
subtitle: "Hypothesis Testing - Errors CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_HypTesting2_CE1")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
library(captioner)
figs <- captioner(prefix="Figure")
```

----

## Body Temperature

<div class="alert alert-warning">
<strong>In the following two applied questions note:</strong> 
<ul>
<li>The hypotheses must be defined as noted in the first assignment for this module.</li>
<li>The summary statement (question 5 in each) is your answer to the research hypothesis/question based on your p-value results. Generally, this is a restatement about whether you believe (i.e., rejected H<sub>0</sub>) or don't believe (i.e., did not reject H<sub>0</sub>) the H<sub>A</sub> because the H<sub>A</sub> is generally formed from the research hypothesis. This statement should be in the context of the problem and should be made clear that it is about the parameter (see "mean" in each statement below) in the population (see "all" in each statement below). </li>
<li>Your p-value may be slightly (though not dramatically) different if you rounded your x&#772; and SE differently than I did.</li>
</ul>
</div>

<ol>
<li>H<sub>A</sub> &mu; &ne; 37 vs. H<sub>0</sub> &mu; = 37 where &mu; is mean normal body temperature for ALL people in the population.</li>
```{r echo=FALSE}
d <- read.csv("https://github.com/droglenc/NCData/raw/master/NormTemp.csv")
```

<li>The p-value is the probability that x&#772;=`r formatC(mean(d$bodytemp),format="f",digits=2)` or "not equals" assuming that &mu;=37.
<li>p-value=`r formatC(2*distrib(round(mean(d$bodytemp),2),mean=37,sd=0.45/sqrt(nrow(d))),format="f",digits=8)`</li>
<li>Reject H<sub>0</sub> because p-value < alpha;.</li>
<li>Average body temperature is less than 37<sup>o</sup>C for all individuals in this population.</li>
</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
2*distrib(36.81,mean=37,sd=0.45/sqrt(130))
```

----

## Beetle Size
```{r echo=FALSE}
d <- read.csv("https://github.com/droglenc/NCData/raw/master/Beetles.csv")
d <- filterD(d,species=="Halticus.oleracea")
```

1. H<sub>A</sub> &mu; > 190 vs. H<sub>0</sub> &mu; = 190, where &mu; is the mean thorax length (&mu;m) for all *Halticus oleracea* beetles in this population.
1. p=value is the probability that x&#772;=`r formatC(mean(d$thorax),format="f",digits=2)` &mu;m or greater assuming that &mu;=190
1. p-value=`r formatC(distrib(round(mean(d$thorax),2),mean=190,sd=15/sqrt(nrow(d)),lower.tail=FALSE),format="f",digits=3)`.
1. Do not reject H<sub>0</sub> because the p-value > &alpha;.
1. The average thorax length for all *Halticus oleracea* beetles in this population does not appear to be greater than 190 &mu;m.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(194.17,mean=190,sd=15/sqrt(18),lower.tail=FALSE)
```
