---
title: "Answer Key"
subtitle: "R Filtering Data Class Example"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_RFilter_CExmpl")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
library(ggplot2)
source("../../../rhelpers/knitr_setup.R")
```

----

# Load Data from Webpage
Nothing to show here ... this is the same as what we did in the previous class example.

```{r eval=FALSE}
firedf <- read.csv("Fire.csv")
```
```{r echo=-1}
firedf <- read.csv("http://derekogle.com/NCMTH107/modules/CE/1_CE_KEYS/Fire.csv")
str(firedf)
```

<br>

# Examining Single Variable
1. The stand type variable is shown below with `firedf$sttype`.
1. The litter depth variable is shown below with `firedf$litdep`.

#### R Code and Results
```{r}
firedf$sttype
firedf$litdep
```

<br>

# Filtering Data Frames
```{r decid, echo=FALSE, results="hide"}
decid <- filterD(firedf,sttype=="d")
str(decid)
```
```{r lt2, echo=FALSE, results="hide"}
lt2 <- filterD(firedf,litdep<2)
str(lt2)
```
```{r cmc, echo=FALSE, results="hide"}
cmc <- filterD(firedf,sttype %in% c("c","mc"))
str(cmc)
```
```{r clt2, echo=FALSE, results="hide"}
clt2 <- filterD(firedf,sttype=="c",litdep<2)
str(clt2)
```
```{r btw115, echo=FALSE, results="hide"}
btw115 <- filterD(firedf,litdep>1,litdep<1.5)
str(btw115)
```

1. There are `r kCounts(nrow(decid))` of deciduous stand types (see `decid` object below).
1. There are `r kCounts(nrow(lt2))` of stand with litter depths less than 2 cm (see `lt2` object below).
1. There are `r kCounts(nrow(cmc))` of coniferous and mixed-coniferous stand types (see `cmc` object below).
1. There are `r kCounts(nrow(clt2))` of coniferous stand types with litter depths of less than 2 cm (see `clt2` object below).
1. There are `r kCounts(nrow(btw115))` of stands with between 1 and 1.5 cm litter depths (see `btw115` object below).

#### R Code and R Results
```{r}
<<decid>>
<<lt2>>
<<cmc>>
<<clt2>>
<<btw115>>
```
