---
title: "Answer Key"
subtitle: "R Getting Data Class Example"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_RData_CExmpl")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
library(ggplot2)
source("../../../rhelpers/knitr_setup.R")
```

----

## Enter Own Data
I created a CSV file with information about members of my family, saved it in the same folder as my Notebook file, loaded the data file into the `famex` object in the Notebook (with `read.csv()`), and showed the complete file in the code below.

```{r eval=FALSE}
famex <- read.csv("Family_Example.csv")
```
```{r echo=-1}
famex <- read.csv("http://derekogle.com/NCMTH107/modules/CE/1_CE_KEYS/Family_Example.csv")
famex
```

<br>

## Load Data from Webpage
I downloaded the `fire.csv` file into the same folders as my Notebook file, loaded the data into the `firedf` object in the Notebook (with `read.csv()`), and showed the stucture of the file with `str()` (because showing the whole file was too big). From the structure one can see that there are 48 individuals (R calls them observations which it abbreviates with "obs."), 6 variables, and the variables are named `sttype`, `tslf`, `tdw`, `litdep`, `fuel1h`, and `tltd`. 

```{r eval=FALSE}
firedf <- read.csv("Fire.csv")
```
```{r echo=-1}
firedf <- read.csv("http://derekogle.com/NCMTH107/modules/CE/1_CE_KEYS/Fire.csv")
str(firedf)
```
