---
title: "Answer Key"
subtitle: "Getting Data Into R"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
---

```{r echo=FALSE, eval=FALSE, message=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_RData_CE")
```

```{r echo=FALSE, results='hide', message=FALSE, message=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
```

----

As part of the assignment template, the following two lines were run before the code run in the sections below.
```{r results="hide",message=FALSE,warning=FALSE}
library(NCStats)
library(ggplot2)
```

<div class="alert alert-warning">
<strong>Note:</strong> 
<ul>
<li>The object name to the left of each assignment operator (i.e., `<-`) can largely be whatever you want. In the examples below, I tried to use names that made sense for the data being imported.</li>
</ul>
</div>

----

# Loading Data from Webpage
```{r echo=FALSE}
bhobj <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/BrainHead.csv")
```

1. There are `r kCounts(ncol(bhobj))` variables in BrainHead.csv.
1. The variables in Brainhead.csv are: `r paste(names(bhobj),collapse=", ")`.
1. There are `r kCounts(nrow(bhobj))` individuals in BrainHead.csv.

#### R Code and Results
```{r eval=FALSE}
bhobj <- read.csv("BrainHead.csv")
```
```{r}
str(bhobj)
```

<br>

----

## Loading Your Own Data

<div class="alert alert-warning">
<strong>Note:</strong> 
<ul>
<li>When entering your own data into R, make sure that variables are organized by columns (so that individuals are organized by rows) and that each column has a useful name in the first/top row. Also make sure that your variable names and your data do not have any spaces in them as this can cause some issues with how the data are loaded into R. So, if your variable is the professor's name don't label that column as `Professors Name`; rather call it `Prof`, `Professor`, `ProfName`, `Prof.Name`, or something similar. Similarly, for the data don't write `Derek Ogle`; rather use `DerekOgle`, `Derek.Ogle`, `Derek_Ogle`, `Ogle`, `ProfessorDoctorOgleSir`, or something similar.</li>
<li>Don't use `str()`, `peek()`, or `headtail()` for data.frames that are very small. Just type the name of the data.frame to show all of the individuals in that data.frame.</li>
</ul>
</div>

Answers will vary by student, but one example is shown below.

#### R Code and Results

```{r eval=FALSE}
famex <- read.csv("Family_Example.csv")
```
```{r echo=-1}
famex <- read.csv("http://derekogle.com/NCMTH107/modules/CE/1_CE_KEYS/Family_Example.csv")
famex
```


<br>

----

## Stacked Data

<div class="alert alert-warning">
<strong>Note:</strong> 
<ul>
<li>In stacked data one variable denotes to which group an individual belongs and the other variable is the variable recorded on each individual.</li>
</ul>
</div>

The data.frame is shown below.

#### R Code and Results

```{r echo=FALSE}
test <- data.frame(
  section=c(rep("Early",7),rep("Late",5)),
  score=c(68,78,64,55,82,63,59,87,92,73,83,91)
)
```
```{r eval=FALSE}
test <- read.csv("Test_Example.csv")
```
```{r}
test
```

