---
title: "Annotated Key"
subtitle: "Univariate EDA Class Example"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_BivEDAQuant_CExmpl")
```
```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
```

## Air Quality
The relationship between ozone concentration and maximum daily temperature in New York City in the summer of 1973 is nonlinear, positive, without strong outliers (the point at approximately 80 degrees and 155 ppb ozone may be considered an outlier), and moderately strong. I did not use the correlation coefficient (r) to assess strength because of the nonlinear relationship.

&nbsp;

## Climate Change Data
The relationship between temperature anomaly and CO<sub>2</sub> in the atmosphere is linear, positive, without outliers, and strong (|r|&gt;0.80). I used the correlation coefficient to assess strength because the relationship was linear and absent of outliers.
