---
title: "Answer Key"
subtitle: "Confidence Regions - Extension CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_ConfRegions2_CE1")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
library(captioner)
figs <- captioner(prefix="Figure")
figs(name="Z1","Normal distribution representing the Z* the first confidence region example.")
figs(name="Z2","Normal distribution representing the Z* the second confidence region example.")
figs(name="Z3","Normal distribution representing the Z* the third confidence region example.")
figs(name="Z4","Normal distribution representing the Z* the fourth confidence region example.")
figs(name="Z5","Normal distribution representing the Z* the fifth confidence region example.")
figs(name="Z6","Normal distribution representing the Z* the sixth confidence region example.")
figs(name="Z7","Normal distribution representing the Z* the seventh confidence region example.")

```

----

<div class="alert alert-warning">
<strong>Note:</strong>
<ul>
<li>In the second part of each question below you MUST not say &mu; -- rather you should say what &mu; represents (e.g., "true population mean body temperature" in the first question).</li>
</ul>
</div>

## Body Temperatures
<ol>
```{r echo=FALSE, fig.show='hide'}
d <- read.csv("https://github.com/droglenc/NCData/raw/master/NormTemp.csv")
cis <- mean(d$bodytemp)+c(-1,1)*round(distrib(0.025,type="q",lower.tail=FALSE),3)*0.45/sqrt(nrow(d))
```
<li>`r formatC(mean(d$bodytemp),format="f",digits=1)`$\pm$`r formatC(distrib(0.975,type="q",lower.tail=FALSE),format="f",digits=3)`$\frac{0.45}{\sqrt{130}}$ = `r formatC(cis[1],format="f",digits=2)`-`r formatC(cis[2],format="f",digits=2)`.</li>
<li>I am 95% confident that the true population mean body temperature is between `r formatC(cis[1],format="f",digits=2)` and `r formatC(cis[2],format="f",digits=2)`<sup>o</sup>C.</li>
</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
( zstar <- distrib(0.025,type="q",lower.tail=FALSE) )
```

----

## Beetle Usage
```{r echo=FALSE, fig.show='hide'}
d <- read.csv("https://github.com/droglenc/NCData/raw/master/Beetles.csv")
d <- filterD(d,species=="Halticus.oleracea")
ci <- mean(d$thorax)+round(distrib(0.95,type="q",lower.tail=FALSE),3)*15/sqrt(nrow(d))
```

1. `r formatC(mean(d$thorax),format="f",digits=2)``r formatC(distrib(0.95,type="q",lower.tail=FALSE),format="f",digits=3)`$\frac{15}{\sqrt{18}}$ = `r formatC(ci,format="f",digits=1)`
1. I am 95% confident that the true population mean thorax length is greater than `r formatC(ci,format="f",digits=1)` mm (i.e., a lower bound).

R Appendix.
```{r eval=FALSE, prompt=FALSE}
( zstar <- distrib(0.95,type="q",lower.tail=FALSE) )
```

----

## Internet Usage
<div class="alert alert-warning">
<strong>Note:</strong>
<ul>
<li>These questions are "sample size" and not "confidence region" calculations.</li>
<li>Note that all sample size calculations should be rounded <strong>UP</strong> to a whole number. For example, in the third question below, note how 742.2 was rounded up to 743.</li>
</ul>
</div>

<ol>
```{r ncalc1, echo=FALSE, results='hide', fig.show='hide'}
sigma <- 7.55
me <- 1
(zstar <- distrib(0.025,type="q",lower.tail=FALSE))
(round(zstar,3)*sigma/me)^2
```
<li>One should take a sample size of `r formatC(ceiling((zstar*sigma/me)^2),format="f",digits=0)` household to achieve the desired level of confidence and margin-of-error.</li>

```{r ncalc2, echo=FALSE, results='hide', fig.show='hide'}
(zstar <- distrib(0.05,type="q",lower.tail=FALSE))
(round(zstar,3)*sigma/me)^2
```
<li>One should take a sample size of `r formatC(ceiling((zstar*sigma/me)^2),format="f",digits=0)` household to achieve the desired level of confidence and margin-of-error.</li>

```{r ncalc3, echo=FALSE, results='hide', fig.show='hide'}
me <- 0.5
(zstar <- distrib(0.025,type="q",lower.tail=FALSE))
(round(zstar,3)*sigma/me)^2
```

<li>One should take a sample size of `r formatC(ceiling((zstar*sigma/me)^2),format="f",digits=0)` household to achieve the desired level of confidence and margin-of-error.</li>
</ol>

R Appendix.
```{r eval=FALSE,prompt=FALSE}
<<ncalc1>>
<<ncalc2>>
<<ncalc3>>
```
