---
title: "Answer Key"
subtitle: "Normal Distributions CE 2"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_NormalDist_CE2")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
```

<div class="alert alert-warning">
<strong>Note:</strong>
<ul>
<li>If it is a reverse question (given a percentage/proportion and asked to find a value of the variable) then you must include <code>type="q"</code> in <code>distrib()</code>.</li>
<li>If it is a "right-of" question then you must include <code>lower.tail=FALSE</code> in <code>distrib()</code>.</li>
<li>All "between" questions require a three-step process with <code>distrib()</code> to answer. If it is a forward question then <code>distrib()</code> is used to find the areas to the left of the two given values and then those two areas are subtracted. If it is a reverse question then the area between must be converted to a left-of area and a right-of area which are then separately put into <code>distrib()</code> to compute two values (which are the final answer).</li>
<li>I strongly urge you to continue to identify whether the question is a "forward" or "reverse" question (which will identify whether you need to use <code>type="q"</code> in <code>distrib()</code> or not) and a "left-of", "right-of", or "between" question (which will identify whether you need to use <code>lower.tail=FALSE</code> in <code>distrib()</code> or not and whether the answer is a three-step process to calculate).</li>
</ul>
</div>

## Carpenter Ants

1. `r formatC(distrib(1900,mean=1400,sd=300,lower.tail=FALSE)*100,format="f",digits=1)`%.
1. `r formatC(distrib(0.15,mean=1400,sd=300,type="q",lower.tail=FALSE),format="f",digits=0)` ants.
1. `r formatC(distrib(0.33,mean=1400,sd=300,type="q"),format="f",digits=0)` ants.
1. `r formatC((distrib(1900,mean=1400,sd=300)-distrib(700,mean=1400,sd=300))*100,format="f",digits=1)`%.
1. `r formatC(distrib(300,mean=1400,sd=300)*100,format="f",digits=2)`%.
1. `r formatC(distrib(0.05,mean=1400,sd=300,type="q",lower.tail=FALSE),format="f",digits=0)` ants.
1. Between `r formatC(distrib(0.10,mean=1400,sd=300,type="q"),format="f",digits=0)` and  `r formatC(distrib(0.10,mean=1400,sd=300,type="q",lower.tail=FALSE),format="f",digits=0)` ants.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(1900,mean=1400,sd=300,lower.tail=FALSE)*100
distrib(0.15,mean=1400,sd=300,type="q",lower.tail=FALSE)
distrib(0.33,mean=1400,sd=300,type="q")
ab <- distrib(1900,mean=1400,sd=300)
a <- distrib(700,mean=1400,sd=300)
(ab-a)*100
distrib(300,mean=1400,sd=300)*100
distrib(0.05,mean=1400,sd=300,type="q",lower.tail=FALSE)
distrib(0.10,mean=1400,sd=300,type="q")
distrib(0.10,mean=1400,sd=300,type="q",lower.tail=FALSE)
```

----

## SAT Scores
1. SAT score is a continuous quantitative variable (this may be discrete, so I counted discrete as correct).
1. The proportion of students not accepted by School A is `r formatC(distrib(500,mean=550,sd=65,plot=FALSE),format="f",digits=3)`.
1. The percentage of students accepted by School B is `r formatC(distrib(650,mean=550,sd=65,lower.tail=FALSE,plot=FALSE)*100,format="f",digits=1)`%.
1. The percentage of students accepted by School A but not School B is `r formatC((distrib(650,mean=550,sd=65,plot=FALSE)-distrib(500,mean=550,sd=65,plot=FALSE))*100,format="f",digits=1)`%.
1. School C should set the acceptance criterion at `r formatC(distrib(0.25,type="q",mean=550,sd=65,plot=FALSE,lower.tail=FALSE),format="f",digits=0)`.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(500,mean=550,sd=65)
distrib(650,mean=550,sd=65,lower.tail=FALSE)
ab <- distrib(650,mean=550,sd=65)
a <- distrib(500,mean=550,sd=65)
ab-a
distrib(0.25,mean=550,sd=65,type="q",lower.tail=FALSE)
```

----

## Urban Deer Relocations

1. The proportion of deer with home ranges between 0.2 and 0.4 km<sup>2</sup> is `r formatC(distrib(0.4,mean=0.3,sd=0.095,plot=FALSE)-distrib(0.2,mean=0.3,sd=0.095,plot=FALSE),format="f",digits=3)`.
1. The proportion of deer with a home range greater than 0.32 km<sup>2</sup> is `r formatC(distrib(0.32,mean=0.3,sd=0.095,lower.tail=FALSE,plot=FALSE),format="f",digits=3)`.
1. The home range such that 17% of the deer have a larger home range is `r formatC(distrib(0.17,type="q",mean=0.3,sd=0.095,lower.tail=FALSE,plot=FALSE),format="f",digits=2)` km<sup>2</sup>.
1. The home range such that 32% of the deer have a smaller home range is `r formatC(distrib(0.32,type="q",mean=0.3,sd=0.095,plot=FALSE),format="f",digits=2)` km<sup>2</sup>.
1. The proportion of deer with a home range less than 0.4 km<sup>2</sup> is `r formatC(distrib(0.4,mean=0.3,sd=0.095,plot=FALSE),format="f",digits=3)`.
1. The most common 48% of home ranges are between `r formatC(distrib(0.26,type="q",mean=0.3,sd=0.095,plot=FALSE),format="f",digits=2)` and `r formatC(distrib(0.26,type="q",mean=0.3,sd=0.095,lower.tail=FALSE,plot=FALSE),format="f",digits=2)` km<sup>2</sup>.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
ab <- distrib(0.4,mean=0.3,sd=0.095)
a <- distrib(0.2,mean=0.3,sd=0.095)
ab-a
distrib(0.32,mean=0.3,sd=0.095,lower.tail=FALSE)
distrib(0.17,mean=0.3,sd=0.095,type="q",lower.tail=FALSE)
distrib(0.32,mean=0.3,sd=0.095,type="q")
distrib(0.4,mean=0.3,sd=0.095)
distrib(0.26,mean=0.3,sd=0.095,type="q")
distrib(0.26,mean=0.3,sd=0.095,type="q",lower.tail=FALSE)
```

<br>
<br>
<br>
<br>
<br>
<br>

----

## iPhone Battery Lifespan 

1. Number of charge-cycles is a discrete quantitative variable.
1. The proportion of betteries that would be rated as "exceptional" is `r formatC(distrib(450,mean=400,sd=20,lower.tail=FALSE,plot=FALSE),format="f",digits=3)`.
1. The percentage of batteries that would be rated as better than "acceptably poor" but not "exceptional" is `r formatC((distrib(450,mean=400,sd=20,plot=FALSE)-distrib(375,mean=400,sd=20,plot=FALSE))*100,format="f",digits=1)`%.
1. The new definition of "exceptional" (i.e., the top 10% of batteries) would be `r formatC(distrib(0.10,mean=400,sd=20,type="q",lower.tail=FALSE,plot=FALSE),format="f",digits=0)` charge-cycles.
1. The new definition of "unacceptably poor" (i.e., the bottom 25% of batteries) would be `r formatC(distrib(0.25,mean=400,sd=20,type="q",plot=FALSE),format="f",digits=0)` charge-cycles.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(450,mean=400,sd=20)
ab <- distrib(450,mean=400,sd=20)
a <- distrib(375,mean=400,sd=20)
ab-a
distrib(0.10,mean=400,sd=20,type="q",lower.tail=FALSE)
distrib(0.25,mean=400,sd=20,type="q")
```

----

## Normal Distribution Characteristics II

1. Below
    1. $\mu$=0, $\sigma$=1
    1. $\mu$=90, $\sigma$=8
    
----

## Hand Calculations II
For Z~N(0,1)

1. Z=1
1. 2.5%
1. 99.85%
1. Between Z=-3 and 3.

----

## Pollen Counts
#### Top questions
1. A day in LaCrosse, WI n early September
1. Pollen count (spores per cubic meter)
1. Discrete quantitative
1. $\mu$=40
1. $\sigma$=8

#### Bottom questions
1. (a) reverse, right-of; (b) `r formatC(distrib(0.1,mean=40,sd=8,type="q",lower.tail=FALSE),format="f",digits=1)` spores per cubic meter.
1. (a) forward, left-of; (b) `r formatC(distrib(50,mean=40,sd=8),format="f",digits=3)`.
1. (a) forward, between; (b) `r formatC(distrib(55,mean=40,sd=8)-distrib(32,mean=40,sd=8),format="f",digits=3)`.
1. (a) reverse, left-of; (b) `r formatC(distrib(0.3,mean=40,sd=8,type="q"),format="f",digits=1)` spores per cubic meter.
1. (a) reverse, between; (b) Between `r formatC(distrib(0.15,mean=40,sd=8,type="q"),format="f",digits=1)` and  `r formatC(distrib(0.15,mean=40,sd=8,type="q",lower.tail=FALSE),format="f",digits=1)` spores per cubic meter.
1. (a) forward, right-of; (b) `r formatC(distrib(35,mean=40,sd=8,lower.tail=FALSE),format="f",digits=3)`.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(0.1,mean=40,sd=8,type="q",lower.tail=FALSE)
distrib(50,mean=40,sd=8)
ab <- distrib(55,mean=40,sd=8)
a <- distrib(32,mean=40,sd=8)
ab-a
distrib(0.3,mean=40,sd=8,type="q")
distrib(0.15,mean=40,sd=8,type="q")
distrib(0.15,mean=40,sd=8,type="q",lower.tail=FALSE)
distrib(35,mean=40,sd=8,lower.tail=FALSE)
```

----


<div class="alert alert-warning">
<strong>Note:</strong>
<ul>
<li>The IQR question below (#4) is the same as asking "what is the most common 50% of times to pass between the two points?".</li>
<li>For question #6 below, the median is equal to the mean for a normal distribution (or any distribution that is perfectly symmetric).</li>
</ul>
</div>

## Driving Speed
#### Top questions
1. A car traveling between the two points
1. Time to travel between the two points
1. Continuous quantitative
1. $\mu$=2.5
1. $\sigma$=0.75

#### Bottom questions
1. (a) reverse, right-of; (b) `r formatC(distrib(0.15,mean=2.5,sd=0.75,type="q",lower.tail=FALSE),format="f",digits=2)` s
1. (a) forward, left-of; (b) `r formatC(distrib(1,mean=2.5,sd=0.75),format="f",digits=3)`.
1. (a) forward, between; (b) `r formatC(distrib(4.5,mean=2.5,sd=0.75)-distrib(1.5,mean=2.5,sd=0.75),format="f",digits=3)`.
1. (a) reverse, between; (b) The IQR is from a Q1 of `r formatC(distrib(0.25,mean=2.5,sd=0.75,type="q"),format="f",digits=2)` s to a Q3 of `r formatC(distrib(0.25,mean=2.5,sd=0.75,type="q",lower.tail=FALSE),format="f",digits=2)` s.
1. (a) forward, right-of; (b) `r formatC(distrib(7,mean=2.5,sd=0.75,lower.tail=FALSE),format="f",digits=12)`.
1. (a) reverse, left-of or right-of; (b) 2.5 s.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(0.15,mean=2.5,sd=0.75,type="q",lower.tail=FALSE)
distrib(1,mean=2.5,sd=0.75)
ab <- distrib(4.5,mean=2.5,sd=0.75)
a <- distrib(1.5,mean=2.5,sd=0.75)
ab-a
distrib(0.25,mean=2.5,sd=0.75,type="q")
distrib(0.25,mean=2.5,sd=0.75,type="q",lower.tail=FALSE)
distrib(7,mean=2.5,sd=0.75,lower.tail=FALSE)
```

----

## Turkey Spur Length

1. The spur length such that 30% of the turkeys have a longer spur length is `r formatC(distrib(0.3,mean=20.9,sd=3.7,type="q",lower.tail=FALSE,plot=FALSE),format="f",digits=0)` mm.
1. The proportion of turkeys with spur lengths between 15 and 25 mm is `r formatC(distrib(25,mean=20.9,sd=3.7,plot=FALSE)-distrib(15,mean=20.9,sd=3.7,plot=FALSE),format="f",digits=3)`.
1. The proportion of turkeys with spur lengths greater than 30 mm is `r formatC(distrib(30,mean=20.9,sd=3.7,lower.tail=FALSE,plot=FALSE),format="f",digits=3)`.
1. The spur length such that 10% of turkeys have a smaller spur length is `r formatC(distrib(0.1,mean=20.9,sd=3.7,type="q",plot=FALSE),format="f",digits=0)` mm.
1. The proportion of urkeys with a spur length less than 18 mm is `r formatC(distrib(18,mean=20.9,sd=3.7,plot=FALSE),format="f",digits=3)`.
1. The most common 80% of spur lengths are between `r formatC(distrib(0.1,mean=20.9,sd=3.7,type="q",plot=FALSE),format="f",digits=0)` and `r formatC(distrib(0.1,mean=20.9,sd=3.7,type="q",lower.tail=FALSE,plot=FALSE),format="f",digits=0)` mm.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
distrib(0.3,mean=20.9,sd=3.7,type="q",lower.tail=FALSE)
ab <- distrib(25,mean=20.9,sd=3.7)
a <- distrib(15,mean=20.9,sd=3.7)
ab-a
distrib(30,mean=20.9,sd=3.7,lower.tail=FALSE)
distrib(0.1,mean=20.9,sd=3.7,type="q")
distrib(18,mean=20.9,sd=3.7)
distrib(0.1,mean=20.9,sd=3.7,type="q")
distrib(0.1,mean=20.9,sd=3.7,type="q",lower.tail=FALSE)
```

----
