---
title: "Answer Key"
subtitle: "Hypothesis Testing Errors CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_HypTestingErrs_CE")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
library(captioner)
figs <- captioner(prefix="Figure")
figs(name="Pwr1","NULL normal distribution representing the rejection region for the power calculation example.")
figs(name="Pwr2","ACTUAL normal distribution representing power in the power calculation example.")
```

----

## Identifying Type I and II Errors

<div class="alert alert-warning">
<strong>Note:</strong> 
<ul>
<li>In these two questions, I would first write the H<sub>0</sub> and H<sub>A</sub> as in the last module. I would then write these hypothesis in words relative to the research hypothesis. For example, in the first question, H<sub>0</sub> is that the chicken is not contaminated whereas H<sub>A</sub> is that the chicken is contaminated. I then write the definitions of Type I (H<sub>0</sub> is true, but it is rejected) and Type II (H<sub>0</sub> is false, but is not rejected) errors, but with the wordy meanings of the hypotheses substituted in appropriately.</li>
</ul>
</div>

<ol>
<li>The H<sub>0</sub>: &mu;=15.4 and H<sub>A</sub>: &mu;>15.4. A Type I error is saying that the chicken is contaminated (rejecting H<sub>0</sub>) when it really is not contaminated (H<sub>0</sub> is true). A Type II error is saying that the chicken is not contaminated (not rejecting H<sub>0</sub>) when it really is contaminated (H<sub>0</sub> is false). In my mind, a Type II error is more egregious as people would be eating contaminated chicken that they have been told is not contaminated.</li>
<li>The H<sub>0</sub> is that the drug does not exceed the specification whereas the H<sub>A</sub> is that the drug does exceed the specification. A Type I error occurs if the drug is not bad (H<sub>0</sub> is true) but it is concluded that is bad (rejecting H<sub>0</sub>). A Type II error occurs if the drug is bad (H<sub>0</sub> is false) but it is concluded that the drug is not bad (not rejecting H<sub>0</sub>). In my mind, a Type II error is more egregious as people would be using a "bad" drug that they were told was not bad.</li>
</ol>

----

## How to Reduce &beta;

<div class="alert alert-warning">
<strong>Note:</strong> 
<ul>
<li>In question 1 below you are asked to prove that the rejection wall is at 97 (within rounding). This requires finding the value of the sample mean that has &alpha; "more extreme" (as dictated by the H<sub>A</sub>) on the "null distribution" (sampling distribution assuming that H<sub>O</sub> is true). This is always a reverse calculation and, given this H<sub>A</sub>, is a reverse-left calculation.</li>
<li>In question 2 below you are asked to prove that &beta; is 0.137 (within rounding). This requires finding the area in the "Do Not Reject H<sub>0</sub>" region of "actual distribution" (sampling distribution assuming that &mu; equals the actual mean provided). This is always a forward calculation and always in the opposite direction of the reverse calculation in quation 1; thus, a forward-right calculation in this instance.</li>
</ul>
</div>

<ol>
<li>See `r figs("Pwr1",display="cite")`.</li>
```{r Pwr1,echo=FALSE}
distrib(0.05,mean=100,sd=10/sqrt(30),type="q")
```

`r figs("Pwr1")`

<li>See `r figs("Pwr2",display="cite")`.</li>
```{r Pwr2,echo=FALSE}
distrib(96.997,mean=95,sd=10/sqrt(30),lower.tail=FALSE)
```

`r figs("Pwr2")`

<li>Effects on &beta;</li>
<ol>
<li>Increasing &alpha; will DEcrease &beta;</li>
<li>Increasing n will DEcrease &beta;</li>
<li>Increasing &sigma; will INcrease &beta;</li>
<li>Increasing how wrong the H<sub>0</sub> is will DEcrease &beta;</li>
</ol>
<li>&beta; my be DEcreased by INcreasing n or INcreasing &alpha;.</li>
<li>Power may be INcreased by INcreasing n or INcreasing &alpha;.</li>
</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
<<Pwr1>>
<<Pwr2>>
```

----
