---
title: "Answer Key"
subtitle: "Linear Regression CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("modules/CE/1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_LinearRegression_CE")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
```

----

## Equation of the Line I

1. **Dishwasher Example**
    1. response - **Height of suds (mm)**
    1. explanatory - **Amount of soap (g)**
    1. slope is **12.4**
    1. For each additional gram of soap, the height of suds increases by 12.4 mm, on average.
    1. For 0 grams of soap, the height of suds is -20.2 mm, on average.  [*This is an extrapolation.*]
    1. Asked to predict height of suds for an amount of soap between 3.5 and 8.0 grams.  Example ... "What is the predicted height of suds for 5 g of soap?"  Answer ... -20.2+12.4&times;5=41.8 mm.
    1. Asked to predict height of suds for an amount of soap not between 3.5 and 8.0 grams.  Example ... "What is the predicted height of suds for 10 g of soap?"

<!----
1. **Livestock Food Example**
    1. response - Costs
    1. explanatory - Livestock food mix in tons
    1. Slope is 190.
    1. Intercept is 12075
    1. For each additional ton of the food mix, the cost increases by $190, on average.
    1. For 0 tons of the food mix, the costs are $12075, on average. [*This may be considered as "start-up" costs.*]
    1. Ask to predict costs for food mix amounts between 36.5 and 42 tons.  Example ... "What is the cost to make 40 tons of the mix?"  Answer ... 190*40+12075 = $19675.
    1. Ask to predict costs for food mix amounts not between 36.5 and 42 tons.  Example ... "What is the cost to make 70 tons of the mix?"
---->

----

<div class="alert alert-warning">
<strong>Note:</strong> In the applied questions below ...
<ul>
<li>The words "response", "explanatory", "y", and "x" are not used at all. These words should be replaced with the actual variable names used in the problem.</li>
<li>The x-axis (explanatory) variable is <strong>increased</strong> by 1 unit when describing the slope. If the slope is positive then the y-axis (response) variable increases by the slope amount for each 1 unit increase of the explanatory variable. If the slope is negative then the y-axis (response) variable decreases by the slope amount (with the negative sign removed) for each 1 unit increase of the explanatory variable.</li>
<li>Interpret the y-intercept value even if it does not make sense, which will happen often as the y-intercept is often an extreme extrapolation.</li>
<li>When computing the residual, make sure to do the observed y-axis (response) value minus the predicted y-axis (response) value. About half the time this will result in a negative value. That is okay, it just means that the observed value was less than the predicted value (i.e., the observed point was below the best-fit line).</li>
<li>The words "proportion of variability explained" <strong>always</strong> imply using r<sup>2</sup> as the answer.</li>
<li>The correlation coefficient is always the value of r. If you compute r by taking the square root of r<sup>2</sup> then remember to put a negative sign on it if the slope (i.e., direction or association) is negative (your calculator will always return a positive value).</li>
<li>When assessing whether anything "concerns" you about the regression then you should address whether the form looks linear and if there is homoscedasticty. Non-linear forms will look obviously curved, heteroscedastic forms will look like a funnel (usually from narrow on the left to wider on the right). If you do not see a curve or a funnel, then explicitly say that you do not have any concerns about the lack of linearity or homoscedasticity.</li>
</ul>
</div>


## Beach Sand
```{r echo=FALSE}
m <- 0.053
b <- 0.16
rsq <- 0.91
```

1. The equation of the best-fit line is **SAND = `r b`+`r m`&times;ANGLE**
1. For each 1 degree increase in beach angle, the median sand diameter increases by `r m` mm, on average.
1. If there is 0 beach angle, then the median sand diameter is `r b` mm, on average.
1. This prediction should not be computed, because 15 degrees of beach angle is outside of the domain of the data used to define the relationship (i.e., this is an **extrapolation**).
1. The predicted median sand diameter for a beach angle of 4 degrees is **`r formatC(b+m*4,format="f",digits=3)` mm**.
1. The residual for an observed median sand diameter of 0.2 mm and beach angle of 5 degrees is **`r formatC(0.2-(b+m*5),format="f",digits=3)` mm**.
1. The correlation coefficient between median sand diameter and beach angle is **`r formatC(-1*sqrt(rsq),format="f",digits=3)`**.
1. The proportion of variability in median sand diameter that is explained by knowing the angle of the beach is **`r formatC(rsq,format="f",digits=3)`**.
1. I would expect the median sand diameter to increase by four slopes or **`r formatC(4*m,format="f",digits=3)` mm**, on average, if the beach angle increased by 4 degrees.
1. The relationship between median sand diameter and beach angle appears to be nonlinear (i.e., curved). I am not concerned about homoscedastiticty as no funnel-shape is apparent.

----


## Everest Temperatures
```{r echo=FALSE}
m <- -0.0059
b <- 29.8
rsq <- 0.923
```

1. The equation of the best-fit line is **AIRTEMP = `r b``r m`&times;ALTLAPSERATE**
1. If the altitude lapse rate is 0, then the mean air temperature is `r b`<sup>o</sup>C, on average.
1. For each 1 degree increase in altitude lapse rate, the mean air temperature decreases by `r -1*m`<sup>o</sup>C, on average.
1. The predicted mean air temperature for an altitude lapse rate of 4000 <sup>o</sup>C/km is **`r formatC(b+m*4000,format="f",digits=2)`<sup>o</sup>C**.
1. The residual for an observed mean air temperature of 0<sup>o</sup>C and altitude lapse rate of 3500<sup>o</sup>C/km is **`r formatC(0-(b+m*3500),format="f",digits=2)`<sup>o</sup>C**.
1. The correlation coefficient between mean air temperature and altitude lapse rate is **`r formatC(sqrt(rsq),format="f",digits=3)`**.
1. I would expect the mean air temperature to decrease by 1000 slopes or **`r formatC(-1000*m,format="f",digits=1)`<sup>o</sup>C**, on average, if the altitude lapse rate increased by 1000<sup>o</sup>C/km.
1. This prediction should not be computed, because 1000<sup>o</sup>C/km of altitude lapse rate is outside of the domain of the data used to define the relationship (i.e., this is an **extrapolation**).
1. The proportion of variability in mean air temperature that is explained by knowing the altitude lapse rate is **`r formatC(rsq,format="f",digits=3)`**.
1. I have **no concerns** as the relationship between mean air temperature and altitude lapse rate appears to be linear (no evident curve) and homoscedastic (no evident funnel).
