---
title: "Answer Key"
subtitle: "R Chi-Square Class Example"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_RChi_CExmpl")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
```

----

## Recycling in Two Communities II
The full 11-steps for this example are in these [annotations for the chi-square test class example](KEY_ChiSquare_CExmpl). However, the calculations required for Steps 5, 6, 7, 8, and 11 may be computed in R if the raw data are available as they were here.

The data are loaded below. Make note of how the data are stacked as in the two-sample t-test example. Also note that the variable of interest here is `method2` (`method1` provided a more detailed method).
```{r}
dfobj <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/PlasticRecycling.csv")
str(dfobj)
headtail(dfobj)
```

The primary R function for this analysis is `chisq.test()`. However, it requires an observed frequency table as its first argument, which must be calculated first with `xtabs()` as you did in the Bivariate EDA - Categorical module (remember that the row variable comes first). The `chisq.test()` also requires `correct=FALSE` so that an unneeded correction factor is not applied.

```{r}
( obs <- xtabs(~city+method2,data=dfobj) )
( chi1 <- chisq.test(obs,correct=FALSE) )
```

From this output one can see

* The observed table needed for Step 6.
* The chi-square test statistic (=`r formatC(chi1$statistic,format="f",digits=1)`) and df (=`r chi1$parameter`) for Step 7.
* The p-value (=`r formatC(chi1$p.value,format="f",digits=11)`) for Step 8.

The expected table for Step 5 can be shown by appending `$expected` to the object saved from `chisq.test()`.

```{r}
chi1$expected
```

And, while it is not needed, you can see the parts that went into the chi-square test statistic calculation with
```{r}
chi1$residuals^2
```

Note that you will complete all 11 steps when you have the raw data, but you should use `chisq.test()` as shown here to perform all of the calculations for you.
