---
title: "Answer Key"
subtitle: "Hypothesis Testing CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_HypTesting_CE")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
library(captioner)
figs <- captioner(prefix="Figure")
figs(name="P1","Normal distribution representing the p-value calculation for the first example.")
figs(name="P2","Normal distribution representing the p-value calculation for the second example.")
figs(name="P3","Normal distribution representing the p-value calculation for the third example.")

figs(name="P4","Normal distribution representing the p-value calculation for the fourth example.")
figs(name="P5","Normal distribution representing the p-value calculation for the fifth example.")
```

----

## Identifying Hypotheses

<ol>
<li>$H_{A}: \mu < 90$ vs. $H_{0}: \mu = 90$, where $\mu$ is the mean number of murders for all burroughs in 2010.</li>
<li>$H_{A}: \mu > 15$ vs. $H_{0}: \mu = 15$, where $\mu$ is the mean length of all oil changes at the Toyota dealership.</li>
<li>$H_{A}: \mu \neq 22500$ vs. $H_{0}: \mu = 22500$, where $\mu$ is the mean cost of all bariatric surgeries.</li>
<li>$H_{A}: p > 0.61$ vs. $H_{0}: p = 0.61$, where $p$ is the proportion of all Washburn High School students taht are involved in at least one extracurricular activity.</li>
</ol>

----

## P-Value Calculations and Decisions I

<ol>
<li>The p-value is "the probability of observing a sample mean of 11.2 or 'different' if the population mean is 14." Reject $H_{0}$ because p-value=`r formatC(2*distrib(11.2,mean=14,sd=6/sqrt(25)),format="f",digits=4)`<&alpha; (`r figs("P1",display="cite")`).</li>
```{r p1, echo=FALSE, results='hide'}
2*distrib(11.2,mean=14,sd=6/sqrt(25))
```

`r figs("P1")`

<li>The p-value is "the probability of observing a sample mean of 79.5 or greater if the population mean is 75." Reject $H_{0}$ because p-value=`r formatC(distrib(79.5,mean=75,sd=12/sqrt(50),lower.tail=FALSE),format="f",digits=4)`<&alpha; (`r figs("P2",display="cite")`).</li>
```{r p2, echo=FALSE}
distrib(79.5,mean=75,sd=12/sqrt(50),lower.tail=FALSE)
```

`r figs("P2")`

<li>The p-value is "the probability of observing a sample mean of 13700 or less if the population mean is 15000." Do not reject $H_{0}$ because p-value=`r formatC(distrib(13700,mean=15000,sd=8000/sqrt(50)),format="f",digits=4)`>&alpha; (`r figs("P3",display="cite")`).</li>

```{r p3, echo=FALSE}
distrib(13700,mean=15000,sd=8000/sqrt(50))
```

`r figs("P3")`

</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
<<p1>>
<<p2>>
<<p3>>
```

----

## Body Temperature
<ol>
<li>$H_{A}: \mu \neq 37$ vs. $H_{0}: \mu = 37$</li>
```{r echo=FALSE}
d <- read.csv("https://github.com/droglenc/NCData/raw/master/NormTemp.csv")
```

<li>x&#772;=`r formatC(mean(d$bodytemp),format="f",digits=3)`</li>
<li>p-value=`r formatC(2*distrib(round(mean(d$bodytemp),3),mean=37,sd=0.41/sqrt(nrow(d))),format="f",digits=9)`</li>
<li>Reject $H_{0}$ because p-value < alpha;.</li>
<li>Average body temperature is somewhat less (but significantly) less than 37C.</li>
</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
d <- read.csv("NormTemp.csv")
Summarize(~bodytemp,data=d,digits=3)
2*distrib(36.808,mean=37,sd=0.41/sqrt(130))
```

----

## Beetle Size
```{r echo=FALSE}
d <- read.csv("https://github.com/droglenc/NCData/raw/master/Beetles.csv")
d <- filterD(d,species=="Halticus.oleracea")
```

1. $H_{A}: \mu > 190$ vs. $H_{0}: \mu = 190$, where $\mu$ is the mean thorax length ($\mu$m) for all *Halticus oleracea* beetles in this population.
1. $\bar{x}$=`r formatC(mean(d$thorax),format="f",digits=3)` $\mu$m.
`. p-value=`r formatC(distrib(round(mean(d$thorax),3),mean=190,sd=14/sqrt(nrow(d)),lower.tail=FALSE),format="f",digits=3)`.
1. Do not reject $H_{0}$ because the p-value $> \alpha$.
1. The average thorax length for all *Halticus oleracea* beetles in this population does not appear to be greater than 190 $\mu$m.

R Appendix.
```{r eval=FALSE, prompt=FALSE}
setwd("c:/stats")
d <- read.csv("Beetles.csv")
d <- filterD(d,species=="Halticus.oleracea")
Summarize(~thorax,data=d,digits=3)
distrib(194.167,mean=190,sd=14/sqrt(18),lower.tail=FALSE)
```

----

## P-Value Calculations and Decisions II

<ol>
<li>The p-value is "the probability of observing a sample mean of 918 or greater if the population mean is 880." Reject $H_{0}$ because p-value=`r formatC(distrib(918,mean=880,sd=90/sqrt(80),lower.tail=FALSE),format="f",digits=4)`<&alpha; (`r figs("P4",display="cite")`).</li>

```{r p4, echo=FALSE}
distrib(918,mean=880,sd=90/sqrt(80),lower.tail=FALSE)
```

`r figs("P4")`

<li>The p-value is "the probability of observing a sample mean of 824 or 'different' if the population mean is 800." Do not reject $H_{0}$ because p-value=`r formatC(2*distrib(824,mean=800,sd=60/sqrt(25),lower.tail=FALSE),format="f",digits=4)`>&alpha; (`r figs("P5",display="cite")`).</li>

```{r p5, echo=FALSE, results='hide'}
2*distrib(824,mean=800,sd=60/sqrt(25),lower.tail=FALSE)
```

`r figs("P5")`

<li>The p-value is "the probabiity of observing a sample mean of 73 or less if the population mean is 80." The null hypothesis is rejected because the p-value (=`r formatC(distrib(73,mean=80,sd=20/sqrt(30)),format="f",digits=4)`)$<\alpha$.</li>
<li>The p-value is "the probabiity of observing a sample mean of 1370 or 'different' if the population mean is 1500." The null hypothesis is not rejected because the p-value (=`r formatC(2*distrib(1370,mean=1500,sd=800/sqrt(40)),format="f",digits=4)`)$>\alpha$.
</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
<<p4>>
<<p5>>
distrib(73,mean=80,sd=20/sqrt(30))
2*distrib(1370,mean=1500,sd=800/sqrt(40))
```
----
