---
title: "Answer Key"
subtitle: "1-Sample Z CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/ce-keys/")
source("../../rhelpers/rhelpers.R")
modHTML("1_Sample_Z")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../rhelpers/knitr_setup.R")
library(captioner)
tbls <- captioner(prefix="Table")
tbls(name="HawksRidge","Results from the one-sample z-test for testing that the mean number of hawks observed at Hawk Ridge, MN is greater than 50000 per year.")
```

----

## Cancer Tumors
```{r TumorCalc, echo=FALSE, results='hide', fig.show='hide'}
xbar <- 31.7; sigma <- 15; n <- 42
se <- sigma/sqrt(n)
t <- (xbar-27)/se
pv <- distrib(t,lower.tail=FALSE)
lcb <- xbar+distrib(0.9,lower.tail=FALSE,type="q")*se
```

<ol>
<li>&alpha;=0.10.</li>
<li>H<sub>A</sub>: &mu;&ne;27 and H<sub>0</sub>: &mu;=27, where &mu; is the mean time of recurrence for ALL bladder cancer patients that received the placebo.</li>
<li>1-sample Z-test because (i) one population (bladder cancer patients that received the placebo) was sampled, (ii) a quantitative variable (time of recurrence) was recorded, and (iii) &sigma; is assumed to be known (=15).</li>
<li>Loosely an "experiment" as the subjects were given a placebo treatment.  There is no obvious randomization apparent in this study.</li>
<li>The assumptions are met because (i) &sigma; is known (=15) and (ii) n=42>30.</li>
<li>The sample mean ($\bar{x}$) is `r xbar` months.</li>
<li>t=`r formatC(t,format="f",digits=4)`.</li>
<li>p-value=`r kPvalue(pv,include.p=FALSE,latex=FALSE)`.</li>
<li>Reject H<sub>0</sub> because the p-value<&alpha;.</li>
<li>It appears that the time of recurrence for the tumor for ALL patients that received the placebo is greater than the hypothesized 27 months.</li>
<li>I am 90% confident that the time of recurrence for the tumor for ALL patients that received the placebo is greater than `r formatC(lcb,format="f",digits=2)` months.</li>
</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
<<TumorCalc>>
```


----

## Raptors at Hawk's Ridge

```{r HawksRidgeCode, echo=FALSE, results='hide'}
hr <- read.csv("https://github.com/droglenc/NCData/raw/master/HawksRidge.csv")
( hr.z <- z.test(hr$Total,mu=50000,sd=37000,alt="greater",conf.level=0.99) )
```

<ol>
<li>$\alpha=0.01$</li>
<li>$H_{O}:\mu=50000$, $H_{A}:\mu>50000$, where $\mu$ is the mean number of raptors seen in ALL years at Hawk's Ridge.</li>
<li>A one-sample Z-test is required because (i) a quantitative variable (number of raptors) was measured, (ii)the individuals are from one population (all years at Hawk's Ridge), and (iii) $\sigma$ is known (=37000).  Thus, the population mean is being compared to a specific value in the null hypothesis.</li>
<li>An observational sample of years that was not random was taken.</li>
<li>The assumptions are met because $n=32>30$ and $\sigma$ is known.</li>
<li>$\bar{x}=$ `r formatC(hr.z$estimate,format="f",digits=0)` (`r tbls("HawksRidge",display="cite")`).</li>

`r tbls("HawksRidge")`

```{r echo=FALSE}
kHtest(hr.z)
```
<li>The test statistic is Z=`r formatC(hr.z$statistic,format="f",digits=3)` (`r tbls("HawksRidge",display="cite")`).</li>
<li>The p-value is `r kPvalue(hr.z$p.value,include.p=FALSE)` (`r tbls("HawksRidge",display="cite")`).</li>
<li>The $H_{O}$ is rejected because the $p-value<\alpha$ (`r tbls("HawksRidge",display="cite")`).</li>
<li>The mean number of raptors seen per year in ALL years appears to be greater than 50000.</li>
<li>I am 99\% confident that the mean number of raptors seen per year in ALL years is more than `r formatC(hr.z$conf.int[1],format="f",digits=0)` (`r tbls("HawksRidge",display="cite")`).</li>
</ol>

R Appendix
```{r eval=FALSE, prompt=FALSE}
<<HawksRidgeCode>>
```

----
