---
title: "Answer Key"
subtitle: "2-Sample t CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/ce-keys/")
source("../../rhelpers/rhelpers.R")
modHTML("2_Sample_t")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../rhelpers/knitr_setup.R")
library(captioner)
tbls <- captioner(prefix="Table")
tbls(name="Ozone","Results from the one-sample t-test for testing that the mean thickness of the ozone layer is different than 300 Dobson units.")
```

----

## Compost Type and Germination Rates
```{r GerminationCalc, echo=FALSE, results='hide', fig.show='hide'}
xbar1 <- 42.67; sigma1 <- 4.33; n1 <- 18
xbar2 <- 28.90; sigma2 <- 3.21; n2 <- 10
df <- n1+n2-2
sp2 <- ((n1-1)*(sigma1^2)+(n2-1)*(sigma2^2))/df
se <- sqrt(sp2*(1/n1+1/n2))
stat <- xbar1-xbar2
t <- stat/se
pv <- distrib(t,distrib="t",df=df,lower.tail=FALSE)
lcb <- stat+distrib(0.9,type="q",distrib="t",df=df,lower.tail=FALSE)*se
```

<ol>
<li>$\alpha$=0.10.</li>
<li>$H_{A}: \mu_{b}-\mu_{h}>0$ and $H_{0}: \mu_{b}-\mu_{h}=0$, where $\mu$ is the mean number of germinated plants, $b$ is the bought compost, and $h$ is the homemade compost.</li>
<li>2-sample t-test because (i) two populations/groups are considered (bought and homemade compost), (ii) a quantitative variable (number germinated) was recorded, and (iii) the samples are independent (there is no known connection between all pairs of bought and homemade individuals.</li>
<li>An experimental study with some randomization (placement of containers in the greenhouse).</li>
<li>The assumptions are met because (i) the samples are independent (as described above) and (ii) $n_{b}+n_{h}$=18+10=28>15 and the two sample distributions (shown on the handout) are not obviously strongly skewed.</li>
<li>$\bar{x}_{1}-\bar{x}_{2}$=`r xbar1`-`r xbar2`=`r stat`.</li>
<li>t=`r formatC(t,format="f",digits=4)` with `r df` df.</li>
<li>p-value`r kPvalue(pv,include.p=FALSE)`.</li>
<li>Reject $H_{0}$ because the p-value<$\alpha$.</li>
<li>It appears that the the number of plants that germinate is significantly lower for the homemage than the bought compost.</li>
<li>I am 90% confident that the number of plants that germinated with the bought compost is more than `r formatC(lcb,format="f",digits=1)` plants greater than the number of plants that germinated with the homemade compost.</li>
</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
<<GerminationCalc>>
```


----

## Thickness of the Ozone Layer

```{r DeliveryCode, echo=FALSE, results='hide'}
del <- data.frame(war=rep(c("Gulf","Bosnia"),each=9),
                  time=c(28.0,20.0,26.5,10.6,9.1,35.2,29.1,41.2,27.5,
                         15.1,6.4,5.0,11.4,6.5,6.5,3.0,7.0,3.5))
(del.lev <- levenesTest(time~war,data=del) )
( del.t <- t.test(time~war,data=del,var.equal=TRUE,alt="less") )
```

<ol>
<li>$\alpha=0.01$</li>
<li>$H_{A}: \mu_{b}-\mu_{g}<0$ and $H_{0}: \mu_{b}-\mu_{g}=0$, where $\mu$ is the mean time to deliver ALL shipmenets, $b$ is for Bosnia, and $g$ is for Persian Gulf War.</li>
<li>2-sample t-test because (i) two populations/groups are considered (Bosnia and Persian Gulf), (ii) a quantitative variable (delivery time) was recorded, and (iii) the samples are independent (there is no known connection between all deliveries to the two wars).</li>
<li>An observational study of 45 random samples of ozone was conducted.</li>
<li>The assumptions are met because $n=45>40$ and $\sigma$ is UNknown.</li>
<li>$\bar{x}=$ `r formatC(ozo.t$estimate,format="f",digits=0)` (`r tbls("Ozone",display="cite")`).</li>

`r tbls("Ozone")`

```{r echo=FALSE}
kHtest(ozo.t)
```
<li>The test statistic is Z=`r formatC(ozo.t$statistic,format="f",digits=3)` with `r formatC(ozo.t$parameter,format="f",digits=0)` df (`r tbls("Ozone",display="cite")`).</li>
<li>The p-value `r kPvalue(ozo.t$p.value,include.p=FALSE)` (`r tbls("Ozone",display="cite")`).</li>
<li>The $H_{O}$ is rejected because the $p-value<\alpha$ (`r tbls("Ozone",display="cite")`).</li>
<li>The mean thickness of the ozone layer at ALL areas abova Arosa, SWI appears to be different (specifically greater) than 300 Dobson units.</li>
<li>I am 99\% confident that the mean thickness of the ozone layer in ALL areas above Arosa, SWI is between `r formatC(ozo.t$conf.int[1],format="f",digits=0)` and  `r formatC(ozo.t$conf.int[2],format="f",digits=0)` Dobson Units (`r tbls("Ozone",display="cite")`).</li>
</ol>

R Appendix
```{r eval=FALSE, prompt=FALSE}
<<OzoneCode>>
```

----
