---
title: "Answer Key"
subtitle: "Confidence Intervals CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/ce-keys/")
source("../../rhelpers/rhelpers.R")
modHTML("Confidence_Intervals")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../rhelpers/knitr_setup.R")
```

----

## Class Excercise 1

Not done.

----

## Class Excercise 2

<ol>
<li>See below.</li>
<ol type="a">
<li>C=0.90</li>
<li>Lower bound</li>
<li>Z*=`r formatC(distrib(0.9,type="q",lower.tail=FALSE),format="f",digits=3)`.</li>
```{r}
zstar <- distrib(0.9,type="q",lower.tail=FALSE)
```
<li>`r formatC(79.5+zstar*12/sqrt(50),format="f",digits=1)`.</li>
<li>I am 90% confident that the true population mean is greater than `r formatC(79.5+zstar*12/sqrt(50),format="f",digits=1)`.</li>
</ol>

<li>See below.</li>
<ol type="a">
<li>C=0.95</li>
<li>Interval</li>
<li>Z*= $\pm$ `r formatC(distrib(0.025,type="q",lower.tail=FALSE),format="f",digits=3)`.</li>
```{r}
zstar <- distrib(0.025,type="q",lower.tail=FALSE)
```
<li>`r formatC(11.2-zstar*6/sqrt(25),format="f",digits=1)`-`r formatC(11.2+zstar*6/sqrt(25),format="f",digits=1)`.</li>
<li>I am 95% confident that the true population mean is between `r formatC(11.2-zstar*6/sqrt(25),format="f",digits=1)` and `r formatC(11.2+zstar*6/sqrt(25),format="f",digits=1)`.</li>
</ol>

<li>See below.</li>
<ol type="a">
<li>C=0.99</li>
<li>Lower bound</li>
<li>Z*=`r formatC(distrib(0.99,type="q",lower.tail=FALSE),format="f",digits=3)`.</li>
```{r}
zstar <- distrib(0.99,type="q",lower.tail=FALSE)
```
<li>`r formatC(918+zstar*90/sqrt(80),format="f",digits=1)`.</li>
<li>I am 99% confident that the true population mean is greater than `r formatC(918+zstar*90/sqrt(80),format="f",digits=1)`.</li>
</ol>

<li>See below.</li>
<ol type="a">
<li>C=0.90</li>
<li>Upper bound</li>
<li>Z*=`r formatC(distrib(0.9,type="q"),format="f",digits=3)`.</li>
```{r}
zstar <- distrib(0.9,type="q")
```
<li>`r formatC(13700+zstar*8000/sqrt(50),format="f",digits=1)`.</li>
<li>I am 90% confident that the true population mean is less than `r formatC(13700+zstar*8000/sqrt(50),format="f",digits=1)`.</li>
</ol>

<li>See below.</li>
<ol type="a">
<li>C=0.99</li>
<li>Interval</li>
<li>Z*= $\pm$ `r formatC(distrib(0.005,type="q",lower.tail=FALSE),format="f",digits=3)`.</li>
```{r}
zstar <- distrib(0.005,type="q",lower.tail=FALSE)
```
<li>`r formatC(824-zstar*60/sqrt(25),format="f",digits=1)`-`r formatC(824+zstar*60/sqrt(25),format="f",digits=1)`.</li>
<li>I am 99% confident that the true population mean is between `r formatC(824-zstar*60/sqrt(25),format="f",digits=1)` and `r formatC(824+zstar*60/sqrt(25),format="f",digits=1)`.</li>
</ol>

<li>See below.</li>
<ol type="a">
<li>C=0.95</li>
<li>Upper bound</li>
<li>Z*=`r formatC(distrib(0.95,type="q"),format="f",digits=3)`.</li>
```{r}
zstar <- distrib(0.95,type="q")
```
<li>`r formatC(73+zstar*20/sqrt(30),format="f",digits=1)`.</li>
<li>I am 95% confident that the true population mean is less than `r formatC(73+zstar*20/sqrt(30),format="f",digits=1)`.</li>
</ol>

<li>See below.</li>
<ol type="a">
<li>C=0.95</li>
<li>Interval</li>
<li>Z*= $\pm$ `r formatC(distrib(0.025,type="q",lower.tail=FALSE),format="f",digits=3)`.</li>
```{r}
zstar <- distrib(0.025,type="q",lower.tail=FALSE)
```
<li>`r formatC(1370-zstar*800/sqrt(40),format="f",digits=1)`-`r formatC(1370+zstar*800/sqrt(40),format="f",digits=1)`.</li>
<li>I am 95% confident that the true population mean is between `r formatC(1370-zstar*800/sqrt(40),format="f",digits=1)` and `r formatC(1370+zstar*800/sqrt(40),format="f",digits=1)`.</li>
</ol>

</ol>

----

## Class Excercise 3
#### Body Temperatures
<ol>
```{r echo=-(4:5), fig.show='hide'}
d <- read.csv("https://github.com/droglenc/NCData/raw/master/NormTemp.csv")
Summarize(d$bodytemp,digits=2)
(zstar <- distrib(0.025,type="q",lower.tail=FALSE))
mn <- mean(d$bodytemp)
n <- nrow(d)
```
<li>`r formatC(mn-zstar*0.41/sqrt(n),format="f",digits=1)`-`r formatC(mn+zstar*0.41/sqrt(n),format="f",digits=1)`.</li>
<li>I am 95% confident that the true population mean body temperature is between `r formatC(mn-zstar*0.41/sqrt(n),format="f",digits=1)` and `r formatC(mn+zstar*0.41/sqrt(n),format="f",digits=1)`C.</li>
</ol>

#### Radius of Jupiter
<ol>
```{r fig.show='hide'}
(zstar <- distrib(0.005,type="q",lower.tail=FALSE))
```
<li>`r formatC(71492-zstar*28/sqrt(40),format="f",digits=1)`-`r formatC(71492+zstar*28/sqrt(40),format="f",digits=1)`.</li>
<li>I am 99% confident that the true population mean equatorial radius of Jupiter is between `r formatC(71492-zstar*28/sqrt(40),format="f",digits=0)` and `r formatC(71492+zstar*28/sqrt(40),format="f",digits=0)` km.</li>
</ol>

----

## Class Excercise 4
#### Internet Usage

<ol>
```{r}
sigma <- 6.95; me <- 1
(zstar <- distrib(0.025,type="q",lower.tail=FALSE))
(zstar*sigma/me)^2
```
<li>One should take a sample size of `r formatC(ceiling((zstar*sigma/me)^2),format="f",digits=0)` household to achieve the desired level of confidence and margin-of-error.</li>

```{r}
sigma <- 6.95; me <- 1
(zstar <- distrib(0.05,type="q",lower.tail=FALSE))
(zstar*sigma/me)^2
```
<li>One should take a sample size of `r formatC(ceiling((zstar*sigma/me)^2),format="f",digits=0)` household to achieve the desired level of confidence and margin-of-error.</li>

```{r}
sigma <- 6.95; me <- 0.5
(zstar <- distrib(0.025,type="q",lower.tail=FALSE))
(zstar*sigma/me)^2
```
<li>One should take a sample size of `r formatC(ceiling((zstar*sigma/me)^2),format="f",digits=0)` household to achieve the desired level of confidence and margin-of-error.</li>
</ol>

#### Blood Pressure in Children

<ol>
```{r}
sigma <- 20; me <- 5
(zstar <- distrib(0.005,type="q",lower.tail=FALSE))
(zstar*sigma/me)^2
```
<li>One should take a sample size of `r formatC(ceiling((zstar*sigma/me)^2),format="f",digits=0)` children to achieve the desired level of confidence and margin-of-error with the "worse-case" standard deviation of 20.</li>
```{r}
sigma <- 15; me <- 5
(zstar <- distrib(0.005,type="q",lower.tail=FALSE))
(zstar*sigma/me)^2
```
<li>One should take a sample size of `r formatC(ceiling((zstar*sigma/me)^2),format="f",digits=0)` children to achieve the desired level of confidence and margin-of-error with the "best-case" standard deviation of 15.</li>
</ol>

----
