---
title: "Answer Key"
subtitle: "Data Production CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/ce-keys/")
source("../../rhelpers/rhelpers.R")
modHTML("Data_Production")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../rhelpers/knitr_setup.R")
```

----

## Class Excercise 1
#### Study Types I
1. Experiment ... the people were given a treatment (placebo or fish oil)
1. Experiment ... the men were given a treatment (training program) with before/after results recorded
1. Observational ... the women were "followed" over time, no treatment was imparted (the B12 deficiency or not was naturally induced)
1. Observational ... no treatment was imparted (the brain cancer factor was naturally induced)
1. Experimental ... a treatment (animations or not) was imparted
1. Observational ... no treatment was imparted (alcohol use was a choice of the women)


#### Study Types II
1. Observational ... no treatment was imparted (work at night was a "choice" of the women)
1. Observational ... no treatment was imparted (observation of church records)
1. Experiment ... a treatment (type of trap) was imparted.

----

## Class Excercise 2
1. Convenience sample.
1. Voluntary response.
1. Simple random sample (probability-based sample).
1. Voluntary response.
1. Convenience sample.

----

## Class Excercise 3
#### Blood Pressure Study
1. Experimental ... a treatment is imposed.
1. A male subject.
1. Below ...
    1. Response ... blood pressure
    1. Factors ... Calcium (pill or not) and Exercise (do or don't)
    1. Levels ... 2 for calcium and 2 for exercise
    1. Treatments ... 4
    1. Replicates ... 8
1. Below
```{r}
alloc <- sample(32)
alloc[1:8]    # first treatment
alloc[9:16]   # second treatment
alloc[17:24]  # third treatment
alloc[25:32]  # fourth treatment
```

#### Softball Bats and Balls
1. Experimental ... imparting treatments (ball types and bats were chosen specifically)
1. Individual .. a "hit"
1. Below
    1. Response ... Distance the ball was hit.
    1. Factors ... Brand of ball and type of bat
    1. Levels ... 2 brands of ball and 3 types of bat
    1. Treatments ... 6
    1. Replicates ... 4 (balls of each brand by each type of bat)
1. I would first randomly order the bats and then randomly order the two brands for each bat.
```{r}
sample(3)  # random order of bats
sample(8)  # first order of balls ... 1-4 is first brand, 5-8 second
sample(8)  # second order of balls
sample(8)  # third order of balls
```

----

## Class Excercise 4
1. Response is "success rate", explanatory is distance of putt.
1. Response is "response time", explanatory is amount of time awake.
1. Response is growth, explanatory is amount of fertilizer.
1. It is not obvious which is the response or expanatory ... would depend on the researcher's question.
1. Response is sales of umbrellas, explanatory is amount of rainfall.
1. Response is weight, explanatory is age.
