---
title: "Answer Key"
subtitle: "Linear Regression CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/ce-keys/")
source("../../rhelpers/rhelpers.R")
modHTML("Linear_Regression")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../rhelpers/knitr_setup.R")
```

----

## Class Excercise 1

1. **Dishwasher Example**
    1. response - Height of suds (mm)
    1. explanatory - Amount of soap (g)
    1. slope is 12.4
    1. intercept is -20.2
    1. For each additional gram of soap, the height of sudes increases by 12.4 mm, on average.
    1. For 0 grams of soap, the height of suds is -20.2 mm, on average.  [*This is an extrapolation.*]
    1. Ask to predict height of suds for an amount of soap between 3.5 and 8.0 grams.  Example ... "What is the predicted height of suds for 5 g of soap?"  Answer ... -20.2+12.4*5=41.8 mm.
    1. Ask to predict height of sudes for an amount of soap not between 3.5 and 8.0 grams.  Example ... "What is the predicted height of suds for 10 g of soap?"
1. **Livestock Food Example**
    1. response - Costs
    1. explanatory - Livestock food mix in pounds
    1. Slope is 190.
    1. Intercept is 12075
    1. For each additional ton of the food mix, the cost increases by $190, on average.
    1. For 0 tons of the food mix, the costs are $12075, on average. [*This may be considered as "start-up" costs.*]
    1. Ask to predict costs for food mix amounts between 36.5 and 42 tons.  Example ... "What is the cost to make 40 tons of the mix?"  Answer ... 190*40+12075 = $19675.
    1. Ask to predict costs for food mix amounts not between 36.5 and 42 tons.  Example ... "What is the cos to make 70 tons of the mix?"
1. **Grey Kangaroo Example**
    1. response - Nasal width (x10)
    1. explanatory - Nasal length (x10)
    1. The slope is 0.2876.
    1. The intercept is 46.451.
    1. For every 1 unit (10 mm) increase in nasal length, the nasal width increases by 0.2876 units (i.e., 2.876 mm), on average.
    1. The nasal width when the nasal length is 0 mm is 46.451 units (i.e., 464.51 mm), on average.
    1. Ask to predict the nasal width when the nasal length (x10) is between 450 and 900.  Example ... "What is the predicted nasal width when the nasal length (x10) is 600 (i.e., 60 mm)?"  Answer ... Y=0.2876*600+46.451=219.01 (or 21.9 mm).
    1. Ask to predict the nasal width when the nasal length (x10) is not between 450 and 900.  Example ... "What is the predicted nasal width when the nasal length (x10) is 250 (i.e., 25 mm)?"

----

## Class Excercise 2

1. 

----

## Class Excercise 3

1. IBI.
1. Percent impervious surface.
1. The relationship between IBI and the percent of impervious surface is negative, nonlinear, and moderately strong with no obvious outliers.  I did not report the value of the correlation coefficient because of the nonlinear form.

```{r fig.width=3, fig.height=3, par1=TRUE}
d <- data.frame(imp=c(60,43,43,34,27,25,21,18,11,8,8,8,7,7,5,4,2),
                IBI=c(9,11,13,23,31,31,21,23,27,37,39,29,31,43,33,35,37))
plot(IBI~imp,data=d,pch=19,xlab="% Impervious Surface",ylab="IBI")
```

----

## Class Excercise 4

<ol>
<li>The relationship between number of coffees sold at the cafe and maximum daily temperature is negative, nonlinear, and moderately strong with no obvious outliers.  I did not report the value of the correlation coefficient because of the nonlinear form.</li>

```{r fig.width=3, fig.height=3, par1=TRUE}
d <- read.csv("cafedata.csv")
plot(coffee.sold~max.temp,data=d,pch=19,xlab="Maximum Daily Temperature",
     ylab="Number of Coffees Sold")
```

<li>The relationship between the number of coffees sold and the number of sodas sold at the cafe is negative, linear, and weak (r=-0.40) with no obvious outliers.  I identified strength from the value of the correlation coefficient because of the linear form and lack of outliers.</li>

```{r fig.width=3, fig.height=3, par1=TRUE}
d <- read.csv("cafedata.csv")
plot(coffee.sold~sodas.sold,data=d,pch=19,xlab="Number of Sodas Sold",
     ylab="Number of Coffees Sold")
corr(coffee.sold~sodas.sold,data=d,use="pairwise.complete.obs")
```

<li>The relationship between the number of coffees sold at the cafe and the number of days since the cafe opened is primarily negative and linear, though there appears to be some cyclic trends in the data (i.e., linear decline for first 20 days, linear increase for next 15 days, linear decline for next 25 days, and then a linear increase for last 12 days).  There is a weak outlier of a larger number of coffees sold on day 30.  The strength is only moderate when compared against an overall linear trend but is somewhat strong when compared against the cyclic trend mentioned previously.  The correlation coefficient was not reported due to outlier and cyclic trend present in these data.</li>

```{r fig.width=3, fig.height=3, par1=TRUE}
d <- read.csv("cafedata.csv")
plot(coffee.sold~days.since.start,data=d,pch=19,xlab="Days Since the Cafe Opened",
     ylab="Number of Coffees Sold")
lines(coffee.sold~days.since.start,data=d,col="gray90")
```

</ol>
