---
title: "Answer Key"
subtitle: "Sampling Distributions CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/ce-keys/")
source("../../rhelpers/rhelpers.R")
modHTML("Sampling_Distributions")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../rhelpers/knitr_setup.R")
```

----

## Class Excercise 1

```{r echo=FALSE, results='hide'}
d <- c(12,14,16,18,20)
mu <- mean(d)
sigma <- sd(d)

d2 <- combn(d,2)
d2mns <- combn(d,2,FUN=mean)
d2m <- mean(d2mns)
d2se <- sd(d2mns)

d3 <- combn(d,3)
d3mns <- combn(d,3,FUN=mean)
d3m <- mean(d3mns)
d3se <- sd(d3mns)

d4 <- combn(d,4)
d4mns <- combn(d,4,FUN=mean)
d4m <- mean(d4mns)
d4se <- sd(d4mns)

## Theoretical SEs need to be corrected for sampling from finite populations
ns <- 2:4
N <- length(d)
fpc <- sqrt((N-ns)/(N-1))
ses <- (sigma/sqrt(ns))*fpc
```

<ol>
  <li> $\mu$=`r formatC(mu,format="f",digits=0)`, $\sigma$=`r formatC(sigma,format="f",digits=3)`.</li>
  <li> All samples of n=2 are shown in columns below.</li>
```{r echo=FALSE}
d2
```
  <li> Means from all samples of n=2 are shown below.</li>
```{r echo=FALSE}
d2mns
```
  <li> The mean of all sample means from n=2 is `r formatC(d2m,format="f",digits=0)`.</li>
  <li> The standard deviation of all sample means from n=2 is `r formatC(d2se,format="f",digits=3)`.  This is really a standard error.</li>
  <li> The above for n=3 are below.</li>
```{r echo=FALSE}
d3
d3mns
d3m
d3se
```
  <li> The above for n=4 are below.</li>
```{r echo=FALSE}
d4
d4mns
d4m
d4se
```

  <li>Same.</li>
  <li>Did not change (were not affected by n).</li>
  <li>All smaller.</li>
  <li>Decreased with increasing n.</li>
</ol>

----

## Class Excercise 2

**NOT DONE**

----

## Class Excercise 3
#### Melatonin and Sleep

1. About a mean from n=4 ...
    1. Sampling.
    1. Can be answered.
    1. N(15,`r formatC(5/sqrt(4),format="f",digits=3)`)
    1. `r formatC(distrib(12,mean=15,sd=5/sqrt(4),lower.tail=FALSE),format="f",digits=4)`.
1. About an individual ...
    1. Population.
    1. Can be answered.
    1. N(15,5)
    1. `r formatC(distrib(12,mean=15,sd=5,lower.tail=FALSE),format="f",digits=4)`.
1. About a mean from n=25 ...
    1. Sampling.
    1. Can be answered.
    1. N(15,`r formatC(5/sqrt(25),format="f",digits=3)`)
    1. `r formatC(distrib(15.5,mean=15,sd=5/sqrt(25),lower.tail=FALSE),format="f",digits=4)`.

#### Pollutants in Auto Exhaust

1. About an individual ...
    1. Population.
    1. Can NOT be answered.
    1. Right-skewed (not normal).
    1. Can't answer.
1. About a mean from n=15 ...
    1. Sampling.
    1. Can NOT be answered.
    1. Not normal.
    1. Can't answer.
1. About a mean from n=50 ...
    1. Sampling.
    1. Can be answered.
    1. N(0.9,`r formatC(0.15/sqrt(50),format="f",digits=4)`)
    1. `r formatC(distrib(0.92,mean=0.9,sd=0.15/sqrt(50),lower.tail=FALSE),format="f",digits=4)`.
1. About a mean from n=50 ...
    1. Sampling.
    1. Can be answered.
    1. N(0.9,`r formatC(0.15/sqrt(50),format="f",digits=4)`)
    1. `r formatC(distrib(0.85,mean=0.9,sd=0.15/sqrt(50)),format="f",digits=4)`.
1. About a mean from n=50 ...
    1. Sampling.
    1. Can be answered.
    1. N(0.9,`r formatC(0.15/sqrt(50),format="f",digits=4)`)
    1. `r formatC(distrib(0.01,mean=0.9,sd=0.15/sqrt(50),lower.tail=FALSE,type="q"),format="f",digits=3)` g/mi.


----

## Class Excercise 4
#### Acorn Length

1. About an individual ...
    1. Population.
    1. Can NOT be answered.
    1. Right-skewed (not normal).
    1. Can't answer.
1. About a mean from n=20 ...
    1. Sampling.
    1. Can be answered.
    1. N(21,`r formatC(6/sqrt(20),format="f",digits=4)`)
    1. `r formatC(distrib(25,mean=21,sd=6/sqrt(20),lower.tail=FALSE),format="f",digits=4)`.
1. About a mean from n=50 ...
    1. Sampling.
    1. Can be answered.
    1. N(21,`r formatC(6/sqrt(50),format="f",digits=4)`)
    1. `r formatC(distrib(25,mean=21,sd=6/sqrt(50))-distrib(20,mean=21,sd=6/sqrt(50)),format="f",digits=4)`.
1. About a mean from n=6 ...
    1. Sampling.
    1. Can NOT be answered.
    1. Not normal.
    1. Can't answer.
1. About an individual ...
    1. Population.
    1. Can NOT be answered.
    1. Right-skewed (not normal).
    1. Can't answer.

#### People in Car

1. No.  Left side will be truncated at 0 (i.e., more than 2 SD away from mean becomes negative, which is not possible).  Discrete data with lots of ones and twos with few higher than that.
1. About a mean from n=40 ...
    1. Sampling.
    1. Can be answered.
    1. N(1.5,`r formatC(0.75/sqrt(40),format="f",digits=4)`)
    1. `r formatC(distrib(1.2,mean=1.5,sd=0.75/sqrt(40)),format="f",digits=4)`.
1. About a mean from n=50 ...
    1. Sampling.
    1. Can be answered.
    1. N(1.5,`r formatC(0.75/sqrt(50),format="f",digits=4)`)
    1. `r formatC(distrib(2,mean=1.5,sd=0.75/sqrt(50),lower.tail=FALSE),format="f",digits=7)`.
1. About a mean from n=5 ...
    1. Sampling.
    1. Can NOT be answered.
    1. Not normal.
    1. Can't answer.
1. About a mean from n=700 (through a total) ...
    1. Sampling.
    1. Can be answered.
    1. N(1.5,`r formatC(0.75/sqrt(700),format="f",digits=4)`)
    1. `r formatC(distrib(1075/700,mean=1.5,sd=0.75/sqrt(700),lower.tail=FALSE),format="f",digits=4)`.

#### Plowing Snow

1. About a mean from n=5 ...
    1. Sampling.
    1. Can NOT be answered.
    1. Not normal.
    1. Can't answer.
1. About an individual ...
    1. Population.
    1. Can NOT be answered.
    1. Right-skewed (not normal).
    1. Can't answer.
1. About a mean from n=40 ...
    1. Sampling.
    1. Can be answered.
    1. N(45,`r formatC(28/sqrt(40),format="f",digits=4)`)
    1. `r formatC(distrib(50,mean=45,sd=28/sqrt(40))-distrib(40,mean=45,sd=28/sqrt(40)),format="f",digits=4)`.
1. About a mean from n=40 ...
    1. Sampling.
    1. Can be answered.
    1. N(45,`r formatC(28/sqrt(40),format="f",digits=4)`)
    1. `r formatC(distrib(0.25,mean=45,sd=28/sqrt(40),type="q"),format="f",digits=1)` secs.
