---
title: "Answer Key"
subtitle: "Univariate EDA - Quantitative CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/ce-keys/")
source("../../rhelpers/rhelpers.R")
modHTML("UnivEDA_Quantitative")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../rhelpers/knitr_setup.R")
```

----

## Class Excercise 2
#### Histograms I

1. An individual 16-years and older in the U.S.
1. Mean Commute time (mins)
1. 51
1. 2
1. 25
1. 26-28
1. Symmetric (possibly slightly right-skewed)
1. No outliers

#### Histograms II

1. (Strongly) Right skewed, outlier at maximum
1. Left skewed, no outliers
1. Right skewed, outlier at maximum
1. Approximately symmetric, no outliers
1. Approximately symmetric, no outliers
1. (Srongly) Left skewed, no outliers
1. Left skewed, outlier at minimum
1. Bimodal (symmetric), no outliers

----

## Class Excercise 2
#### Hand Calculations

<ol>
<li>Shown below</li>
```{r}
sdCalc(c(18,28,25,21,16,24))
```
<li>Shown above</li>
<li>Shown below</li>
```{r}
iqrCalc(c(4,54,16,85,52,29,24,71,61,60,2))
```
<li>Shown above</li>
<li>Shown below</li>
```{r}
iqrCalc(c(93,81,34,5,54,84,54,13,1,35,79,63,97,71))
```

<li>Shown above</li>
<li>Range is 5 to 97.</li>
</ol>

----

## Class Excercise 3
1. NA
1. The mean is greater than the median for right-skewed distributions, less than the median for left-skewed distributions, and approximately equal to the median for approximately symmetric distributions.
1. Outliers attract the mean but have little effect on the median.  For example, if the outlier is at the maximum value then the mean will be greater than the median.
1. See below.
    1. Mean > Median
    1. Mean < Median (only slightly)
    1. Mean > Median
    1. Mean = Median (approximately)
    1. Mean = Median (approximately)
    1. Mean < Median
    1. Mean < Median
    1. Mean = Median (approximately)
1. The mean uses the actual values whereas the median is only concerned with relative positions.

----

## Class Excercise 4
<ol>
<li>The diameter of CWD is largely right-skewed with a separate group of 4 individuals at higher diameters (Figure 1).  With this definition of shape, there are no outliers.  The center as measured by the median is 18 cm and the IQR is from a Q1 of 16.25 to a Q3 of 20.75 cm.  I chose to use the median and IQR because of the right-skewed shape.</li>
```{r fig.width=3, fig.height=3, par1=TRUE}
d <- read.csv("cwd.csv")
hist(~diameter,data=d,xlab="Diameter of CWD")
Summarize(~diameter,data=d,digits=2)
```

<li>The diameter of CWD is largely right-skewed for both low- and medium-exposure sites (Figure 2).  There are no obvious outliers.  [<i>Note that it is difficult to describe shape and outliers in this case because the sample sizes in the two groups are so small.</i>]  The medians are 19 and 18, respectively, for the low- and medium-exposure sites.  The IQR is from a Q1 of 16 to a Q3 of 24 for the low-exposure sites and from a Q1 of 17 to a Q3 of 20 for the medium-exposure sites.  The median and IQR were used because of the skewed shape for both distributions.</li>
```{r fig.width=6, fig.height=3, par1=TRUE}
hist(diameter~exposure,data=d,xlab="Diameter of CWD")
Summarize(diameter~exposure,data=d,digits=2)
```
</ol>

----

## Class Excercise 5

**Not Yet Completed**
