---
layout: page
title: "Why Statistics Is Important"
subtitle: Handout
author: "Derek H. Ogle"
css: "/css/modules.css"
output:
  html_document:
    fig_width: 3.5
    fig_height: 3.5
    self_contained: no
---
```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/Why_Statistics_is_Important")
source("../../rhelpers/rhelpers.R")
fnm <- "CE"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message=FALSE}
fnm <- "CE"
source("../../rhelpers/knitr_setup.R")
```

The following code is used to produce a "live" example for class.  The meaning of this code will be explained later.  For now, we simply want the results that are produced.

Type the code exactly as shown in each grey box and attempt to answer the follow-up questions.

\ 

## Load the NCStats package
```{r echo=-2, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
data(ABCens90)
```

\ 

## Obtain a Random Sample of Residents from Ashland-Bayfield Counties
```{r results='hide'}
data(ABCens90)
( smp1 <- srsdf(ABCens90,50) )
```

* Were the individuals in your sample identical?  Why or why not?
* Can you quickly explain the "general" age and sex of individuals in your sample?  Why or why not?

\ 

## Summarize that Sample
```{r results='hide'}
Summarize(~age,data=smp1,digits=1)
Summarize(~sex,data=smp1)
```

* Can you quickly explain the "general" age and sex of individuals in your sample?  Why or why not?

\ 

## Obtain a Second Random Sample of Residents from Ashland-Bayfield Counties
```{r results='hide'}
( smp2 <- srsdf(ABCens90,50) )
Summarize(~age,data=smp2,digits=1)
Summarize(~sex,data=smp2)
```

* Were the mean ages from the two samples identical?
* Were the percentage of females from the two samples identical?
* Are your results typical or expected?  Why or why not?

\ 

## Compare Your Results to the Truth
```{r}
Summarize(~age,data=ABCens90,digits=1)
Summarize(~sex,data=ABCens90)
```

* Did your sample mean ages exactly equal this true value for the population?
* Did you sample percentages of females exactly equal this true percentage of females in the population?
* Are your results typical or expected?  Why or why not?

\ 

## Examine Many Random Samples of Residents from Ashland-Bayfield Counties
```{r echo=FALSE, results='hide', par1=TRUE}
set.seed(14787834)
hist(~replicate(1000,mean(srsdf(ABCens90,50)$age)),xlab="Mean Age",main="",col="gray90",breaks=15)
abline(v=mean(ABCens90$age),lwd=2,lty=2)
hist(~replicate(1000,Summarize(srsdf(ABCens90,50)$sex)["male","perc"]),xlab="Percent Male",main="",col="gray90",breaks=15)
abline(v=Summarize(ABCens90$sex)["male","perc"],lwd=2,lty=2)
```

```{r echo=FALSE, results="hide", message=FALSE, warnings=FALSE}
purl2(paste0(fnm,".Rmd"))
```
