\chapter{Statistical Symbols}
\vspace{-40pt}
This appendix contains statistical symbols used throughout the book.  The order of symbols generally follows the order in which the symbol was introduced in the text, although similar symbols are placed next to each other regardless of when they were introduced.

\begin{description}
  \item[$n$] -- Sample size.
  \item[$N$] -- Population size (generally unknown).
  \item[$\bar{x}$] -- Sample mean.
  \item[$\mu$] -- Population mean.
  \item[$Q1$] -- First quartile.
  \item[$Q3$] -- Third quartile.
  \item[$IQR$] -- Inter-quartile range.
  \item[$s$] -- Sample standard deviation.
  \item[$\sigma$] -- Population standard deviation.
  \item[$s^{2}$] -- Sample variance.
  \item[$\sigma^{2}$] -- Population variance.
  \item[$X\sim N(\mu,\sigma)$] -- The variable $X$ is normally distributed with a mean of $\mu$ and a standard deviation of $\sigma$.
  \item[$N(0,1)$] -- Standard normal distribution.
  \item[$Z$] -- (1) A variable that follows a $N(0,1)$  (2) The test statistic in a one-sample Z-test.
  \item[$C$] -- Level of confidence for a confidence interval.
  \item[$c$] -- Number of columns in the frequency table of a chi-square test.
  \item[$H_{0}$] -- Null hypothesis.
  \item[$H_{A}$] -- Alternative hypothesis.
  \item[$\alpha$] -- alpha.  (1) The probability of rejecting a true null hypothesis.  (2) The user-defined size of the rejection region in statistical hypothesis testing.
  \item[$\beta$] -- beta.  The probability of not rejecting a false null hypothesis.
  \item[$\mu_{0}$] -- Specific value of $\mu$ found in the null hypothesis for 1-sample z- and 1-sample t-tests.
  \item[$t$] -- (1) A variable that follows a t-distribution.  (2) The test statistic in t-tests.
  \item[$df$] -- Degrees-of-freedom (used for t- and $\chi^{2}$ distributions.
  \item[$\hat{p}$] -- Sample proportion.
  \item[$p$] -- Population proportion.
%  \item[$p_{0}$] -- Specific value of $p$ found in a null hypothesis.
  \item[$r$] -- (1) Sample correlation coefficient.  (2) Number of rows in the frequency table of a chi-square test.
%  \item[$\rho$] -- Population correlation coefficient.
  \item[$\hat{b}_{0}$] -- Sample y-intercept of regression line.
  \item[$\hat{b}_{1}$] -- Sample slope of regression line.
  \item[$\beta_{0}$] -- Population y-intercept of regression line.
  \item[$\beta_{1}$] -- Population slope of regression line.
  \item[$r^{2}$] -- Coefficient of determination.
  \item[$\hat{y}_{i}$] -- Predicted value of the response variable for the $i$th individual.
  \item[$RSS$] -- Residual sum-of-squares.
  \item[$s_{p}^{2}$] -- Pooled sample variance used in a 2-sample t-test.
%  \item[$\bar{x}_{diff}$] -- Sample mean of the difference in paired measurements.
%  \item[$s_{diff}$] -- Sample standard deviation of the differences in paired measurements.
  \item[$\chi^{2}$] -- (1) A variable that follows a chi-square distribution.  (2) The test statistic in chi-square tests.
\end{description}
