---
layout: page
title: "Introduction"
author: "Derek H. Ogle"
output:
  html_document:
    fig_width: 3.5
    self_contained: no
  pdf_document:
    fig_width: 3.5
  word_document:
    fig_width: 3.5
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/lecture/HOs"); source("../../rhelpers/rhelpers.R")
fnm <- "UnivariateEDA"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message=FALSE}
fnm <- "UnivariateEDA"
source("../../rhelpers/knitr_setup.R")
```

## Background
Measurements of the levels of arsenic in the drinking water, cooking water, and toenail samples, as well as related covariates, were measured on 21 individuals with private wells in a New Hampshire community.  The variables below were recorded in the \href{https://sites.google.com/site/ncstats/data/Arsenic.txt}{Arsenic.txt} file located on the R Resources web page.

* `age`: Age (yrs) of person
* `sex`: Sex of person
* `usedrink`: Household well used for drinking (A="$$<$$1/4", B="$$\approx$$1/4",  C="$$\approx$$1/2",  D="$$\approx$$3/4",  E="$$>$$3/4")
* `usecook`: Household well used for cooking (A="$$<$$1/4", B="$$\approx$$1/4",  C="$$\approx$$1/2",  D="$$\approx$$3/4",  E="$$>$$3/4")
* `arswater`: Arsenic in water (ppm)
* `arsnails`: Arsenic in toenails (ppm)

## Initialization}
```{r message=FALSE, warning=FALSE}
library(NCStats)
setwd("C:/aaaWork/Class Materials/MTH107/Lecture/HOs/")
Ars <- read.csv("Arsenic.csv")
str(Ars)
view(Ars)
```


## Univariate EDA -- Quantitative}
```{r hist}
Summarize(~age,data=Ars,digits=2)
hist(~age,data=Ars,main="",xlab="Age (yrs)")
```

\vspace{18pt}
## Univariate EDA -- Quantitative (Separated by Groups)}
```{r mhist, fig.width=7, fig.height=3.5, out.width='.8\\linewidth', par2=TRUE}
Summarize(age~sex,data=Ars,digits=2)
hist(age~sex,data=Ars,xlab="Age (yrs)",col="gray90")
```

## Univariate EDA -- Categorical}
```{r barchart}
( tbl.drink <- xtabs(~usedrink,data=Ars) )
percTable(tbl.drink,digits=1)
barplot(tbl.drink,xlab="Rating of Use for Drinking",ylab="Frequency",col="gray90")
```

```{r echo=FALSE, results="hide", message=FALSE, warnings=FALSE}
purl2(paste0(fnm,".Rmd"))
```
