---
layout: page
title: "Answer Key"
subtitle: "1-Sample Z CE"
author: Derek H. Ogle
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/1_Sample_Z/")
source("../../rhelpers/rhelpers.R")
modHTML("CE_Keys")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../rhelpers/knitr_setup.R")
library(captioner)
tbls <- captioner(prefix="Table")
tbls(name="HawksRidge","Results from the one-sample z-test for testing that the mean number of hawks observed at Hawk Ridge, MN is greater than 50000 per year.")
```

----

## Diet and Weight Gain
```{r DietCalc, echo=FALSE, results='hide', fig.show='hide'}
xbar <- 6.7; sigma <- 7.1; n <- 35
se <- sigma/sqrt(n)
t <- (xbar-5)/se
pv <- distrib(t,lower.tail=FALSE)
lcb <- xbar+distrib(0.95,lower.tail=FALSE,type="q")*se
```

<ol>
<li>&alpha;=0.05.</li>
<li>H<sub>A</sub>: &mu;>5 and H<sub>0</sub>: &mu;=5, where &mu; is the mean weight gain for all women that followed the diet.</li>
<li>1-sample Z-test because (i) one population (women that followed this diet) was sampled, (ii) a quantitative variable (weight gain) was recorded, and (iii) &sigma; is assumed to be known (=7.1).</li>
<li>An observational study as the women chose to be on the diet.  There is clear randomization in this study.</li>
<li>The assumptions are met because (i) &sigma; is known (=7.1) and (ii) n=35>30.</li>
<li>x&#772;=`r xbar` pounds gained.</li>
<li>Z=`r formatC(t,format="f",digits=4)`.</li>
<li>p-value=`r kPvalue(pv,include.p=FALSE,latex=FALSE)`.</li>
<li>Do not reject H<sub>0</sub> because the p-value>&alpha;.</li>
<li>There is not enough evidence to conclude that the mean weight gain for ALL women on this diet is greater than 5 pounds.</li>
<li>Not required</li>
</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
<<TumorCalc>>
```

----

## Raptors at Hawk's Ridge

```{r HawksRidgeCode, echo=FALSE, results='hide'}
hr <- read.csv("https://github.com/droglenc/NCData/raw/master/HawksRidge.csv")
( hr.z <- z.test(hr$Total,mu=50000,sd=37000,alt="greater",conf.level=0.99) )
```

<ol>
<li>&alpha;=0.01</li>
<li>H<sub>O</sub>: &mu;=50000, H<sub>A</sub>: &mu;>50000, where &mu; is the mean number of raptors seen in ALL years at Hawk's Ridge.</li>
<li>A one-sample Z-test is required because (i) a quantitative variable (number of raptors) was measured, (ii)the individuals are from one population (all years at Hawk's Ridge), and (iii) &sigma; is known (=37000).  Thus, the population mean is being compared to a specific value in the null hypothesis.</li>
<li>An observational sample of years that was not random was taken.</li>
<li>The assumptions are met because n=32>30 and &sigma; is known.</li>
<li>x&#772;=`r formatC(hr.z$estimate,format="f",digits=0)` (`r tbls("HawksRidge",display="cite")`).</li>

`r tbls("HawksRidge")`

```{r echo=FALSE}
kHtest(hr.z)
```
<li>The test statistic is Z=`r formatC(hr.z$statistic,format="f",digits=3)` (`r tbls("HawksRidge",display="cite")`).</li>
<li>The p-value is `r kPvalue(hr.z$p.value,include.p=FALSE,latex=FALSE)` (`r tbls("HawksRidge",display="cite")`).</li>
<li>The H<sub>O</sub> is rejected because the p-value<&alpha; (`r tbls("HawksRidge",display="cite")`).</li>
<li>The mean number of raptors seen per year in ALL years appears to be greater than 50000.</li>
<li>I am 99\% confident that the mean number of raptors seen per year in ALL years is more than `r formatC(hr.z$conf.int[1],format="f",digits=0)` (`r tbls("HawksRidge",display="cite")`).</li>
</ol>

R Appendix
```{r eval=FALSE, prompt=FALSE}
<<HawksRidgeCode>>
```

----
