---
title: "Answer Key"
subtitle: "Bivariate EDA Categorical CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/BivEDA_Categorical/")
source("../../rhelpers/rhelpers.R")
modHTML("CE_Keys")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../rhelpers/knitr_setup.R")
library(captioner)
tbls <- captioner(prefix="Table")
tbls(name="TreesRaw",'Raw freuency table for stands by type ("sttype": c="coniferous",mc="mostly coniferous",md="mostly deciduous", and d="deciduous" and years since last fire category (tslf).')
tbls(name="TreesRow",'Row percentages table for stands by type ("sttype": c="coniferous",mc="mostly coniferous",md="mostly deciduous", and d="deciduous" and years since last fire category (tslf).')
tbls(name="TreesCol",'Column percentages table for stands by type ("sttype": c="coniferous",mc="mostly coniferous",md="mostly deciduous", and d="deciduous" and years since last fire category (tslf).')
tbls(name="TreesTot",'Total percentages table for stands by type ("sttype": c="coniferous",mc="mostly coniferous",md="mostly deciduous", and d="deciduous" and years since last fire category (tslf).')
```

----

## Hand and R Calculations

<ol>
<li>Two-way frequency table is below.</li>
```{r echo=FALSE}
d <- data.frame(VAR1 <- c("D","C","A","C","A","D","C","A","D","B","A","B",
                          "B","A","C","D","A","C","A","B","B","A","D","C","A","C","B","B"),
                VAR2 <- c("c","c","b","c","a","c","a","b","c","b","c","c","b","c","c",
                          "c","c","c","b","b","a","b","c","b","a","b","a","b"))
```
```{r eval=FALSE}
d <- read.csv("BivEDA_CE1.csv")
```
```{r}
( tbl <- xtabs(~VAR2+VAR1,data=d))
```

<li>Row Percentages Table</li>
```{r}
percTable(tbl,margin=1,digits=1)
```

<li>Column Percentages Table</li>
```{r}
percTable(tbl,margin=2,digits=1)
```

<li>Total Percentages Table</li>
```{r}
percTable(tbl,digits=1)
```
</ol>

----

## Forest Composition

The questions below use the following tables.
```{r echo=FALSE}
d <- read.csv("https://github.com/droglenc/NCData/raw/master/Fire.csv")
```
```{r eval=FALSE}
d <- read.csv("Fire.csv")
```
```{r}
d$sttype <- factor(d$sttype,levels=c("c","mc","md","d"))
## Raw Frequency Table
(tbl <- xtabs(~tslf+sttype,data=d))
## Row Percentage Table
percTable(tbl,margin=1,digits=1)
## Column Percentage Table
percTable(tbl,margin=2,digits=1)
## TotalPercentage Table
percTable(tbl,digits=1)
```

1. 15 stands
1. 37.5%
1. 50%
1. 14.6%
1. 46.7%
1. 31.2%
1. 15.4%
1. Those with more recent fires tended to be deciduous, those without recent fires tended to be coniferous.

----

## Class Excercise 3

The questions below use the following tables.

```{r echo=FALSE}
d <- read.csv("https://github.com/droglenc/NCData/raw/master/PopularKids.csv")
```
```{r eval=FALSE}
d <- read.csv("PopularKids.csv")
```
```{r}
## Raw Frequency Table
(tbl <- xtabs(~grade+goals,data=d))
## Row Percentage Table
percTable(tbl,margin=1,digits=1)
## Column Percentage Table
percTable(tbl,margin=2,digits=1)
## TotalPercentage Table
percTable(tbl,digits=1)
```

1. 119 students
1. 11.5%
1. 21.0%
1. 50.0%
1. 51.7%
1. 39.0%
1. Fairly even distribution of students into the goals categories among the grades.  Thus, there is no real difference in goals by grade in school.
