---
title: "Review Exercise Key"
subtitle: "1-Sample t-Test"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/RE/zRevExAns/")
source("../../../rhelpers/rhelpers.R")
modHTML("1Samplet")
```
```{r echo=FALSE, message=FALSE}
library(NCStats)
library(captioner)
tabs <- captioner(prefix="Table")
figs <- captioner(prefix="Figure")
```

----

## Achievement Test Scores

1. &alpha; = 0.05.
1. H<sub>0</sub> &mu;=80, H<sub>A</sub> &mu;>80 where &mu; is the average achievement test score for all students at the superintendent's school.
1. A one-sample t-test is required because a quantitative variable (achievement test score) was measured on individuals from one group/population (one school), the population mean is compare to a specific value in the null hypothesis, and &sigma; is UNknown.
1. An observational study with randomly selected subjects was used.
1. The &sigma; is unknown. The sample size is not greater than 40; however, it is greater than 15. Unfortunately the only clue that the sample distribution is not strongly skewed is the fact that the mean and median are approximately equal. I would prefer to have a histogram to look at.
1. The statistic is x&#772; = 83.2.
1. The test statistic is t=$\frac{83.2-80}{\frac{5.5}{\sqrt{32}}}$ = 3.29 with 32-1 = 31 df.
1. The p-value is `r kPvalue(pt(3.29,df=31,lower.tail=FALSE),include.p=FALSE,latex=FALSE)`.
1. The H<sub>0</sub> is rejected because the p-value < &alpha;.
1. It appears that the mean achievement score of all students at the superintendent's school is greater than 80 points.
1. A 95% lower confidence bound is warranted in this situation. The t<sup>*</sup> = `r formatC(qt(0.95,df=31,lower.tail=FALSE),format="f",digits=3)`. Thus, 83.2-1.696 $\frac{5.5}{\sqrt{32}}$ or 83.2-1.65 = 81.455. One is 95% confident that the mean achievement score for all students at the superintendent's school is greater than 81.46 points.

### R Appendix
```{r eval=FALSE}
( pval <- distrib(3.29,distrib="t",df=31,lower.tail=FALSE) )
( tstar <- distrib(0.95,distrib="t",type="q",df=31,lower.tail=FALSE) )
```

----

## Salaries of College Graduates

1. &alpha; = 0.10.
1. H<sub>0</sub> &mu; = 32, H<sub>A</sub> &mu; > 32 where &mu; is the mean salary in 1000s of dollars.
1. A one-sample t-test is required because a quantitative variable (salary) was measured on individuals from one group/population (Northwestern University), the population mean is compared to a specific value in the null hypothesis, and &sigma; is UNknown.
1. An observational study with apparent random selection of individuals (background says "random surveys") was used.
1. The &sigma; is unknown. The sample size is greater than 40; thus, the sampling distribution of the test statistic should follow a t-distribution.
1. The statistic is x&#772; = 32.511.
1. The test statistic is t=$\frac{32.511-32}{\frac{1.713}{\sqrt{42}}}$ = 1.933 with 42-1 = 41 df.
1. The p-value is `r kPvalue(pt(1.933251,df=41,lower.tail=FALSE),include.p=FALSE,latex=FALSE)`.
1. The H<sub>0</sub> is rejected because the p-value < &alpha;.
1. It appears that the average salary for all recent graduates from the College of Liberal Arts is greater than $32,000; thus, the Dean's statement is supported.
1. A 90% lower confidence bound is warranted in this situation. The t<sup>*</sup> = `r formatC(qt(0.90,df=41,lower.tail=FALSE),format="f",digits=3)`
Thus, 32.511-1.303 $\frac{1.713}{\sqrt{42}}$ or 32.511-0.344 = 32.167. One is 90% confident that the mean salary for all recent graduates from the College of Liberal Arts is greater than $32,167.

### R Appendix
```{r eval=FALSE}
( pval <- distrib(2.085,distrib="t",df=41,lower.tail=FALSE) )
( tstar <- distrib(0.90,distrib="t",type="q",df=41,lower.tail=FALSE) )
```

----

## Fishing Line Strength
```{r echo=FALSE, results='hide'}
fl <- data.frame(lbs=c(6.1,5.3,5.5,4.9,6.2,6.5,5.7,5.5,4.7,6.2,6.8,5.9,5.8,6.7,6.3,6.2,5.4,5.5,6.7,5.9))
figs(name="FishingLine","Histogram of the breaking strength (lbs) of 6-lb fishing line.")
tabs(name="FishingLine","Results of 1-Sample t-Test for breaking strength (lbs) of 6-lb fishing line.")
```
```{r tFLS, echo=FALSE, results='hide'}
( fl.t <- t.test(fl$lbs,mu=6,alt="less",conf.level=0.90) )
```

1. &alpha; = 0.10.
1. H<sub>0</sub> &mu; = 6, H<sub>A</sub> &mu; < 6 where &mu; is the mean breaking weight of the  fishing line.
1. A one-sample t-test is required because a quantitative variable (breaking weight) was measured on individuals from one group/population (one type of line), the population mean is compared to a specific value in the null hypothesis, and &sigma; is UNknown.
1. An experimental study with randomly selected pieces of line was used.
1. The &sigma; is unknown. The sample size (=`r nrow(fl)`) is not greater than 40, however, it is greater than 15. The histogram suggests that the sample distribution is roughly symmetric (`r figs("FishingLine",display="cite")`). Thus, the assumptions are adequately met and the sampling distribution of the test statistic should follow a t-distribution.
1. The statistic is x&#772; = `r formatC(fl.t$estimate,format="f",digits=2)` (`r tabs("FishingLine",display="cite")`).
1. The test statistic is t = `r formatC(fl.t$statistic,format="f",digits=3)` with `r fl.t$parameter` df (`r tabs("FishingLine",display="cite")`).
1. The p-value is `r kPvalue(fl.t$p.value,include.p=FALSE,latex=FALSE)` (`r tabs("FishingLine",display="cite")`).
1. The H<sub>0</sub> is not rejected because the p-value > &alpha;.
1. It does not appear that the line breaks at average pressures less than 6 lbs; the manufacturer's claim is supported.
1. A confidence region is not required because H<sub>0</sub> was not rejected.

```{r histFLS, echo=FALSE}
hist(~lbs,data=fl,xlab="Fishing Line Break Weight (lbs)")
```

`r figs("FishingLine")`

`r tabs("FishingLine")`

```{r echo=FALSE, background='white', comment=""}
kHtest(fl.t)
```

### R Appendix
```{r eval=FALSE}
fl <- read.csv("data/FishingLine.csv")
<<histFLS>>
<<tFLS>>
```

----

## Strawberries Produced
```{r echo=FALSE, results='hide'}
sb <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/Strawberries.csv")
tabs(name="Strawberries","Results of 1-Sample t-Test for number of strawberries produced.")
```
```{r tSB, echo=FALSE, results='hide'}
( sb.t <- t.test(sb$berries,mu=12,alt="less",conf.level=0.90) )
```

1. &alpha; = 0.10.
1. H<sub>0</sub> &mu; = 12, H<sub>A</sub> &mu; < 12 where &mu; is the mean number of strawberries produced per plant.
1. A one-sample t-test is required because a quantitative variable (number of berries) was measured on individuals from one group/population (just my strawberries), the population mean is compared to a specific value in the null hypothesis, and &sigma; is UNknown.
1. An observational study with randomly selected individuals was used.
1. The &sigma; is unknown. The sample size (=`r nrow(sb)`) is greater than 40. Thus, the sampling distribution of the test statistic should follow a t-distribution.
1. The statistic is x&#772; = `r formatC(sb.t$estimate,format="f",digits=2)` (`r tabs("Strawberries",display="cite")`).
1. The test statistic is t = `r formatC(sb.t$statistic,format="f",digits=3)` with `r sb.t$parameter` df (`r tabs("Strawberries",display="cite")`).
1. The p-value is `r kPvalue(sb.t$p.value,include.p=FALSE,latex=FALSE)` (`r tabs("Strawberries",display="cite")`).
1. The H<sub>0</sub> is rejected because the p-value < &alpha;.
1. It appears that the plants produce fewer than 12 berries per plant, on average. The companies claim does not seem to be supported.
1. A 90% upper confidence bound is `r formatC(sb.t$conf.int[2],format="f",digits=2)`. Thus, one is 90% confident that the mean number of berries produces is less than `r formatC(sb.t$conf.int[2],format="f",digits=2)` (`r tabs("Strawberries",display="cite")`).

`r tabs("Strawberries")`

```{r echo=FALSE, background='white', comment=""}
kHtest(sb.t)
```

### R Appendix
```{r eval=FALSE}
sb <- read.csv("data/Strawberries.csv")
<<tSB>>
```

----

## Putting Together Toys
```{r echo=FALSE, results='hide'}
tt <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/ToyTime.csv")
figs(name="Toys","Histogram of the amount of time needed to put a toy together.")
tabs(name="Toys","Results of 1-Sample t-Test for the amount of time needed to put a toy together.")
```
```{r tTT, echo=FALSE, results='hide'}
( tt.t <- t.test(tt$time,mu=20,alt="greater",conf.level=0.90) )
```

1. &alpha; = 0.10.
1. H<sub>0</sub> &mu; = 20, H<sub>A</sub> &mu; > 20 where &mu; is the mean time to put the toy together.
1. A one-sample t-test is required because a quantitative variable was measured (time) on individuals from one group/population (this toy), the population mean is compared to a specific value in the null hypothesis, and &sigma; is UNknown.
1. An observational study with randomly selected individuals was used.
1. The &sigma; is unknown. The sample size (=`r nrow(tt)`) is not greater than 40, however it is greater than 15. The histogram suggests that the sample distribution is roughly symmetric (`r figs("Toys",display="cite")`). Thus, the assumptions are adequately met and the sampling distribution of the test statistic should follow a t-distribution.
1. The statistic is x&#772; = `r formatC(tt.t$estimate,format="f",digits=2)` (`r tabs("Toys",display="cite")`).
1. The test statistic is t = `r formatC(tt.t$statistic,format="f",digits=3)` with `r tt.t$parameter` df (`r tabs("Toys",display="cite")`).
1. The p-value is `r kPvalue(tt.t$p.value,include.p=FALSE,latex=FALSE)` (`r tabs("Toys",display="cite")`).
1. The H<sub>0</sub> is not rejected because the p-value > &alpha;.
1. It appears that the toy does not take more than 20 minutes, on average, to assemble; thus, this toy should not be rated as "difficult" to assemble.
1. No confidence region is required because H<sub>0</sub> was not rejected.

```{r histTT, echo=FALSE}
hist(~time,data=tt,xlab="Time to Assemble Toy (mins)")
```

`r figs("Toys")`

`r tabs("Toys")`

```{r echo=FALSE, background='white', comment=""}
kHtest(tt.t)
```

### R Appendix
```{r eval=FALSE}
tt <- read.csv("data/ToyTime.csv")
<<histTT>>
<<tTT>>
```

----

## Boat Traffic
```{r echo=FALSE, results='hide'}
y <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/Yahara.csv")
tabs(name="Boats","Results of 1-Sample t-Test for the number of boats on the Yahara locks.")
```
```{r prepYahara, echo=FALSE, results='hide', warning=FALSE}
y1 <- filterD(y,year==2005 & mon>5 & mon<9)
xtabs(~year+mon,data=y1)  # A check
```
```{r tYahara, echo=FALSE, results='hide'}
( y.t <- t.test(y1$boats.total,mu=75,alt="greater",conf.level=0.95) )
```

1. &alpha; = 0.05.
1. H<sub>0</sub> &mu; = 75, H<sub>A</sub> &mu; > 75 where &mu; is the mean number of boats through the locks.
1. A one-sample t-test is required because a quantitative variable (number of boats) was measured on individuals from one group/population (Yahara locks), the population mean is compare to a specific value in the null hypothesis, and &sigma; is UNknown.
1. An observational study that is very likely not random was used.
1. The &sigma; is unknown. The sample size (=`r nrow(y1)`) is greater than 40. Thus, the assumptions are met and the sampling distribution of the test statistic should follow a t-distribution.
1. The statistic is x&#772; = `r formatC(y.t$estimate,format="f",digits=2)` (`r tabs("Boats",display="cite")`).
1. The test statistic is t = `r formatC(y.t$statistic,format="f",digits=3)` with `r y.t$parameter` df (`r tabs("Boats",display="cite")`).
1. The p-value is `r kPvalue(y.t$p.value,include.p=FALSE,latex=FALSE)` (`r tabs("Boats",display="cite")`).
1. The H<sub>0</sub> is rejected because the p-value < &alpha;.
1. It appears that the mean number of boats through the locks per day during June, July, and August of 2005 is indeed greater than 75.
1. In fact, one is 95% confident that the average number of boats through the locks per day during June, July, and August of 2005 is at least `r formatC(y.t$conf.int[1],format="f",digits=2)` (`r tabs("Boats",display="cite")`).

`r tabs("Boats")`

```{r echo=FALSE, background='white', comment=""}
kHtest(y.t)
```


### R Appendix
```{r eval=FALSE}
y <- read.csv("data/Yahara.csv")
<<prepYahara>>
<<tYahara>>
```

----

## Golden Triangle in Beads?
```{r echo=FALSE, results='hide'}
d <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/Shoshoni.csv")
figs(name="Beads","Histogram of the width-to-length ratio of Shoshoni beads.")
tabs(name="Beads","Results of 1-Sample t-Test for the width-to-length ratio of Shoshoni beads")
```
```{r tBeads, echo=FALSE, results='hide'}
( d.t <- t.test(d$ratios,mu=0.618) )
```

1. &alpha; = 0.05.
1. H<sub>0</sub> &mu; = 0.618, H<sub>A</sub> &mu; &ne; 0.618 where &mu; is the mean width-to-length ratio of the Shoshoni beads.
1. A one-sample t-test is required because a quantitative variable (width-to-length ratio) was measured on individuals from one group/population (Shoshoni), the population mean is compare to a specific value in the null hypothesis, and &sigma; is UNknown.
1. An observational study that is very likely not random was used.
1. The &sigma; is unknown. The sample size (=`r nrow(d)`) is greater than 15 and the sample distribution is somewhat right-skewed with minor outliers (`r figs("Beads",display="cite")`). The assumptions are marginally met and the sampling distribution of the test statistic will likely follow a t-distribution.
1. The statistic is x&#772; = `r formatC(d.t$estimate,format="f",digits=2)` (`r tabs("Beads",display="cite")`).
1. The test statistic is t = `r formatC(d.t$statistic,format="f",digits=3)` with `r d.t$parameter` df (`r tabs("Beads",display="cite")`).
1. The p-value is `r kPvalue(d.t$p.value,include.p=FALSE,latex=FALSE)` (`r tabs("Beads",display="cite")`).
1. The H<sub>0</sub> is not rejected because the p-value > &alpha;.
1. It appears that the mean width-to-length ratio of the Shoshoni beads is NOT different from 0.618. Thus, it appears that the Shoshoni beads do follow the golden rectangle.
1. Confidence region not required by H<sub>0</sub> was not rejected.

```{r histBeads, echo=FALSE}
hist(~ratios,data=d,xlab="Width-to-Length Ratio of Beads")
```

`r figs("Beads")`

`r tabs("Beads")`

```{r echo=FALSE, background='white', comment=""}
kHtest(d.t)
```

### R Appendix
```{r eval=FALSE}
d <- read.csv("data/shoshoni.csv")
<<histBeads>>
<<tBeads>>
```

----
