---
title: "Review Exercise Key"
subtitle: "2-Sample t-Test"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/RE/zRevExAns/")
source("../../../rhelpers/rhelpers.R")
modHTML("2Samplet")
```
```{r echo=FALSE, message=FALSE}
library(NCStats)
library(captioner)
tabs <- captioner(prefix="Table")
figs <- captioner(prefix="Figure")
```

----

## Popcorn Yield by Variety

1. &alpha; = 0.10.
1. The H<sub>0</sub>:&mu;<sub>w</sub>-&mu;<sub>y</sub>=0 where &mu; is the mean number of unpopped kernels, w represents white kernels, and y represents yellow kernels (thus, positive numbers represent more unpopped white kernels). The H<sub>A</sub>:&mu;<sub>w</sub>-&mu;<sub>y</sub>&ne;0.
1. A two-sample t-test is required because (i) a quantitative variable (kernels unpopped) was measured, (ii) two groups/populations (yellow and white) exist, and (iii) those populations were independent (there is no connection between all of the individuals of white corn and those of yellow corn).
1. The data appear to be part of an observational study with the individuals randomly selected.
1. Individuals in the groups/populations are independent (as stated above). The sample size (n<sub>w</sub>+n<sub>y</sub>=60) is &ge; 40. The population variances appear to be equal because the p-value for Levene's test (given as 0.972) is "large" (i.e., > &alpha;). Therefore, the test statistic computed below should reasonably follow a t-distribution with n<sub>w</sub>+n<sub>y</sub>-2 = 58 df.
1. The statistic is x&#772;<sub>w</sub>-x&#772;<sub>y</sub> = 4.267-3.567 = 0.700. The pooled sample variance is $s_{p}^{2}=\frac{(30-1)4.456^{2}+(30-1)4.485^{2}}{30+30-2}$ = 19.986. The standard error of the statistic is $SE_{\bar{x}_{w}-\bar{x}_{y}}=\sqrt{19.986\left(\frac{1}{30}+\frac{1}{30} \right)}$ = 1.154.
1. The test statistic is t = $\frac{0.700-0}{1.154}$ = 0.606 with 58 df.
1. The p-value is `r kPvalue(2*pt(0.606,df=58,lower.tail=FALSE),include.p=FALSE,latex=FALSE)`.
1. The H<sub>0</sub> is not rejected because p-value > &alpha;.
1. There does not appear to be a difference in the mean number of unpopped kernels between the white and yellow varieties.
1. One is 90% confident that the difference in mean number of unpopped kernels between the white and yellow varieties is between -1.610 and 3.010 (as computed with 0.700&pm;2.002*1.154). This further illustrates that there is not a difference in mean number of unpopped kernels between the white and yellow varieties (because 0 is in the confidence region).

### R Appendix
```{r eval=FALSE}
( pval <- 2*distrib(0.60,distrib="t",df=58,lower.tail=FALSE) )
( tstar <- distrib(0.025,distrib="t",df=58,lower.tail=FALSE,type="q") )
```

----

## Heating System Efficiency

1. &alpha; = 0.10.
1. The H<sub>0</sub>:&mu;<sub>e</sub>-&mu;<sub>t</sub>=0 where &mu; is the mean BTUs and the subscripts represent the two types of vents (positive numbers represent more BTUs used for houses fit with electric vent). The H<sub>A</sub>:&mu;<sub>e</sub>-&mu;<sub>t</sub>&ne;0.
1. A two-sample t-test is required because (i) a quantitative variable (BTUs) was measured, (ii) two groups/populations (electric and thermal vents) exist, and (iii) those populations were independent (there is no connection between all of the vents in the two groups).
1. The data are part of an observational study where it is not obvious that the devices were randomly allocated to houses.
1. The individuals in the two groups are independent (no vent was placed in the same house). The sample size (n<sub>e</sub>+n<sub>t</sub> = 90) is &ge; 40. The population variances appear to be equal because the p-value for Levene's test of the homogeneity of variance test (given as 0.996) is "large" (i.e., > &alpha;). Therefore, the test statistic computed below should reasonably follow a t-distribution with n<sub>e</sub>+n<sub>t</sub>-2 = 88 df.
1. The statistic is x&#772;<sub>e</sub>-x&#772;<sub>t</sub> = 9.908 - 10.143 = -0.235. The pooled sample variance is $s_{p}^{2}=\frac{(40-1)3.020^{2}+(50-1)2.767^{2}}{40+50-2}$ = 8.305. The standard error of the statistic is $SE_{\bar{x}_{e}-\bar{x}_{t}}=\sqrt{8.305\left(\frac{1}{40}+\frac{1}{50} \right)}$ = 0.611.
1. The test statistic is t = $\frac{-0.235-0}{0.611}$ = -0.385 with 88 df.
1. The p-value is `r kPvalue(2*pt(-0.385,df=88),include.p=FALSE,latex=FALSE)`.
1. The H<sub>0</sub> is not rejected because the p-value > &alpha;.
1. There does not appear to be a difference in average BTUs between houses equipped with the electrically and the thermally activated vent.
1. One is 90% confident that the difference in average BTUs between houses equipped with the electrically and the thermally activated vent is between -0.875 and 0.405 (as computed with -0.235&pm;1.662*0.385). This further indicates that there is no difference in average BTUs between houses equipped with the electrically and the thermally activated vent (as 0 is in the confidence region).

### R Appendix
```{r eval=FALSE}
( pval <- 2*distrib(0.384,distrib="t",df=88,lower.tail=FALSE) )
( tstar <- distrib(0.05,distrib="t",df=88,lower.tail=FALSE,type="q"))
```

----

## Dioxins in Milk Cartons
```{r echo=FALSE, results="hide"}
mc <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/MilkCartons.csv")
tabs(name="2tMilk","Results of 2-Sample t-Test for dioxin levels in lined and unlined milk cartons.")
```
```{r 2tMilk, echo=FALSE, results="hide"}
levenesTest(dioxin~type,data=mc)
( mc.t <- t.test(dioxin~type,data=mc,var.equal=TRUE,alt="less",conf.level=0.99) )
```

1. &alpha; = 0.01.
1. The H<sub>0</sub>:&mu;<sub>L</sub>-&mu;<sub>UL</sub>=0 where &mu; is the mean dioxin level, L represents lined, and UL represents unlined cartons (thus, negative numbers represent lower dioxin levels in the lined cartons). The H<sub>A</sub>:&mu;<sub>L</sub>-&mu;<sub>UL</sub><0 (representing that the lined cartons will reduce the dioxin level).
1. A two-sample t-test is required because (i) a quantitative variable (dioxin level) was measured, (ii) two groups/populations (cartons with and without foil) exist, and (iii) those populations were independent (there is no connection between all of the cartons in the two groups).
1. The data appear to be part of an observational study; however there is no mention of randomization in this study although it would have been easy to randomly select cartons.
1. The individuals in the two groups are independent (cartons could not be both linead and unlined). The sample size (n<sub>L</sub>+n<sub>UL</sub>=`r nrow(mc)`) is &ge; 40. The variances appear to be UNequal because the Levene's test p-value (`r kPvalue(levenesTest(dioxin~type,data=mc)[1,"Pr(>F)"],include.p=FALSE,latex=FALSE)`) is less than &alpha;. Even though the assumptions are not fully met in this case, I am going to continue with the analysis and assume that the test statistic computed below should reasonably follow a t-distribution with n<sub>L</sub>+n<sub>UL</sub>-2 = `r nrow(mc) -2` df.
1. The statistic is x&#772;<sub>L</sub>-x&#772;<sub>UL</sub> = `r formatC(mc.t$estimate[1],format="f",digits=3)` - `r formatC(mc.t$estimate[2],format="f",digits=3)` = `r formatC(mc.t$estimate[1]-mc.t$estimate[2],format="f",digits=3)` (`r tabs("2tMilk",display="cite")`).
1. The test statistic is t = `r formatC(mc.t$statistic,format="f",digits=3)` with `r formatC(mc.t$parameter,format="f",digits=0)` df (`r tabs("2tMilk",display="cite")`).
1. The p-value is `r kPvalue(mc.t$p.value,include.p=FALSE,latex=FALSE)` (`r tabs("2tMilk",display="cite")`).
1. The H<sub>0</sub> is rejected because the p-value < &alpha;.
1. The lined milk cartons have lower levels of dioxin, on average, than the unlined cartons.
1. One is 99% confident that the mean dioxin levels is more than `r formatC(-1*mc.t$conf.int[2],format="f",digits=2)` GREATER in the unlined than in the lined cartons (`r tabs("2tMilk",display="cite")`).

`r tabs("2tMilk")`

```{r echo=FALSE, background='white', comment=""}
kHtest(mc.t)
```

### R Appendix
```{r eval=FALSE}
mc <- read.csv("data/MilkCartons.csv")
<<2tMilk>>
```

----

## Low Grades in Math
```{r echo=FALSE, results="hide"}
mg <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/UNCgrades.csv")
tabs(name="2tGrades","Results of 2-Sample t-Test for gpa in math and other classes at the University of North Carolina.")
```
```{r 2tMathGrades, echo=FALSE, results="hide"}
levenesTest(gpa~class.type,data=mg)
( mg.t <- t.test(gpa~class.type,data=mg,var.equal=TRUE,alt="less",conf.level=0.95) )
```

1. &alpha; = 0.05.
1. The H<sub>0</sub>:&mu;<sub>m</sub>-&mu;<sub>o</sub>=0 where &mu; is the mean gpa, m represents math class, and o represents other classes (thus, negative numbers represent lower grades in math classes). The H<sub>A</sub>:&mu;<sub>m</sub>-&mu;<sub>o</sub><0 (representing that the mean gpa is lower in the math classes).
1. A two-sample t-test is required because (i) a quantitative variable (gpa) was measured, (ii) two groups/populations (math and other class) exist, and (iii) those populations were independent (there is no connection between all of the students in the two types of clases).
1. The data appear to be part of an observational study, probably without randomization as the students were not "forced" to take the certain classes.
1. The individuals in the two groups are independent (no classes consisted of all the same students). The sample size (n<sub>m</sub>+n<sub>o</sub> = `r nrow(mg)`) is &ge; 40.  The population variances appear to be equal because the Levene's test p-value (`r kPvalue(levenesTest(gpa~class.type,data=mg)[1,"Pr(>F)"],include.p=FALSE,latex=FALSE)`) is larger than &alpha;. Therefore, the test statistic computed below should reasonably follow a t-distribution with n<sub>m</sub>+n<sub>o</sub>-2 = `r nrow(mg)-2` df.
1. The statistic is x&#772;<sub>m</sub>-x&#772;<sub>o</sub> = `r formatC(mg.t$estimate[1],format="f",digits=3)` - `r formatC(mg.t$estimate[1],format="f",digits=3)` = `r formatC(mg.t$estimate[1]-mg.t$estimate[2],format="f",digits=3)` (`r tabs("2tMathGrades",display="cite")`).
1. The test statistic is t = `r formatC(mg.t$statistic,format="f",digits=3)` with `r formatC(mg.t$parameter,format="f",digits=0)` df (`r tabs("2tMathGrades",display="cite")`).
1. The p-value is `r kPvalue(mg.t$p.value,include.p=FALSE,latex=FALSE)` (`r tabs("2tMathGrades",display="cite")`).
1. The H<sub>0</sub> is rejected because the p-value < &alpha;.
1. The mean gpa is lower in math classes than it is in the other classes.
1. One is 95% confident that the mean gpa of students in the math class is more than `r formatC(-1*mg.t$conf.int[2],format="f",digits=3)` LOWER than the gpa of students in the other classes (`r tabs("2tMathGrades",display="cite")`).

`r tabs("2tGrades")`

```{r echo=FALSE, background='white', comment=""}
kHtest(mg.t)
```

### R Appendix
```{r eval=FALSE}
mg <- read.csv("data/UNCGrades.csv")
<<2tMathGrades>>
```

----

## Hours Worked by Medical Interns
```{r echo=FALSE, results="hide"}
med <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/MedInternHrs.csv")
tabs(name="2tMedInterns","Results of 2-Sample t-Test for numbers of hours worked by medical interns in two cities.")
figs(name="2tMedInterns","Histogram of numbers of hours worked by medical interns in two cities.")
```
```{r 2tMedInterns, echo=FALSE, results="hide"}
levenesTest(hrs.worked~city,data=med)
( med.t <- t.test(hrs.worked~city,data=med,var.equal=TRUE,conf.level=0.90) )
```

1. &alpha; = 0.10.
1. The H<sub>0</sub>:&mu;<sub>1</sub>-&mu;<sub>2</sub>=0 where &mu; is the mean number of hours worked, 1 represents the first city, and 2 represents the second city (thus, positive numbers represent a larger mean for the first city). The H<sub>A</sub>:&mu;<sub>1</sub>-&mu;<sub>2</sub>&ne;0.
1. A two-sample t-test is required because (i) a quantitative variable (hours worked) was measured, (ii) two groups/populations (two cities) exist, and (iii) those populations were independent (there is no connection between all of the interns in the two cities).
1. The data appear to be part of an observational study with the individuals randomly selected.
1. The individuals in the two groups are independent (all interns were not in both cities). The sample size (n<sub>1</sub>+n<sub>2</sub>=`r nrow(med)`) is < 40 but &ge; 15. Histograms constructed for both groups (`r figs("2tMedInterns",display="cite")`) suggested a right-skewness but not an overly strong skew (this is very difficult to ascertain because of the small sample size). The population variances appear to be equal because the Levene's test p-value (`r kPvalue(levenesTest(hrs.worked~city,data=med)[1,"Pr(>F)"],include.p=FALSE,latex=FALSE)`) is larger than &alpha;. Therefore, the test statistic computed below should reasonably follow a t-distribution with n<sub>1</sub>+n<sub>2</sub>-2 = `r nrow(med)`-2 = `r nrow(med)-2` df.
1. The statistic is x&#772;<sub>1</sub>-x&#772;<sub>2</sub> = `r formatC(med.t$estimate[1],format="f",digits=2)` - `r formatC(med.t$estimate[2],format="f",digits=2)` = `r formatC(med.t$estimate[1]-med.t$estimate[2],format="f",digits=2)` (`r tabs("2tMedInterns",display="cite")`).
1. The test statistic is t =`r formatC(med.t$statistic,format="f",digits=3)` with `r formatC(med.t$parameter,format="f",digits=0)` df (`r tabs("2tMedInterns",display="cite")`).
1. The p-value is `r kPvalue(med.t$p.value,include.p=FALSE,latex=FALSE)` (`r tabs("2tMedInterns",display="cite")`).
1. The H<sub>0</sub> is rejected because the p-value < &alpha;.
1. It appears that interns in the second city work more hours, on average, than interns in the first city.
1. One is 90% confident that the interns in the second city work an average of between `r formatC(-1*med.t$conf.int[2],format="f",digits=2)` and `r formatC(-1*med.t$conf.int[1],format="f",digits=2)` hours more than interns in the first city (`r tabs("2tMedInterns",display="cite")`).

```{r 2tMedInternsHist, echo=FALSE, fig.width=7}
hist(hrs.worked~city,data=med,xlab="Hours Worked")
```

`r figs("2tMedInterns")`

`r tabs("2tMedInterns")`

```{r echo=FALSE, background='white', comment=""}
kHtest(med.t)
```

### R Appendix
```{r eval=FALSE}
med <- read.csv("data/MedInternHrs.csv")
<<2tMedInterns>>
<<2tMedInternsHist>>  
```

----

## Corn Crop Yield
```{r echo=FALSE, results="hide"}
yld <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/CropYield.csv")
tabs(name="2tYield","Results of 2-Sample t-Test for yields of corn between fields treated with an insecticide and sterile males.")
```
```{r 2tYield, echo=FALSE, results="hide"}
levenesTest(yield~group,data=yld)
( yld.t <- t.test(yield~group,data=yld,var.equal=TRUE,conf.level=0.95) )
```

1. &alpha; = 0.05.
1. The H<sub>0</sub>:&mu;<sub>i</sub>-&mu;<sub>s</sub>=0 where &mu; represents the mean yield, i represents the insecticide plots, and s represents the sterile male plots (thus, positive numbers represent a larger yield in the insecticide treatments). The H<sub>A</sub>:&mu;<sub>i</sub>-&mu;<sub>s</sub>&ne;0.
1. A two-sample t-test is required because (i) a quantitative variable (yield) was measured, (ii) two groups/populations (insecticide and sterile male plots) exist, and (iii) those populations were independent (there is no connection between plots that received the insecticide and those that received the sterile males).
1. The data appear to be part of an experimental study with two treatments and 40 replicates. Plots were randomly allocated to treatments.
1. The individuals in the two groups are independent as there is no connection between the two types of plots. The sample size (n<sub>i</sub>+n<sub>s</sub>=`r nrow(yld)`) is &ge; 40. The variances appear to be equal because the Levene's test p-value (`r kPvalue(levenesTest(yield~group,data=yld)[1,"Pr(>F)"],include.p=FALSE,latex=FALSE)`) is larger than &alpha;. Therefore, the test statistic computed below should reasonably follow a t-distribution with n<sub>i</sub>+n<sub>s</sub>-2 = `r nrow(yld)`-2 = `r nrow(yld)-2` df.
1. The statistic is x&#772;<sub>i</sub>-x&#772;<sub>s</sub> = `r formatC(yld.t$estimate[1],format="f",digits=3)`-`r formatC(yld.t$estimate[1],format="f",digits=3)` = `r formatC(yld.t$estimate[1]-yld.t$estimate[2],format="f",digits=3)` (`r tabs("2tYield",display="cite")`).
1. The test statistic is t = `r formatC(yld.t$statistic,format="f",digits=3)` with `r formatC(yld.t$parameter,format="f",digits=0)` df (`r tabs("2tYield",display="cite")`).
1. The p-value is `r kPvalue(yld.t$p.value,include.p=FALSE,latex=FALSE)` (`r tabs("2tYield",display="cite")`).
 1. The H<sub>0</sub> is rejected because the p-value < &alpha;.
1. It appears that yield is greater, on average, in the plots with the sterile males as compared to the plots with the insecticide.
1. One is 95% confident that the mean yield in plots with sterile males is between `r formatC(-1*yld.t$conf.int[2],format="f",digits=2)` and `r formatC(-1*yld.t$conf.int[1],format="f",digits=2)` bushels HIGHER than in plots with the insecticide (`r tabs("2tYield",display="cite")`).

`r tabs("2tYield")`

```{r echo=FALSE, background='white', comment=""}
kHtest(yld.t)
```

### R Appendix
```{r eval=FALSE}
yld <- read.csv("data/CropYield.csv")
<<2tYield>>
```

----

## Death Anxiety and Organ Donation
```{r echo=FALSE, results="hide"}
das <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/DeathAnxiety.csv")
tabs(name="2tDAS","Results of 2-Sample t-Test for death anxiety score between organ and nonorgan donors.")
```
```{r 2tDAS, echo=FALSE, results="hide"}
das$donor1 <- factor(das$donor,levels=c("organ","non.organ"))
levenesTest(DAS~donor1,data=das)
( das.t <- t.test(DAS~donor1,data=das,var.equal=TRUE,conf.level=0.99) )
```

1. &alpha; = 0.01.
1. The H<sub>0</sub>:&mu;<sub>o</sub>-&mu;<sub>n</sub>=0 where &mu; is the mean DAS, o represents the organ donors, and n represents the non-organ donors (thus, positive numbers represent a higher anxiety among organ donors). The H<sub>A</sub>:&mu;<sub>o</sub>-&mu;<sub>n</sub>&ne;0.
1. A two-sample t-test is required because (i) a quantitative variable (DAS) was measured, (ii) two groups/populations (donor or not) exist, and (iii) those populations were independent (there is no connection between all individuals in the two groups).
1. The data appear to be part of an observational study where it is not clear if random samples were taken or not.
1. The individuals in the two groups are independent because the individual cannot both be a donor and a non donor. The sample size (n<sub>o</sub>+n<sub>n</sub> = `r nrow(das)`) is &ge; 40. Therefore, the test statistic computed below should reasonably follow a t-distribution with n<sub>o</sub>+n<sub>n</sub>-2 = `r nrow(das)`-2 = `r nrow(das)-2` df. The variances appear to be equal because the Levene's test p-value (`r kPvalue(levenesTest(DAS~donor,data=das)[1,"Pr(>F)"],include.p=FALSE,latex=FALSE)`) is larger than &alpha;.
1. The statistic is x&#772;<sub>o</sub>-x&#772;<sub>n</sub> = `r formatC(das.t$estimate[1],format="f",digits=2)`-`r formatC(das.t$estimate[2],format="f",digits=2)` = `r formatC(das.t$estimate[1]-das.t$estimate[2],format="f",digits=2)` (`r tabs("2tDAS",display="cite")`).
1. The test statistic is t = `r formatC(das.t$statistic,format="f",digits=3)` with `r formatC(das.t$parameter,format="f",digits=0)` df (`r tabs("2tDAS",display="cite")`).
1. The p-value is `r kPvalue(das.t$p.value,include.p=FALSE,latex=FALSE)` (`r tabs("2tDAS",display="cite")`).
1. The H<sub>0</sub> is not rejected because the p-value > &alpha;.
1. There does not appear to be a significant difference in average anxiety about death between organ donors and non-organ donors, at least at the 1% level.
1. One is 99% confident that the difference in mean death anxiety score between organ and non-organ donors is between `r formatC(das.t$conf.int[1],format="f",digits=2)` and `r formatC(das.t$conf.int[2],format="f",digits=2)`. This further illustrates that there is no difference in mean death anxiety score between organ and non-organ donors (because 0 is contained in the confidence region)

`r tabs("2tDAS")`

```{r echo=FALSE, background='white', comment=""}
kHtest(das.t)
```

### R Appendix
```{r eval=FALSE}
das <- read.csv("dta/DeathAnxiety.csv")
<<2tDAS>>
```

----

## Pork Quality

1. &alpha; = 0.05.
1. The H<sub>0</sub>:&mu;<sub>1</sub>-&mu;<sub>2</sub>=0 where &mu; is the mean backfat thickness and the subscripts represent the two diets (thus, positive numbers represent more backfat thickness for pigs fed diet 1). The H<sub>A</sub>:&mu;<sub>1</sub>-&mu;<sub>2</sub>&ne;0.
1. A two-sample t-test is required because (i) a quantitative variable (backfat thickness) was measured, (ii) two groups/populations (different diets) exist, and (iii) those populations were independent (there is no connection between all pigs in the two diet groups).
1. The data are part of an experimental study where the pigs were randomly allocated to treatments.
1. The individuals in the two groups are independent (no pig was fed both diets). The sample size (n<sub>1</sub>+n<sub>2</sub>=25) is < 40 but &ge; 15 and the sample distributions (i.e., histograms) are not strongly skewed. The population variances appear to be equal because the p-value for Levene's test of the homogeneity of variance test (given as 0.532) is "large" (i.e., > &alpha;). Therefore, the test statistic computed below should reasonably follow a t-distribution with n<sub>1</sub>+n<sub>2</sub>-2 = 22 df.
1. The statistic is x&#772;<sub>1</sub>-x&#772;<sub>2</sub> = 3.420-2.989 = 0.431. The pooled sample variance is $s_{p}^{2}=\frac{(12-1)0.295^{2}+(12-1)0.375^{2}}{12+12-2}$ = 0.1138. The standard error of the statistic is $SE_{\bar{x}_{1}-\bar{x}_{2}}=\sqrt{0.1138\left(\frac{1}{12}+\frac{1}{12} \right)}$ = 0.1377.
1. The test statistic is t = $\frac{0.431-0}{0.1377}$ = 3.13 with 22 df.
1. The p-value is `r kPvalue(2*pt(3.13,df=22,lower.tail=FALSE),include.p=FALSE,latex=FALSE)`.
1. The H<sub>0</sub> is rejected because the p-value < &alpha;.
1. There does appear to be a difference in average backfat thickness between pigs fed the two diets, with pigs fed diet 2 having a lower mean backfat thickness and, thus, a higher-quality pork.
1. One is 95% confident that the mean backfat thickness on all pigs fed diet 1 is between 0.145 and 0.717 cm greater than the mean backfat thickness of all pigs fed diet 2 (as computed with 0.431&pm;2.074*0.138). Thus, diet 2 produces higher-quality pork.

### R Appendix
```{r eval=FALSE}
( pval <- 2*distrib(3.13,distrib="t",df=22,lower.tail=FALSE) )
( tstar <- distrib(0.025,distrib="t",type="q",df=22,lower.tail=FALSE) )
```
