---
title: "Review Exercise Key"
subtitle: "Confidence Regions"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/RE/zRevExAns/")
source("../../../rhelpers/rhelpers.R")
modHTML("ConfRegions")
```
```{r echo=FALSE}
library(NCStats)
```

----

## BOD in Effluent
```{r CIbod, echo=FALSE, results='hide', fig.show='hide'}
zstar <- distrib(0.90,type="q",lower.tail=FALSE)
```

1. 90% (=100(1-0.10))
1. Lower confidence bound (because $H_{A}$ is "greater than").
1. $Z^{*}$=`r formatC(zstar,format="f",digits=3)`
1. 2504`r formatC(zstar,format="f",digits=3)`$\frac{1200}{\sqrt(118)}$ = `r formatC(2504+zstar*1200/sqrt(118),format="f",digits=0)`.
1. I am 90% confident that the population mean BOD is greater than `r formatC(2504+zstar*1200/sqrt(118),format="f",digits=0)`. By confident I mean that 90% of all 90% confidence regions will contain the true population mean.

### R Appendix
```{r eval=FALSE}
library(NCStats)
<<CIbod>>
```

----

## Medical School Admissions
```{r CImed, echo=FALSE, results='hide', fig.show='hide'}
zstar <- distrib(0.95,type="q",lower.tail=FALSE)
```

1. 95% (=100(1-0.05))
1. Upper confidence bound (because $H_{A}$ is "less than").
1. $Z^{*}$=`r formatC(zstar,format="f",digits=3)`
1. 3.60+`r formatC(zstar,format="f",digits=3)`$\frac{0.35}{\sqrt(40)}$ = `r formatC(3.6+zstar*0.35/sqrt(40),format="f",digits=2)`.
1. I am 95% confident that the population mean GPA is less than `r formatC(3.6+zstar*0.35/sqrt(40),format="f",digits=2)`. By confident I mean that 95% of all 95% confidence regions will contain the true population mean.

### R Appendix
```{r eval=FALSE}
library(NCStats)
<<CImed>>
```

----

## Banff Snow Depth
```{r echo=FALSE}
banff <- data.frame(snow=c(29.00,45.51,30.18,45.83,39.54,80.39,32.64,32.89,
                           46.84,45.79,62.92,67.24,30.96,46.08,33.28))
```
```{r CIbanff, echo=FALSE, results='hide', fig.show='hide'}
Summarize(~snow,data=banff,digits=2)
zstar <- distrib(0.995,type="q")
```

1. 99%
1. Confidence interval.
1. $Z^{*}$=`r formatC(zstar,format="f",digits=3)`
1. `r formatC(mean(banff$snow),format="f",digits=2)` $\pm$ `r formatC(zstar,format="f",digits=3)` $\frac{15}{\sqrt(15)}$ = (`r formatC(mean(banff$snow)-zstar*15/sqrt(15),format="f",digits=2)`,`r formatC(mean(banff$snow)+zstar*15/sqrt(15),format="f",digits=2)`).
1. I am 95% confident that the population mean snowpack depth is between `r formatC(mean(banff$snow)-zstar*15/sqrt(15),format="f",digits=2)` and `r formatC(mean(banff$snow)+zstar*15/sqrt(15),format="f",digits=2)` cm. By confident I mean that 99% of all 99% confidence regions will contain the true population mean.

### R Appendix
```{r eval=FALSE}
library(NCStats)
banff <- read.csv("data/banff.csv")
<<CIbanff>>
```

----

## Creatine Phosphate Concentrations
```{r echo=FALSE}
cp <- data.frame(conc=c(121,82,100,151,68,58,95,145,64,
                        119,104,110,113,118,203,62,83,67,
                        201,101,163,84,57,139,60,78,94,
                        93,92,110,25,123,70,48,95,42))
```
```{r CIcp, echo=FALSE, results='hide', fig.show='hide'}
Summarize(~conc,data=cp,digits=1)
zstar <- distrib(0.95,type="q",lower.tail=FALSE)
```

1. 95% (=100(1-0.05)).
1. Lower confidence bound (because $H_{A}$ is a "greater than").
1. $Z^{*}$=`r formatC(zstar,format="f",digits=3)`.
1. `r formatC(mean(cp$conc),format="f",digits=1)``r formatC(zstar,format="f",digits=3)`$\frac{40}{\sqrt(36)}$ = `r formatC(mean(cp$conc)+zstar*40/sqrt(36),format="f",digits=1)`.
1. I am 95% confident that the population mean creatine phosphate concentration is greater than `r formatC(mean(cp$conc)+zstar*40/sqrt(36),format="f",digits=1)`. By confident I mean that 95% of all 95% confidence regions will contain the true population mean.

### R Appendix
```{r eval=FALSE}
library(NCStats)
cp <- read.csv("data/creatine.csv")
<<CIcp>>
```

----

## Brule River Gauge Heights
```{r echo=FALSE}
br <- data.frame(hts=c(1.56,1.54,1.54,1.57,1.58,1.61,1.60,1.69,
                       1.99,2.11,1.98,1.76,1.69,1.99,1.86,1.53))
```
```{r CIbr, echo=FALSE, results='hide', fig.show='hide'}
Summarize(~hts,data=br,digits=2)
zstar <- distrib(0.975,type="q")
```

1. 95% (=100(1-0.05)).
1. Confidence interval (because $H_{A}$ is a "not equals").
1. $Z^{*}$= $\pm$ `r formatC(zstar,format="f",digits=3)`.
1. `r formatC(mean(br$hts),format="f",digits=2)` $\pm$ `r formatC(zstar,format="f",digits=3)`$\frac{0.2}{\sqrt(16)}$ = (`r formatC(mean(br$hts)-zstar*0.2/sqrt(16),format="f",digits=2)`,`r formatC(mean(br$hts)+zstar*0.2/sqrt(16),format="f",digits=2)`).
1. I am 95% confident that the population mean gauge height is between `r formatC(mean(br$hts)-zstar*0.2/sqrt(16),format="f",digits=2)` and `r formatC(mean(br$hts)+zstar*0.2/sqrt(16),format="f",digits=2)` feet. By confident I mean that 95% of all 95% confidence regions will contain the true population mean.

### R Appendix
```{r eval=FALSE}
library(NCStats)
br <- read.csv("data/bruleriver.csv")
<<CIbr>>
```

----

## Population Density in Wisconsin Counties
```{r echo=FALSE}
wic <- data.frame(dens=c(429.0,67.8,52.1,97.4,57.9,354.9,16.2,19.1,
                         127.0,27.6,10.2,54.6,28.8,30.1,20.2))
```
```{r CIwic, echo=FALSE, results='hide', fig.show='hide'}
Summarize(~dens,data=wic,digits=2)
zstar <- distrib(0.90,type="q")
```

1. 90% (=100(1-0.10)).
1. Upper confidence bound (because $H_{A}$ is a "less than").
1. $Z^{*}$=`r formatC(zstar,format="f",digits=3)`.
1. `r formatC(mean(wic$dens),format="f",digits=1)`+`r formatC(zstar,format="f",digits=3)`$\frac{125}{\sqrt(15)}$ = `r formatC(mean(wic$dens)+zstar*125/sqrt(15),format="f",digits=1)`.
1. I am 90% confident that the population mean density of people per county in Wisconsin is less than `r formatC(mean(wic$dens)+zstar*125/sqrt(15),format="f",digits=1)`. By confident I mean that 90% of all 90% confidence regions will contain the true population mean.

### R Appendix
```{r eval=FALSE}
library(NCStats)
cp <- read.csv("data/creatine.csv")
<<CIcp>>
```

----

## Pebble Size
```{r npebble, echo=FALSE, results='hide', fig.show='hide'}
zstar <- distrib(0.995,type="q")
me <- 0.1
sigma <- 4
( n <- (zstar*sigma/me)^2 )
```

The required sample size is `r formatC(ceiling(n),format="f",digits=0,big.mark=",")` pebbles.

### R Appendix
```{r eval=FALSE}
library(NCStats)
<<npebble>>
```

----

## Internet Usage
```{r nisp, echo=FALSE, results='hide', fig.show='hide'}
zstar <- distrib(0.975,type="q")
me <- 1
sigma <- 6.95
( n <- (zstar*sigma/me)^2 )
```

The required sample size is `r ceiling(n)` households.

### R Appendix
```{r eval=FALSE}
library(NCStats)
<<nisp>>
```

----

## Counting Plants in Plots


\vspace{24pt}
\section*{Counting Plants in Plots}
```{r nplants, echo=FALSE,results="hide", fig.show="hide"}
me <- 10
sd <- 50
( zstar <- distrib(0.95,type="q") )
( n <- (zstar*sd/me)^2 )
```

A total of `r formatC(ceiling(n),format="f",digits=0)` plots would be required.

### R Appendix
```{r eval=FALSE}
library(NCStats)
<<nplants>>
```
