---
title: "Review Exercise Key"
subtitle: "Hypothesis Testing"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/RE/zRevExAns/")
source("../../../rhelpers/rhelpers.R")
modHTML("HypTesting")
```
```{r echo=FALSE, message=FALSE, results='hide'}
library(NCStats)
```

## Computing P-Values 1

The p-value is `r kPvalue(distrib(12.1,mean=10,sd=5/sqrt(25),lower.tail=FALSE,plot=FALSE))`, reject $H_{O}$.
```{r fig.show='hide', results='hide'}
( distrib(12.1,mean=10,sd=5/sqrt(25),lower.tail=FALSE) )
```

----

## Computing P-Values 2

The p-value is `r kPvalue(distrib(43.8,mean=50,sd=20/sqrt(50),lower.tail=TRUE,plot=FALSE))`, reject $H_{O}$.
```{r fig.show='hide', results='hide'}
( distrib(43.8,mean=50,sd=20/sqrt(50),lower.tail=TRUE) )
```

----

## Computing P-Values 3

The p-value is `r kPvalue(2*distrib(98,mean=100,sd=15/sqrt(100),lower.tail=TRUE,plot=FALSE))`, DNR $H_{O}$.
```{r fig.show='hide', results='hide'}
( 2*distrib(98,mean=100,sd=15/sqrt(100),lower.tail=TRUE) )
```

----

## BOD in Effluent

```{r pvalBOD, echo=FALSE, results='hide', fig.show='hide'}
( pval <- distrib(2504,mean=2200,sd=1200/sqrt(118),lower.tail=FALSE) )
```

1. $H_{O}:\mu=2200$ vs. $H_{O}:\mu>2200$.
1. $z=\frac{2504-2200}{\frac{1200}{\sqrt{118}}}=2.75$.
1. The p-value is \Sexpr{kPvalue(pval)}.
1. The p-value is less than $\alpha$, therefore $H_{O}$ is rejected.
1. There is evidence that the mean BOD level is greater than 2200 lbs/day.  Thus, the plant managers will need to make corrective actions.

### R Appendix
```{r eval=FALSE}
<<pvalBOD>>
```

----

## Medical School Admissions
```{r pvalMED, echo=FALSE, results='hide', fig.show='hide'}
( pval <- distrib(3.6,mean=3.7,sd=0.35/sqrt(40)) )
```

1. $H_{O}:\mu=3.7$ vs. $H_{A}:\mu<3.7$.
1. $z=\frac{3.6-3.7}{\frac{0.35}{\sqrt{40}}}=-1.807016$.
1. The p-value is \Sexpr{kPvalue(pval)}.
1. The p-value is less than $\alpha$, therefore $H_{O}$ is rejected.
1. It appears that the mean gpa in non-science courses is less than 3.7.  Thus, the administrator's concern was warranted.

### R Appendix
```{r eval=FALSE}
<<pvalMED>>
```

----

## Power Calculations 1
```{r PowerCalc1, echo=FALSE, results='hide', fig.show='hide'}
( rejreg <- distrib(0.05,mean=50,sd=20/sqrt(50),type="q") )
( pwr <- distrib(rejreg,mean=45,sd=20/sqrt(50)) )
```

The statistical power for this situation is `r formatC(pwr,format="f",digits=4)`.

### R Appendix
```{r eval=FALSE}
<<PowerCalc1>>
```

----

## Power Calculations 2
```{r PowerCalc2, echo=FALSE, results='hide', fig.show='hide'}
( rejreg <- distrib(0.10,mean=10,sd=5/sqrt(25),type="q",lower.tail=FALSE) )
( pwr <- distrib(rejreg,mean=12,sd=5/sqrt(25),lower.tail=FALSE) )
```

The statistical power for this situation is `r formatC(pwr,format="f",digits=4)`.

### R Appendix
```{r eval=FALSE}
<<PowerCalc2>>
```

----

## Power Calculations 3
```{r PowerCalc3, echo=FALSE, results='hide', fig.show='hide'}
( rejreg <- distrib(0.01,mean=75,sd=15/sqrt(30),type="q",lower.tail=FALSE) )
( pwr <- distrib(rejreg,mean=82,sd=15/sqrt(30),lower.tail=FALSE) )
```

The statistical power for this situation is `r formatC(pwr,format="f",digits=4)`.

### R Appendix
```{r eval=FALSE}
<<PowerCalc3>>
```

----
