---
title: "Review Exercise Key"
subtitle: "Probability"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/RE/zRevExAns/")
source("../../../rhelpers/rhelpers.R")
modHTML("Probability")
```
```{r echo=FALSE, message=FALSE}
library(NCStats)
```

----

## Coin Purse

1. The probability of randomly selecting a nickel from this purse is `r formatC(17/32,format="f",digits=3)` (i.e., $\frac{17}{32}$).
1. The probability of randomly selecting a dime from this purse is `r formatC(15/32,format="f",digits=3)` (i.e., $\frac{15}{32}$).
1. The probability of randomly selecting a dime if two nickels and three dimes had been removed is `r formatC(12/27,format="f",digits=3)` (i.e., $\frac{12}{27}$).

----

## Greenhouse Plants

1. The probability of randomly selected a tomato plant is `r formatC(10/30,format="f",digits=3)` (i.e., $\frac{10}{30}$).
1. The probability of randomly selected a cauliflower plant is `r formatC(8/30,format="f",digits=3)` (i.e., $\frac{8}{30}$).
1. The probability of randomly selected a pea plant assuming that all tomato plants were gone is `r formatC(12/20,format="f",digits=3)` (i.e., $\frac{12}{20}$).

----

## Raccoons Consuming Corn
```{r NormRocky, echo=FALSE, include=FALSE}
( q2 <-  distrib(10,mean=8,sd=2/sqrt(25),lower.tail=FALSE) )
( q3 <- distrib(8.5,mean=8,sd=2/sqrt(60),lower.tail=FALSE) )

```

1. This question is about an individual and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of  $\bar{x}\sim N(8,\frac{2}{\sqrt{25}})$ because $n=25 \geq 15$ and the population distribution is not strongly skewed.  Thus, the probability is `r formatC(q2,digits=3,format="fg")`.
1. This question is ultimately about a statistic as the total of 510 should be converted an $\bar{x}$ of $\frac{510}{60}=8.5$ kg.  The sampling distribution of $\bar{x}\sim N(8,\frac{2}{\sqrt{60}})$ or $\bar{x}\sim N(8,0.2582)$ because $n=60 \geq 30$.  Thus, the probability is `r formatC(q3,digits=4,format="f")`.

### R Appendix
```{r eval=FALSE}
<<NormRocky>>
```

----

## NFL Running Backs 
```{r NormNFL, echo=FALSE, include=FALSE}
( q2 <- distrib(100,mean=82,sd=26/sqrt(16),lower.tail=FALSE) )
q3ab <- distrib(90,mean=82,sd=26/sqrt(16))
q3a <- distrib(70,mean=82,sd=26/sqrt(16))
q3ab-q3a
( q4 <- distrib(70,mean=82,sd=26/sqrt(32),lower.tail=FALSE) )
( q6 <- distrib(0.05,mean=82,sd=26/sqrt(16),type="q",lower.tail=FALSE) )

```

1. This question is about an individual (a game) and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(82,\frac{26}{\sqrt{16}})$ because $n=16 \geq 15$ and the population distribution is not strongly skewed.  Thus, the probability is `r formatC(q2,digits=4,format="f")`.
1. This question is about a statistic and requires the same sampling distribution as the previous question --  $\bar{x}\sim N(82,\frac{26}{\sqrt{16}})$..  Thus, the probability is `r formatC(q3ab-q3a,digits=4,format="f")`.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(82,\frac{26}{\sqrt{32}})$ because $n=32 \geq 30$.  Thus, the probability is `r formatC(q4,digits=4,format="f")`.
1. This question is about an individual (a game) and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(82,\frac{26}{\sqrt{16}})$ because $n=16 \geq 15$ and the population distribution is not strongly skewed.  Thus, the mean yards gained is `r formatC(q6,digits=1,format="f")`.

### R Appendix
```{r eval=FALSE}
<<NormNFL>>
```

----

## Stock Returns
```{r NormStocks, echo=FALSE, include=FALSE}
( q1 <- distrib(0,mean=4.2,sd=4.9/sqrt(5),lower.tail=FALSE) )
( q2 <- distrib(0,mean=4.2,sd=4.9,lower.tail=FALSE) )
( q3 <- distrib(2,mean=4.2,sd=4.9/sqrt(10)) )
( q4 <- distrib(0.10,mean=4.2,sd=4.9,type="q",lower.tail=FALSE) )
( q5 <- distrib(0.10,mean=4.2,sd=4.9/sqrt(10),type="q",lower.tail=FALSE) )
```

1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(4.2,\frac{4.9}{\sqrt{5}})$ because the population distribution is approximately normal.  Thus, the probability is `r formatC(q1,digits=4,format="f")`.
1. This question is about an individual (a stock) and requires the population distribution. Because the population distribution is approximately normally distributed, the probability can be computed.  Thus, the probability is `r formatC(q2,digits=4,format="f")`.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(4.2,\frac{4.9}{\sqrt{10}})$ because the population distribution is approximately normal.  Thus, the probability is `r formatC(q3,digits=4,format="f")`.
1. This question is about an individual (a stock) and requires the population distribution. Because the population distribution is approximately normally distributed, the probability can be computed.  Thus, the mean return is `r formatC(q4,digits=1,format="f")`.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(4.2,\frac{4.9}{\sqrt{10}})$ because the population distribution is approximately normal.  Thus, the mean return is `r formatC(q5,digits=1,format="f")`.

### R Appendix
```{r eval=FALSE}
<<NormStocks>>
```

----

## HMF in Honey
```{r NormHoney, echo=FALSE, include=FALSE}
( q3 <- distrib(10,mean=9.5,sd=13.5/sqrt(50)) )
( q4 <- distrib(0.20,mean=9.5,sd=13.5/sqrt(50),type="q") )
```

1. This question is about an individual (one kg of honey) and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution, however, is not normal because $n < 30$ and the population distribution is strongly skewed.  Thus, this question CANNOT BE COMPUTED.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(9.5,\frac{13.5}{\sqrt{50}})$ because $n \geq 30$.  Thus, the probability is `r formatC(q3,digits=4,format="f")`.
1. This question is about a statistic and requires the same sampling distribution as in the previous question -- $\bar{x}\sim N(9.5,\frac{13.5}{\sqrt{50}})$.  Thus, the average amount is `r formatC(distrib(0.20,mean=9.5,sd=13.5/sqrt(50),type="q",plot=FALSE),digits=1,format="f")`.

### R Appendix
```{r eval=FALSE}
<<NormHoney>>
```

----

## Size of Farms in England
```{r NormFarms, echo=FALSE, include=FALSE}
( q2a <- distrib(0.45,mean=65.13,sd=108.71/sqrt(60),type="q") )
( q2b <- distrib(0.55,mean=65.13,sd=108.71/sqrt(60),type="q") )
( q3 <- distrib(50,mean=65.13,sd=108.71/sqrt(60)) )
```

1. This question is about an individual (a farm) and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(65.13,\frac{108.71}{\sqrt{60}})$ because $n \geq 30$.  Thus, the most common 10\% of means is between `r formatC(q2a,digits=2,format="f")` and `r formatC(q2b,digits=2,format="f")`.
1. This question is about a statistic and requires the same sampling distribution as in the previous question -- $\bar{x}\sim N(65.13,\frac{108.71}{\sqrt{60}})$.  Thus, the probability is `r formatC(q3,digits=4,format="f")`.
1. This question is about an individual (a farm) and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.

### R Appendix
```{r eval=FALSE}
<<NormFarms>>
```

----

## Painted Turtle Hatchling Size
```{r NormTurtle, echo=FALSE, include=FALSE}
( q2 <- distrib(4.5,mean=4.46,sd=0.13/sqrt(20),lower.tail=FALSE) )
( q3 <- distrib(4.5,mean=4.46,sd=0.13/sqrt(50),lower.tail=FALSE) )
( q4 <- distrib(0.20,mean=4.46,sd=0.13/sqrt(50),type="q") )
```

1. This question is about an individual (a turtle) and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(4.46,\frac{0.13}{\sqrt{20}})$ because $n \geq 15$ and the population is only slightly skewed.  Thus, the probability that a sample of 20 turtles will have an average number of days until hatching that is greater than 4.5 days is `r formatC(q2,digits=4,format="f")`.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(4.46,\frac{0.13}{\sqrt{50}})$ because $n \geq 30$.  Thus, the probability that a sample of 50 turtles will have an average number of days until hatching that is greater than 4.5 days is `r formatC(q3,digits=4,format="f")`.
1. This question is about a statistic and requires the same sampling distribution as in the previous question -- $\bar{x}\sim N(4.46,\frac{0.13}{\sqrt{50}})$.  Thus, the mean number of days until hatching such that 20\% of samples of 50 turtles have a smaller mean is `r formatC(q4,digits=2,format="f")`.
1. This question is about an individual (a turtle) and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.

### R Appendix
```{r eval=FALSE}
<<NormTurtle>>
```

----

## Moose Hunting
```{r NormMoose, echo=FALSE, include=FALSE}
( q4 <- distrib(48,mean=40,sd=15/sqrt(25),lower.tail=FALSE) )
```

1. An individual is a Minnesota moose hunter.
1. Time spent hunting is a continuous quantitative variable.
1. This question is about a quantitative variable and an individual; thus, the population distribution would be used.  This question cannot be answered because the population distribution is NOT normal (it is only known that it is symmetric).
1. This question is about a quantitative variable and a statistic; thus, the sampling distribution of $\bar{x}$ would be used.  Because the population distribution is only symmetric, the sample size must be greater than 15, which it is in this question.  Thus, $\bar{x}\sim N(40,\frac{15}{\sqrt{25}})$ or $\bar{x}\sim N(40,3)$.  Thus, the probability is `r formatC(q4,digits=4,format="f")`.

### R Appendix
```{r eval=FALSE}
<<NormMoose>>
```

----

## Lengths of Pine Needles
```{r ProbLPN, echo=FALSE, results='hide', fig.show='hide'}
( q1ab <- distrib(80,mean=75,sd=8) )
( q1a <- distrib(70,mean=75,sd=8) )
q1ab-q1a
( q2 <- distrib(90,mean=75,sd=8,lower.tail=FALSE) )
( q3 <- distrib(50,mean=75,sd=8) )
```

1. The probability that a randomly selected needle is between 70 and 80 mm long is
`r formatC(q1ab-q1a,digits=3,format="f")`.
1. The probability that a randomly selected needle is longer than 90 mm is `r formatC(q2,digits=3,format="f")`.
1. The probability that a randomly selected needle is less than 50 mm is `r formatC(q3,digits=4,format="f")`.

### R Appendix
```{r fig.show='hide', results='hide'}
<<ProbLPN>>
```

----
