---
title: "Review Exercise Key"
subtitle: "Univariate EDA - Categorical"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/RE/zRevExAns/")
source("../../../rhelpers/rhelpers.R")
modHTML("UEDACat")
```
```{r echo=FALSE, results='hide'}
library(NCStats)
library(captioner)
tabs <- captioner(prefix="Table")

```

## Drinking Water Source

The approximate frequency table is shown below.
```{r echo=FALSE, comment=""}
d <- c("Groundwater"=15.9,"Lake/Reservoir"=29.2,"River"=9.6,"Muliple"=15.7,"Don't Know"=29.4,"Refused to Answer"=0.2)  
round(d/100*751,0)
```

A very large number of respondents did not know where their water supply was derived.  Of those that did, most were derived from a lake or reservoir, with about the same number evenly split between groundwater and multiple sources.

----

## Curbside Recycling

```{r echo=FALSE, results='hide'}
recycle <- factor(c("C","C","B","B","B","B","C","E","B","B",
                    "C","B","C","B","C","C","C","E","B","B",
                    "B","C","B","B","C","C","C","B","C","B",
                    "C","B","B","C","B","C","B","B","B","C",
                    "E","B","E","B","B","C","C","B","B","B",
                    "E","B","C","C","C","B","B","C","B","B","B","B","B"))
hi <- data.frame(recycle)
t.hi <- xtabs(~recycle,data=hi)
pt.hi <- percTable(t.hi,digits=1)
tabs(name="Recycle","Percentage table for how much curbside recycling has reduced overall refuse. Note that B=25%, C=50%, and E='too early to tell'.")
```

More than half of the respondents felt that the curbside recycling program has reduced their waste by approximately 25%, with approximately another third saying approximately 50% (`r tabs("Recycle",display="cite")`). The only other response offered was "too early to tell."

`r tabs("Recycle")`
```{r Recycle, echo=FALSE, comment=""}
pt.hi
```

### R Code Appendix
```
library(NCStats)
df <- read.csv("data/Recycling.csv")
t.hi <- xtabs(~recycle,data=df)
percTable(t.hi,digits=1)
```

----

## Zoo Animal Types
```{r echo=FALSE, results='hide'}
df <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/Zoo1.csv")
freq <- xtabs(~type,data=df)
perc <- percTable(freq,digits=1)
tabs(name="Zoos","Percentage table of animal types in a sample of several national zoos.")
```

Animals were nearly evenly split among the three categories, with somewhat more birds and somewhat fewer amphibian/reptiles (`r tabs("Water",display="cite")`).

`r tabs("Zoos")`
```{r Zoos, echo=FALSE, comment=""}
perc
```

### R Code Appendix
```
library(NCStats)
df <- read.csv("data/Zoo1.csv")
freq <- xtabs(~type,data=df)
percTable(freq,digits=1)
```

----
