---
layout: page
title: "Inference Concepts"
author: "Derek H. Ogle"
css: "/css/handouts.css"
output:
  html_document:
    fig_width: 3.5
    fig_height: 3.5
    self_contained: no
  pdf_document:
    fig_width: 3.5
    fig_height: 3.5
  word_document:
    fig_width: 3.5
    fig_height: 3.5
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/class/HOs")
source("../../../rhelpers/rhelpers.R")
fnm <- "InferenceConcepts"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message='FALSE'}
fnm <- "InferenceConcepts"
source("../../../rhelpers/knitr_setup.R")
```

## Background
Barrels designed to hold 200-gallons of gasoline were recently found in an abandoned warehouse.  The contents of a random sample of 38 barrels were carefully measured to determine if the barrels had leaked a significant amount of gasoline.   Assume that it is known that the actual content of the barrels has a standard deviation of 10 gallons.  The results for the sample are found in [barrels.csv](https://github.com/droglenc/NCData/blob/master/Barrels.csv) on the class webpage.  Use results computed from the sample to determine, at the 10% level, if there is evidence that the barrels had leaked.

## Getting the Data
```{r message=FALSE, warning=FALSE}
library(NCStats)
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/class/HOs")
brls <- read.csv("Barrels.csv")
str(brls)
```

## Quick EDA
```{r par1=TRUE}
Summarize(~gasoline,data=brls,digits=1)
hist(~gasoline,data=brls,xlab="Gallons of Gasoline")
```

## 1-Sample Z-test
```{r par1a=TRUE}
( gas.z <- z.test(brls$gasoline,sd=10,mu=200,alt="less",conf.level=0.90) )
plot(gas.z)
```

```{r echo=FALSE, results="hide", message=FALSE, warnings=FALSE}
purl2(paste0(fnm,".Rmd"))
```
