---
layout: page
title: Assignment
subtitle: 2-Sample t-test
author: Derek H. Ogle
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/CE/")
source("../../rhelpers/rhelpers.R")
modHTML("2Samplet_CE1")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../rhelpers/knitr_setup.R")
```

<div class="alert alert-success">For all "hypothesis testing" questions, make sure to provide explicit answers with your work shown for all <a href="../11-steps">11 steps of a hypothesis test</a>.
</div>

<br>

## Compost Type and Germination Rates
The owner of a greenhouse developed a methodology to produce sterilized compost for growing plants that was cheaper than the compost she could buy while also allowing her to reduce waste production. However, these benefits could be offset by losses related to poor plant germination and growth. Thus, the owner attempted to answer the queston about germination success by planting 50 seeds of [marigolds (*Tagetes* spp.)](https://en.wikipedia.org/wiki/Tagetes) into individual containers that were filled with either the home-made sterilized compost or store-bought compost. The containers were randomly filled with one type of compost and were randomly placed into a large greenhouse where all environmental conditions were controlled. After five days, the number of (the 50) seeds in each container that had germinated was recorded. The results are shown below.

```{r echo=FALSE, fig.width=6, fig.height=3.5}
tmp <- data.frame(compost=factor(rep(c("bought","homemade"),c(18,10))),
                  germination=c(36,43,44,39,44,50,39,44,46,39,42,50,42,46,
                                39,38,49,38,28,31,25,28,28,27,32,24,33,33))
Summarize(germination~compost,data=tmp,digits=2)
levenesTest(germination~compost,data=tmp)
par(mar=c(3.05,3.05,1.3,0.65),mgp=c(1.9,0.3,0),tcl=-0.2,las=1,cex.lab=0.95,cex.axis=0.9)
hist(germination~compost,data=tmp,xlab="Number Germinated",same.breaks=FALSE)
```

Use this information to determine, at the 1% level, if the number of seeds that germinated was lower in the home-made compost.[^1] [*NOTE: you are testing if the germination rate was lower in the home-made compost because that would suggest that it performs worse -- fewer germinated plants -- than the other compost and would increase costs to the greenhouse.*]

----

## Commitment to Adult Animals
A researcher constructed a survey to determine an individual person's "commitment to adult animals." Each survey is summarized to a single number that measures that individual's "commitment", with larger numbers representing "greater commitment." The researcher wanted to determine if the mean "commitment" according to this measure was greater for people who evacuated all or some of their pets versus those who did not evacuate any pets during a propane tanker derailment in Weyauwega, WI in 1996. The table below shows the results for the "commitment" measure for 116 individuals that evacuated all or some of their pets (i.e., `DidEvac`) and for 125 individuals that evacuated none of their pets (i.e., `NoEvac`).

<pre>
Variable   n  Mean Median StDev SE Mean   Min    Max    Q1    Q3
DidEvac  116 7.694  7.658 3.410  0.317  -0.863 14.763 5.035 10.204
NoEvac   125 6.640  6.599 3.102  0.277  -1.214 14.444 4.568  8.696

Levene's p=0.678
</pre>

Use these results to examine the researcher's hypothesis at the 1% significance level.

&nbsp;

[^1]: This question was adapted from [McClave *et al.* (1998) *First Course in Business Statistics*](http://www.amazon.com/First-Course-Business-Statistics-TEXT/dp/B000MLT2MY/).
