---
layout: page
title: Class Example
subtitle: 2-Sample t-test
author: Derek H. Ogle
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
    lib_dir: zlibs

---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/CE/")
source("../../rhelpers/rhelpers.R")
modHTML("2Samplet_CExmpl")
```

```{r echo=FALSE, results='hide', message=FALSE}
source("../../rhelpers/knitr_setup.R")
library(NCStats)
library(ggplot2)
library(captioner)
tbls <- captioner(prefix="Table")
tbls(name="ldl","Summary statistics for the reduction in LDL for subjects that received Befizal and Lipanthyl.")
figs <- captioner(prefix="Table")
figs(name="ldl","Histogram for the reduction in LDL for subjects that received Befizal and Lipanthyl.")
```

## Cholesterol Drug Treatments I
[Low density lipoprotein (LDL)](https://en.wikipedia.org/wiki/Low-density_lipoprotein) is often called "bad cholesterol" because they can cause lipids to be deposited in artery walls, thus causing arteries to become blocked.  Levels of LDL may be lowered with lifestyle changes and pharmaceuticals.  Researchers conducted an experiment where 25 men with high cholesterol were randomly assigned one of two drugs (Befizal and Lipanthyl) designed to lower LDL levels.  The results from this test are shown below.


```{r echo=FALSE}
d <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/Cholesterol2.csv")
d$redux <- d$before-d$after
Summarize(redux~tx,data=d,digits=3)
```
```{r warning=FALSE}
levenesTest(redux~tx,data=d)
```
```{r echo=FALSE, fig.width=7}
ggplot(data=d,mapping=aes(x=redux)) +
  geom_histogram(binwidth=0.1,boundary=0,color="black",fill="lightgray") +
  labs(x="Reduction in LDL",y="Frequency of Subjects") +
  scale_y_continuous(expand=expansion(mult=c(0,0.05))) +
  theme_NCStats() +
  facet_wrap(vars(tx))
```

Determine, at the 5% level, if there is a difference in the mean decrease in LDL between the two drugs.  Make sure to provide explicit answers with your work shown for all [11 steps of a hypothesis test](../11-steps.html).
