---
layout: page
title: Assignment
subtitle: Hypothesis Testing - Errors
author: Derek H. Ogle
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/CE/")
source("../../rhelpers/rhelpers.R")
modHTML("HypTesting2_CE1")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
source("../../rhelpers/knitr_setup.R")
library(NCStats)
library(captioner)
source("../../rhelpers/knitr_setup.R")
tbls <- captioner(prefix="Table")
tbls(name="bodytemp","Summary statistics for the body temperature of a sample of men and women.")
tbls(name="beetles","Summary statistics for the thorax length for two species of beetles.")
```

## Body Temperature
[Machowiak *et al.* (1992)](http://www.ncbi.nlm.nih.gov/pubmed/1302471)[^1] provided a critical examination of whether normal body temperature was 37<sup>o</sup>C. They recorded the orally-determined body temperatures of 65 male and 65 female subjects. Use their results in `r tbls("bodytemp",display="cite")` to answer the questions further below.

`r tbls("bodytemp")`

```{r echo=FALSE, background="white"}
d <- read.csv("https://github.com/droglenc/NCData/raw/master/NormTemp.csv")
Summarize(~bodytemp,data=d,digits=2)
```

1. construct statistical hypotheses (*make sure to define* &mu;),
1. write the definition of the p-value specific to this question,
1. compute the p-value assuming that &sigma;=0.45 (*Show a drawing and the `distrib()` code that you used*), 
1. make a decision about the hypotheses using &alpha;=0.05 (*show you work*), and 
1. summarize your findings in terms of body temperatures.

<br>

## Beetle Size
Researchers examined the size of two different species of beetles. They hypothesized that the thorax length of the *Halticus oleracea* species would be greater than 190 &mu;m. Use their results for the *Halticus oleracea* species in `r tbls("beetles",display="cite")` to answer the questions further below.

`r tbls("beetles")`

```{r echo=FALSE, background="white"}
d <- read.csv("https://github.com/droglenc/NCData/raw/master/Beetles.csv")
Summarize(thorax~species,data=d,digits=2)
```

1. construct statistical hypotheses (*make sure to define* &mu;),
1. write the definition of the p-value specific to this question,
1. compute the p-value assuming that &sigma;=15 (*Show a drawing and the `distrib()` code that you used*),
1. make a decision about the hypotheses using &alpha;=0.05 (*show your work*), and 
1. summarize your findings in terms of the beetle's thorax size.

<br>

## Identifying Type I and II Errors

In each situation below, (i) write what H<sub>0</sub> and H<sub>A</sub> mean in words,  (ii) describe what a Type I and a Type II error is, and (iii) make an argument for which error is more egregious. [*Hint: See the example about the rare plant in the module reading.*]

1. The United States Department of Agriculture's (USDA) limit for salmonella contamination for chicken parts is 15.4%.[^2] Suppose that a meat inspector reports that the chicken produced by a particular company exceeds the USDA limit. A hypothesis test will be performed to determine whether the meat inspector's claim appears to be true or not.
1. The process of manufacturing a certain drug is quite complex. Small departures from the standard procedure can result in a drug that does not meet specifications and is dangerous to the user. Before being released for use, a sample of a batch of the drug is tested to determine if the mean amount of a particular chemical in the drug meets specifications. A hypothesis test will be performed to determine if the drug met specifications or not.

&nbsp;

[^1]: This question was adapted from [Shoemaker (1996)](http://www.amstat.org/publications/jse/v4n2/datasets.shoemaker.html).

[^2]: Here is the [source](http://www.foodsafetynews.com/2015/01/fsis-releases-new-salmonella-campylobacter-performance-standards-for-poultry/) for salmonella threshold.
