---
layout: page
title: Additional Exercises
subtitle: Normal Distributions Introduction
author: Derek H. Ogle
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/CE/")
source("../../rhelpers/rhelpers.R")
modHTML("NormalDist1_CE2")
```

```{r echo=FALSE, results='hide', message=FALSE}
source("../../rhelpers/knitr_setup.R")
library(ggplot2)
```

&nbsp;

1. **General Understanding**
    1. [Normal Distribution Characteristics II](#normal-distribution-characteristics-ii)
    1. [Normal Distribution Characteristics III](#normal-distribution-characteristics-iiI)
    1. [Normal Distribution Characteristics IV](#normal-distribution-characteristics-iv)
    1. [68-95-99.7 (Empirical) Rule](#68-95-997-empirical-rule)
    1. [Hand Calculations II](#hand-calculations-ii)

----

# General Understanding
## Normal Distribution Characteristics II
For each situation below, (i) identify &mu;, (ii) identify &sigma;, and (iii) draw the normal distribution with an approximately accurate scale on the x-axis.
1. Z~N(0,1)
1. Y~N(90,8)


----

## Normal Distribution Characteristics III

Identify the mean and standard deviation for each population on the graph below (HINT: "eyeball" integers).

```{r NormDistGuess1, echo=FALSE, fig.width=3.5, fig.height=3.5}
x <- seq(-5,7,length.out=500)
B <- dnorm(x,-1,1)
A <- dnorm(x,1,2)
df <- data.frame(x=rep(x,2),y=c(A,B),grp=rep(c("A","B"),each=500))
ggplot(data=df,mapping=aes(x=x,y=y,color=grp,fill=grp)) +
  geom_line(size=1.5) +
  scale_y_continuous(expand=expansion(mult=c(0,0.01))) +
  scale_x_continuous(expand=expansion(mult=0),breaks=seq(-5,7.5,1)) +
  annotate(geom="text",x=c(-1.7,2),y=c(0.38,0.2),label=c("A","B"),size=5) +
  theme_bw() +
  theme(legend.position="none",
        axis.title.y=element_blank(),axis.text.y=element_blank())
```

----

## Normal Distribution Characteristics IV

Identify the mean and standard deviation for each population on the graph below (HINT: "eyeball" integers).

```{r NormDistGuess2, echo=FALSE, fig.width=3.5, fig.height=3.5}
x <- seq(-20,60,length.out=500)
A <- dnorm(x,20,10)
B <- dnorm(x,30,5)
df <- data.frame(x=rep(x,2),y=c(A,B),grp=rep(c("A","B"),each=500))
ggplot(data=df,mapping=aes(x=x,y=y,color=grp,fill=grp)) +
  geom_line(size=1.5) +
  scale_y_continuous(expand=expansion(mult=c(0,0.01))) +
  scale_x_continuous(expand=expansion(mult=0),breaks=seq(-20,60,10)) +
  annotate(geom="text",x=c(12,35),y=c(0.035,0.075),label=c("B","A"),size=5) +
  theme_bw() +
  theme(legend.position="none",
        axis.title.y=element_blank(),axis.text.y=element_blank())
```

----

## 68-95-99.7 (Empirical) Rule

You can click on each question to see an answer. Click on it again to close the answer.

<div class="panel-group">

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#EmpRule1">On any normal distribution, what percentage of individuals is within &plusmn;1&sigma; of &mu;?</a>
</div>
</div>
<div id="EmpRule1" class="panel-collapse collapse">
<div class="panel-body">68%</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#EmpRule2">On any normal distribution, what percentage of individuals is within &plusmn;3&sigma; of &mu;?</a>
</div>
</div>
<div id="EmpRule2" class="panel-collapse collapse">
<div class="panel-body">99.7%</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#EmpRule3">On any normal distribution, what percentage of individuals is greater than &mu;+&sigma;?</a>
</div>
</div>
<div id="EmpRule3" class="panel-collapse collapse">
<div class="panel-body">16%</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#EmpRule4">On any normal distribution, what percentage of individuals is less than &mu;-2&sigma;?</a>
</div>
</div>
<div id="EmpRule4" class="panel-collapse collapse">
<div class="panel-body">2.5%</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#EmpRule5">On any normal distribution, what percentage of individuals is between &mu;-2&sigma; and &mu;+&sigma;?</a>
</div>
</div>
<div id="EmpRule5" class="panel-collapse collapse">
<div class="panel-body">81.5% (using the results from the previous two questions ... 100-16-2.5).</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#EmpRule6">On a N(1,1), what percentage of individuals are negative?</a>
</div>
</div>
<div id="EmpRule6" class="panel-collapse collapse">
<div class="panel-body">16% (Note that negative means less than 0, 0 is one &sigma; below &mu; on this distribution, and 16% of individuals are less than one &sigma; below &mu;.</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#EmpRule7">On a N(100,20), what percentage of individuals are less than 80?</a>
</div>
</div>
<div id="EmpRule7" class="panel-collapse collapse">
<div class="panel-body">16% (16% of individuals are less than one &sigma; below &mu;.</div>
</div>
</div>
  
<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#EmpRule8">On a N(-20,100), what percentage of individuals are greater than 80?</a>
</div>
</div>
<div id="EmpRule8" class="panel-collapse collapse">
<div class="panel-body">16% (16% of individuals are greater than one &sigma; above &mu;.</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#EmpRule9">On a N(50,20), what percentage of individuals are less than 60?</a>
</div>
</div>
<div id="EmpRule9" class="panel-collapse collapse">
<div class="panel-body">This question cannot be answered with the 68-95-99.7 Rule because 60 is not an integer number of &sigma;s away from &mu;.</div>
</div>
</div>

</div>

----

## Hand Calculations II
Answer each question below, **without using R**, assuming that Z~N(0,1). Show your work with a careful drawing for each question.

1. What is the value of Z such that 16% of Z values are greater?
1. What percent of Z values are greater than 2?
1. What percent of Z values are less than 3?
1. What is the most common 99.7% of Z values?
