---
title: "20 - R Filtering Data"
description: |
  Introduction to creating subsets of data frames in R.
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,fig.align='center',comment="#R>  ")
library(NCStats)
```

In the [Getting Data into R module](RData.html), you learned how to retrieve data from the class webpage, enter your own data into a CSV file, load that data into R, and how to view that data in R. In this module, we will learn how to create subsets (i.e., filter) of a data frame into smaller data frames. For example, you may want to create a data frame that contains just male bears from a data frame with both male and female bears, or a data frame that contains only sales during summer months from a data frame that contains all sales. Less often you may wish to eliminate a particular individual from the data frame, perhaps if it is considered to be erroneous.

This module also uses the `bears` data frame from the [Getting Data into R module](RData.html), which is shown below for reference.

```{r eval=FALSE}
library(NCStats)
setwd{"c:/data"}
bears <- read.csv("Bears.csv")
bears
```
```{r echo=FALSE}
( bears <- read.csv("data/Bears.csv") )
```


&nbsp;

# Filtering a data frame
It is common to create a new data frame that contains only some of the individuals from an existing data frame. The process of creating the newer, smaller data frame is called filtering (or subsetting) and is accomplished with `filterD()`.^[`filterD()` requires the `NCStats` package.] The `filterD()` function requires the original data frame as the first argument and a condition statement as the second argument. The condition statement is used to either include or exclude individuals from the original data frame. Condition statements consist of the name of a variable in the original data frame, a comparison operator, and a comparison value (Table \@ref(tab:filterComps)). The results from `filterD()` should be assigned to an object, which is then the name of the new data frame.

<aside>
`filterD()` is used to create a subset of a data frame.
</aside>

&nbsp;

```{r echo=FALSE}
tmp <- data.frame(
  c('`var==value`','`var!=value`','`var %in% c(value1,value2)`',
    '`var>value`','`var>=value`','`var<value`','`var<=value`',
    'condition1,condition2','condition1 | condition2'),
  c('All rows where `var` **IS equal** to `value`',
    'All rows where `var` **is NOT equal** to `value`',
    'All rows where `var` **IS IN** vector of `value`s^[`value` should be a character, factor, or integer.]',
    'All rows where `var` is **greater than** `value`^[`value` must be numeric.]',
    'All rows where `var` is **greater than or equal to** `value`^[`value` must be numeric.]',
    'All rows where `var` is **less than** `value`^[`value` must be numeric.]',
    'All rows where `var` is **less than or equal to** `value`^[`value` must be numeric.]',
    'All rows where **BOTH** conditions are true',
    'All rows where **ONE or BOTH** conditions are true^[Note that this "or" operator is a "vertical line"" which is typed with the shift-backslash key.]')
)
names(tmp) <- c("Comparison Operator",
                "Rows Returned from Original Data Frame")
knitr::kable(tmp,label="filterComps",caption="Comparison operators used in `filterD()` and their results. Note that `var` generically represents a variable in the original data frame and `value` is a generic value or level. Both `var` and `val` would be replaced with specific items (see examples in main text).")
```


&nbsp;

The following are examples of new data frames created from `bears`. The name of the new data frame (i.e., object left of the assignment operator) can be any valid object name. As demonstrated below, the new data frame should be examined after each filtering to ensure that the data frame actually contains the items that you desire.^[For larger data.frames you should check the structure (using `str()`) or the `headtail()` of the new data frame.]

<aside>
Examine the new data frame after subsetting to ensure that it contains the data you intended.
</aside>

* Only individuals from *Bayfield* county.

```{r}
bf <- filterD(bears,loc=="Bayfield")
bf
```

* Individuals from both *Bayfield* and *Ashland* counties.

```{r}
bfash <- filterD(bears,loc %in% c("Bayfield","Ashland"))
bfash
```

* Individuals with a weight greater than 100 kg.

```{r}
gt100 <- filterD(bears,weight.kg>100)
gt100
```

* Individuals from *Douglas* County that weighed at least 150 kg.

```{r}
do150 <- filterD(bears,loc=="Douglas",weight.kg>=150)
do150
```


&nbsp;

# Selecting Individuals
In some instances, you may need to select or exclude an individual from a data frame. Positions within an object are identified within square brackets. As data frames are two-dimensional objects they are indexed by a row and a column, in that order. For example, the item in the third row and second column of `bears` is selected below.

```{r}
bears[3,2]
```

An entire row or column may be selected by omitting the other dimension. For example, one could select the entire second column with `bears[,2]`, but this is also the `weight.kg` variable and is better selected with `bears$weight.kg`. As a better example, the entire third row is selected below (note that the column designation was omitted).

```{r}
bears[3,]
```

Multiple rows are selected by combining row indices together with `c()`. For example, the third, fifth, and eighth rows are selected below (again, the column index is omitted).

```{r}
bears[c(3,5,8),]
```

Finally, rows can be excluded by preceding the row indices with a negative sign.

```{r}
bears[-c(3,5,8,10,12),]
```


# Corrections {.appendix}
If you see any errors (code, typographical, or logical) please bring these to [my attention](mailto:dogle@northland.edu).
