---
title: "Getting Started with R"
description: |
  Installing R and a Brief Introduction.
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```


# Setting Up R
R is a software environment for performing statistical analyses. RStudio is a helper program that makes it easier to use R. Herein, I will refer to R but you will interact with R through RStudio.

Detailed methods for downloading, installing, and configuring R, RStudio, and `NCStats` on your personal computer are given on the [Resources page of the course website](http://derekogle.com/NCMTH107/resources#computer).


&nbsp;

# R Basics
## Saving
Results are not saved in R. Rather, "scripts" of successful R commands are saved and, then, if the analysis needs to be re-done, the entire set of commands is opened and run again. When writing a report, all tabular and graphical output should be copied from R and pasted into your report document. This document will serve as your analysis report and can be modified to include answers to questions, references to the tables and graphs, etc. Most data used in this class will be entered into R through "comma-separated values" text file (see \@ref(sect:REnterData)).

R does allow one to save a "workspace", though I urge you not to do that. Rather, save your "good" commands in a script and save your "good" results in a report document; do not save the workspace.

<aside>
Do NOT save the workspace.
</aside>

## Expressions and Assignments
Expressions in R are mathematical "equations" that are evaluated by R with a result seen immediately. An example of an expression in R is

```{r}
5+log(7)-pi
```

where `log()` and `pi` are built-in functions used to compute the natural log and find the value of $\pi$, respectively. Expressions in R are like using a calculator where the result is shown, but not saved for subsequent analyses. In addition, expressions in R follow the same order of operations and use of parentheses as expressions entered into your calculator.

Results from an expression are typically saved for further computations by assigning the results to an object with the assignment operator (i.e., `<-`). The general form for saving the result of an expression into an object is `object <- expression`. The result of the expression will not be seen unless the object name is subsequently typed into R (but see below). For example, the result of the previous expression is saved into an object called `x` and then viewed with

```{r}
x <- 5+log(7)-pi
x
```

The result of an expression can be both assigned and printed by surrounding the command in parentheses. For example, the following assigns the result of the expression to `y` and prints the result.^[The spaces between the expression and the parentheses are only needed to increase legibility.]

```{r}
( y <- 15*exp(2) )
```

An object can be named whatever you want, with the exception that it cannot start with a number or symbol (e.g., `!`, `@`), contain a space, or be the name of a reserved word or function (e.g., `pi` or `log`). Object names should be short and simple enough that you can remember what is contained in the object. It is also good practice to view the object immediately after making the assignment to make sure that it contains results that seem appropriate.

<aside>
Object names should be short and informative.
</aside>


&nbsp;

# Functions and Arguments
R contains many "programs," or functions, to perform particular tasks. A function is "called" by typing the function name followed by open and closed parentheses.^[Regular curved parentheses have two primary uses in R: (1) to control order of operations in expressions (as with a calculator) and (2) to contain the arguments sent to a function.] The `log()` function, used in the previous section, is an example of a function. Arguments, which the function will use to perform its task, are contained within the parentheses. In `log()` above, the argument is the number in parentheses (and for which the natural log will be computed). Many other functions will be described below and in subsequent modules.

<aside>
Function names are followed by parentheses.

Arguments are within function parentheses.
</aside>

&nbsp;

# NCStats Package
`NCStats` is a set of R functions that makes the statistical methods used in this class easier. You need to load the `NCStats` package at the beginning of every R script used in this class. Packages are loaded by including the package name within `library()`.

```{r}
library(NCStats)
```
