---
layout: page
title: Confidence Regions (Z)
author: Derek H. Ogle
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE, warning=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/Explanations/")
source("../../rhelpers/rhelpers.R")
modHTML("Calc_CI_Z")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
source("../../rhelpers/knitr_setup.R")
```

----

## Steps to Follow

Constructing a confidence region is a five-step process as described below.

1. **Determine level of confidence (C) with 100(1-&alpha;)%.**
    * For example, if &alpha;=0.05 then C=100(1-0.95)=95%.
1. **Determine type of confidence region.**
    * If H<sub>A</sub> is a "less than" then construct an "upper confidence bound."
    * If H<sub>A</sub> is a "greater than" then construct an "lower confidence bound."
    * If H<sub>A</sub> is a "not equals" then construct an "confidence interval."
1. <strong>Determine Z<sup>*</sup> by shading on the standard normal distribution (mean=0,SD/SE=1) an area of C according to H<sub>A</sub>.</strong>
    * If H<sub>A</sub> is a "less than" then shade C to the left and Z<sup>*</sup> is positive.
    * If H<sub>A</sub> is a "greater than" then shade C to the right and Z<sup>*</sup> is negative.
    * If H<sub>A</sub> is a "not equals" then shade C in the middle to get two Z<sup>*</sup>s that are positive and negative. (To use `distrib()` here you will need to put half of the area outside of C into each tail of the distribution.)
1. **Construct confidence region value(s) with $\bar{x}+Z^{*}\frac{\sigma}{\sqrt{n}}$.**
1. **Interpret the confidence region.** *If the question has a context then the parameter should be stated within the context of the question.*
    * If H<sub>A</sub> is a "less than" then the parameter (i.e., population mean) is said to be less than the confidence region value.
    * If H<sub>A</sub> is a "greater than" then the parameter (i.e., population mean) is said to be greater than the confidence region value
    * If H<sub>A</sub> is a "not equals"  then the parameter (i.e., population mean) is said to be between the two confidence region values.

----

## First Example
Suppose that I hypothesized that the mean time to take a quiz is less than 100 minutes and that, in a sample of 30 students, the mean time to take the quiz was 95.2 minutes. Further, suppose that &sigma;=20 and &alpha;=0.10.

1. C=100(1-0.10)=90%
1. An upper confidence bound because H<sub>A</sub> is a "greater than."
1. Z<sup>*</sup>=`r formatC(distrib(0.9,type="q"),format="f",digits=3)` as shown below.
1. The confidence region value is thus 95.2+`r formatC(distrib(0.9,type="q"),format="f",digits=3)`$\frac{20}{\sqrt{30}}$ = 95.2+`r formatC(distrib(0.9,type="q")*20/sqrt(30),format="f",digits=1)`=`r formatC(95.2+distrib(0.9,type="q")*20/sqrt(30),format="f",digits=1)`.
1. I am 90% confident that the mean time for all students to take the quiz is less than `r formatC(95.2+distrib(0.9,type="q")*20/sqrt(30),format="f",digits=1)` minutes.

```{r}
distrib(0.9,type="q")
```

----

## Second Example
Suppose that I hypothesized that the mean time to walk between the Science Center and the Ponzio Center is more than 7 minutes and that, in a sample of 50 students, the mean time to make this walk was 8.3 minutes. Further, suppose that &sigma;=1.3 and &alpha;=0.05.

1. C=100(1-0.05)=95%
1. A lower confidence bound because H<sub>A</sub> is a "less than."
1. Z<sup>*</sup>=`r formatC(distrib(0.9,type="q",lower.tail=FALSE),format="f",digits=3)` as shown below.
1. The confidence region value is thus 8.3-`r formatC(distrib(0.95,type="q"),format="f",digits=3)`$\frac{1.3}{\sqrt{50}}$ = 8.3+`r formatC(distrib(0.95,type="q")*1.3/sqrt(50),format="f",digits=2)`=`r formatC(8.3+distrib(0.95,type="q")*1.3/sqrt(50),format="f",digits=2)`.
1. I am 90% confident that the mean time for all students to walk between the Science Center and the Ponzio Center is more than `r formatC(8.3+distrib(0.95,type="q")*1.3/sqrt(50),format="f",digits=2)` minutes.

```{r}
distrib(0.95,type="q",lower.tail=FALSE)
```


----

## Third Example
Suppose that I hypothesized that the mean time students spend preparing for the prep check is different than 30 minutes and that, in a sample of 40 students, the mean preparation time was 37.8 minutes. Further, suppose that &sigma;=9.7 and &alpha;=0.01.

1. C=100(1-0.01)=99%
1. A confidence interval because H<sub>A</sub> is a "not equals."
1. Z<sup>*</sup>=&plusmn;`r formatC(distrib(0.995,type="q"),format="f",digits=3)` as shown below.
1. The confidence region value is thus 37.8&plusmn;`r formatC(distrib(0.995,type="q"),format="f",digits=3)`$\frac{9.7}{\sqrt{40}}$ = 37.8&plusmn;`r formatC(distrib(0.995,type="q")*9.7/sqrt(40),format="f",digits=2)`=(`r formatC(37.8-distrib(0.995,type="q")*9.7/sqrt(50),format="f",digits=2)`,`r formatC(37.8+distrib(0.995,type="q")*9.7/sqrt(50),format="f",digits=2)`).
1. I am 99% confident that the mean time that all students spend preparing for the prep checks is between `r formatC(37.8-distrib(0.995,type="q")*9.7/sqrt(50),format="f",digits=2)` and `r formatC(37.8+distrib(0.995,type="q")*9.7/sqrt(50),format="f",digits=2)` minutes.

```{r}
distrib(0.005,type="q")
```

----

