---
layout: page
title: Review Exercise
subtitle: Hypothesis Testing Errors
author: Derek H. Ogle
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/RE/")
source("../../rhelpers/rhelpers.R")
modHTML("HypTestingErrs_RevEx")
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../rhelpers/knitr_setup.R")
```

----

&nbsp;

1. [Power and Type II Error Relationships](#power-and-type-ii-error-relationships)
1. [Power Calculations (Without Context)](#power-calculations)

&nbsp;

&nbsp;

----

## Power and Type II Error Relationships

You can click on each question to see an answer. Click on it again to close the answer.

<div class="panel-group">

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#PowerBeta">What is $\beta$ if power=0.875?</a>
</div>
</div>
<div id="PowerBeta" class="panel-collapse collapse">
<div class="panel-body">$\beta$=1-power = 1-0.875 = 0.125.</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#poweralpha">For a constant sample size, $\sigma$, and difference between the hypothesized and actual means, what happens to power, if $\alpha$ is INcreased?</a>
</div>
</div>
<div id="poweralpha" class="panel-collapse collapse">
<div class="panel-body">Power INcreases.</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#powern">For a constant $\alpha$, $\sigma$, and difference between the hypothesized and actual means, what happens to power, if the sample size INcreases?</a>
</div>
</div>
<div id="powern" class="panel-collapse collapse">
<div class="panel-body">Power INcreases.</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#powermu">For a constant $\alpha$, $\sigma$, and sample size, what happens to power if the difference between the hypothesized and actual means increases?</a>
</div>
</div>
<div id="powermu" class="panel-collapse collapse">
<div class="panel-body">Power INcreases.</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#betalpha">For a constant sample size, $\sigma$, and difference between the hypothesized and actual means, what happens to $\beta$, if $\alpha$ is INcreased?</a>
</div>
</div>
<div id="betalapha" class="panel-collapse collapse">
<div class="panel-body">$\beta$ DEcreases.</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#betan">For a constant $\alpha$, $\sigma$, and difference between the hypothesized and actual means, what happens to $\beta$, if the sample size is increased?</a>
</div>
</div>
<div id="betan" class="panel-collapse collapse">
<div class="panel-body">$\beta$ DEcreases.</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#betamu">For a constant $\alpha$, $\sigma$, and sample size, what happens to $\beta$ if the difference between the hypothesized and actual means INcreases?</a>
</div>
</div>
<div id="betamu" class="panel-collapse collapse">
<div class="panel-body">$\beta$ DEcreases.</div>
</div>
</div>

<div class="panel panel-default">
<div class="panel-heading">
<div class="panel-title">
<a data-toggle="collapse" href="#reallife">Describe a real-life situation where you think that making a Type II error would be much more ``costly'' than making a Type I error. Completely describe the situation at hand and what Type I and a Type II errors mean in terms of the situation you describe.</a>
</div>
</div>
<div id="reallife" class="panel-collapse collapse">
<div class="panel-body">The classic example deals with the population dynamics of an animal.  The null hypothesis says that the animal's population size is not declining, it is remaining stable over time.  The alternative hypothesis is that the animal's population is declining.  If the null hypothesis is true and you reject it (i.e., you say the population is declining when it really isn't) you have made a type I error.  This will likely lead resource managers to take actions to protect the animal.  This is not necessarily a bad thing for the animal, although it may be costly.  Conversely, if the null hypothesis is false and you fail to reject it (i.e., the animal is really declining but you say that it isn't), then you have made a type II error.  This is much more egregious mistake because now you will take no action to protect the animal; it may go extinct.  To me, losing an animal is much more costly (important) than overprotecting the animal.</div>
</div>
</div>


</div>

----

## Power Calculations

1. Compute power given the following information: $\alpha=0.05$, $H_{0}:\mu=50$, $H_{A}:\mu<50$, $\sigma=20$, $n=50$, and $\mu_{A}=45$. [See answer here](zRevExAns/HypTestingErrs.html#power-calculations-1)
1. Compute power given the following information: $\alpha=0.10$, $H_{0}:\mu=10$, $H_{A}:\mu>10$, $\sigma=5$, $n=25$, and $\mu_{A}=12$. [See answer here](zRevExAns/HypTestingErrs.html#power-calculations-2)
1. Compute power given the following information: $\alpha=0.01$, $H_{0}:\mu=75$, $H_{A}:\mu>75$, $\sigma=15$, $n=30$, and $\mu_{A}=82$. [See answer here](zRevExAns/HypTestingErrs.html#power-calculations-3)

----
