---
layout: page
author: Derek H. Ogle
title: MTH107 FAQ
subtitle: When I load NCStats, R cannot find some other packages
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/FAQ/FAQs")
source("../../../rhelpers/rhelpers.R")
fnm <- "install-other-packages"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message='FALSE'}
source("../../../rhelpers/knitr_setup.R")
```

----

`NCStats` is a set of functions that I have written to make aspects of using R for statistics classes at Northland easier.  A method for installing `NCStats` is described [in this FAQ](install-NCStats.html).  `NCStats` depends on other packages being installed and, at times, the installation of `NCStats` will not properly install these other pacakges.  If this is the case, then typing

```{r eval=FALSE}
library(NCStats)
```

will return an error where R complains that some package is not installed.  In this instance, one should note the name of the package that is not installed and then follow [these directions](http://derekogle.com/IFAR/supplements/installations/InstallPackagesRStudio.html) for installing the package.
