---
title: "Review Exercise Key"
subtitle: "Univariate EDA"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/RE/zRevExAns/")
source("../../../rhelpers/rhelpers.R")
modHTML("UEDAQuant2_EDAs")
```
```{r echo=FALSE}
library(NCStats)
library(captioner)
figs <- captioner(prefix="Figure")
tabs <- captioner(prefix="Table")
```

## Time Spent Playing Video Games

The time spent playing video games is very strongly right-skewed with an extreme outlier at 30 hours and another outlier at approximately 15 hours. The center is at a median of 0 hours and the dispersion is from a Q1 of 0 hours to a Q3 of 1.2 hours. The median and IQR were used because of the skew and presence of outliers. [*Results are derived from the figure and table presented in the question.*]

----

## Bank Transaction Times

The bank transaction times are strongly right-skewed with no obvious outliers. The center is at a median of 93.0 hours and the dispersion is from a Q1 of 60.3 hours to a Q3 of 145.2 hours. The median and IQR were used because of the skew and presence of outliers. [*Results are derived from the figure and table presented in the question.*]

----

## Creatine Phosphate Concentrations
```{r echo=FALSE, results='hide'}
cp <- c(121,82,100,151,68,58,95,145,64,119,104,110,113,118,203,62,83,67,201,
101,163,84,57,139,60,78,94,93,92,110,25,123,70,48,95,42)
cpsum <- Summarize(~cp,digits=1)
figs(name="CPHist","Histogram of the creatine phosphate concentrations in 36 male volunteers.")
tabs(name="CPSum","Summary statistics of the creatine phosphate concentrations in 36 male volunteers.")
```

The distribution of creatine phosphate concentrations is slightly right-skewed with no obvious outliers (`r figs("CPHist",display="cite")`). The center is at a mean of `r formatC(cpsum["mean"],format="f",digits=1)` and the dispersion is a standard deviation of `r formatC(cpsum["sd"],format="f",digits=1)` (`r tabs("CPSum",display="cite")`). The mean and standard deviation were used because the distribution was not strongly skewed and outliers were not present.  [*Note that I did not call the individuals above 200 outliers because they were only one bar removed, there were two individuals, and they "felt" like an extension of the right-tail.*]

```{r CPHist, echo=FALSE, comment=""}
par(mar=c(3.05,3.05,0.65,0.65),mgp=c(1.9,0.3,0),tcl=-0.2)
hist(~cp,xlab="Creatine Phosphate Concentration")
```

`r figs("CPHist")`

`r tabs("CPSum")`
```{r CPSum, echo=FALSE, comment=""}
cpsum
```


### R Code Appendix
```
library(NCStats)
df <- read.csv("data/CreatinePhosphate.csv")
hist(~cp,data=df,xlab="Creatine Phosphate Concentration")
Summarize(~cp,data=df,digits=1)
```

----

## Air Pollution
```{r echo=FALSE, results='hide'}
polln <- c(30,30,34,36,37,38,40,42,43,43,45,52,55,58,58,58,59,63,63,71,75,85,86,86,99,102,102,141,153,261,21)
pollnsum <- Summarize(~polln,digits=1)
figs(name="POLLNHist","Histogram of the carbon monoxide levels arising from the oil refinery stack.")
tabs(name="POLLNSum","Summary statistics of the carbon monoxide levels arising from the oil refinery stack.")
```

The distribution of carbon monoxide levels arising from the oil refinery stack is strongly right-skewed with a possible outlier at 261 ppm (`r figs("POLLNHist",display="cite")`). The center is at a median of `r formatC(pollnsum["median"],format="f",digits=1)` and the dispersion is from a Q1 of `r formatC(pollnsum["Q1"],format="f",digits=1)` to a Q3 of `r formatC(pollnsum["Q3"],format="f",digits=1)` (`r tabs("POLLNSum",display="cite")`). The median and IQR were used because the distribution was strongly skewed and an outlier was present.

```{r POLLNHist, echo=FALSE, comment=""}
par(mar=c(3.05,3.05,0.65,0.65),mgp=c(1.9,0.3,0),tcl=-0.2)
hist(~polln,xlab="Carbon Monoxide Levels")
```

`r figs("POLLNHist")`

`r tabs("POLLNSum")`
```{r POLLNSum, echo=FALSE, comment=""}
pollnsum
```


### R Code Appendix
```
library(NCStats)
df <- read.csv("data/Polln.csv")
hist(~polln,data=df,xlab="Carbon Monoxide Levels")
Summarize(~polln,data=df,digits=1)
```

----

## Dow Jones
```{r echo=FALSE, results='hide'}
dj <- c(152,180,167,119,115,113,119,135,127,161,140,126,114,133,205,104,149,124,106,179)
djsum <- Summarize(~dj,digits=1)
figs(name="DJHist","Histogram of the Dow Jones Travel Index for May, 1996.")
tabs(name="DJSum","Summary statistics of the Dow Jones Travel Index for May, 1996.")
```

The distribution of the May, 1996 Dow Jones Travel Index is strongly right-skewed with no obvious outliers present (`r figs("DJHist",display="cite")`). The center is at a median of `r formatC(djsum["median"],format="f",digits=1)` and the dispersion is from a Q1 of `r formatC(djsum["Q1"],format="f",digits=1)` to a Q3 of `r formatC(djsum["Q3"],format="f",digits=1)` (`r tabs("DJSum",display="cite")`). The median and IQR were used because the distribution was strongly skewed.

```{r DJHist, echo=FALSE, comment=""}
par(mar=c(3.05,3.05,0.65,0.65),mgp=c(1.9,0.3,0),tcl=-0.2)
hist(~dj,xlab="Dow Jones Travel Index")
```

`r figs("DJHist")`

`r tabs("DJSum")`
```{r DJSum, echo=FALSE, comment=""}
djsum
```


### R Code Appendix
```
library(NCStats)
df <- read.csv("data/DowJonesTravel.csv")
hist(~djt,data=df,xlab="Dow Jones Travel Index")
Summarize(~djt,data=df,digits=1)
```

----

## Zoo Sizes
```{r echo=FALSE, results='hide'}
zoo <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/Zoo2.csv")
zoosum <- Summarize(~size,data=zoo,digits=1)
figs(name="ZOOHist","Histogram of the size (acres) of a sample of zoos.")
tabs(name="ZOOSum","Summary statistics of the size (acres) of a sample of zoos.")
```

The distribution of the physical size of zoos is strongly right-skewed with possible outliers between 500 and 600 acres (`r figs("ZOOHist",display="cite")`). The center is at a median of `r formatC(zoosum["median"],format="f",digits=1)` acres and the dispersion is from a Q1 of `r formatC(zoosum["Q1"],format="f",digits=1)` to a Q3 of `r formatC(zoosum["Q3"],format="f",digits=1)` acres (`r tabs("ZOOSum",display="cite")`). The median and IQR were used because the distribution was strongly skewed and there were possible outliers.

```{r ZOOHist, echo=FALSE, comment=""}
par(mar=c(3.05,3.05,0.65,0.65),mgp=c(1.9,0.3,0),tcl=-0.2)
hist(~size,data=zoo,xlab="Size (acres)",w=50)
```

`r figs("ZOOHist")`

`r tabs("ZOOSum")`
```{r ZOOSum, echo=FALSE, comment=""}
zoosum
```


### R Code Appendix
```
library(NCStats)
df <- read.csv("data/Zoos.csv")
hist(~size,data=df,xlab="Size (acres)",w=50)
Summarize(~size,data=df,digits=1)
```
