---
title: "Review Exercise Key"
subtitle: "Summaries for One Quantitative Variable"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: false
    lib_dir: zlibs
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/RE/zRevExAns/")
source("../../../rhelpers/rhelpers.R")
modHTML("UEDAQuant1")
```
```{r echo=FALSE, results="hide"}
library(NCStats)
library(captioner)
figs <- captioner(prefix="Figure")
figs(name="CPHist","Histogram of the creatine phosphokinase concentrations in 36 male volunteers.")
figs(name="POLLNHist","Histogram of the carbon monoxide levels arising from the oil refinery stack.")
```

## Number of States Visited

The calculation of the median and IQR is shown below.
```{r echo=FALSE, comment=""}
states <- c(8, 3, 17, 15, 4, 19)
iqrCalc(states)
```

The calculation of the mean and standard deviation is shown below.
```{r echo=FALSE, comment=""}
sdCalc(states)
```

----

## Home Range of Red Kangaroos

The calculation of the median and IQR is shown below.
```{r echo=FALSE, comment=""}
kang <- c(258, 310, 560, 470, 502)
iqrCalc(kang)
```

The calculation of the mean and standard deviation is shown below.
```{r echo=FALSE, comment=""}
sdCalc(kang)
```

----

## Brule River Gauge Heights

The calculation of the median and IQR is shown below.
```{r echo=FALSE, comment=""}
hts <- c(1.56,1.54,1.54,1.57,1.58,1.61,1.60,1.69,
         1.99,2.11,1.98,1.76,1.69,1.99,1.86,1.53)
iqrCalc(hts)
```

The calculation of the mean and standard deviation is shown below.
```{r echo=FALSE, comment=""}
sdCalc(hts)
```

### R Code Appendix
```
library(NCStats)
df <- read.csv("data/BruleRiver.csv")
Summarize(~gauge,data=df)
```

----

## Population Density in Wisconsin Counties

The calculation of the median and IQR is shown below.
```{r echo=FALSE, comment=""}
dens <- c(429.0,67.8,52.1,97.4,57.9,354.9,16.2,19.1,127.0,27.6,10.2,54.6,28.8,30.1,20.2)
iqrCalc(dens)
```

The calculation of the mean and standard deviation is shown below.
```{r echo=FALSE, comment=""}
sdCalc(dens)
```

### R Code Appendix
```
library(NCStats)
df <- read.csv("data/WIDensity.csv")
Summarize(~dens,data=df)
```

----

## Creatine Phosphate Concentrations

The calculation of the median and IQR is shown below.
```{r echo=FALSE, comment=""}
cp <- c(121,82,100,151,68,58,95,145,64,119,104,110,113,118,203,62,83,67,201,
101,163,84,57,139,60,78,94,93,92,110,25,123,70,48,95,42)
iqrCalc(cp)
```

The calculation of the mean and standard deviation is shown below.
```{r echo=FALSE, comment=""}
sdCalc(cp)
```

The histogram is in `r figs("CPHist",display="cite")`.

```{r CPHist, echo=FALSE, comment=""}
hist(~cp,xlab="Creatine Phosphate Concentration (ppm)")
```

`r figs("CPHist")`

### R Code Appendix
```
library(NCStats)
df <- read.csv("data/CreatinePhosphate.csv")
Summarize(~cp,data=df)
<<CPHist>>
```

----

## Air Pollution

The calculation of the median and IQR is shown below.
```{r echo=FALSE, comment=""}
polln <- c(30,30,34,36,37,38,40,42,43,43,45,52,55,58,58,58,59,63,63,71,75,85,86,86,99,102,102,141,153,261,21)
iqrCalc(polln)
```

The calculation of the mean and standard deviation is shown below.
```{r echo=FALSE, comment=""}
sdCalc(polln)
```

The histogram is in `r figs("POLLNHist",display="cite")`.

```{r POLLNHist, echo=FALSE, comment=""}
hist(~polln,xlab="Carbon Monoxide Levels (ppm)")
```

`r figs("POLLNHist")`

# R Code Appendix
```
library(NCStats)
df <- read.csv("data/Polln.csv")
Summarize(~polln,data=df)
<<POLLNHist>>
```

----
