---
layout: page
title: "Class Example"
subtitle: "Filtering Data In R"
author: "Derek H. Ogle"
css: "/css/modules.css"
output:
  html_document:
    fig_width: 3.5
    fig_height: 3.5
    self_contained: false
    lib_dir: libs
---
```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/CEx")
source("../../rhelpers/rhelpers.R")
fnm <- "FilteringDataInR_CExmpl"
modHTML(fnm)
purl2(paste0(fnm,".Rmd"))
```
```{r echo=FALSE, results='hide', message=FALSE}
fnm <- "FilteringDataInR"
source("../../rhelpers/knitr_setup.R")
```

----

## Load Data From Webpage

Load **Iris.csv** into an R data.frame called `dfiris`. *Open your script and copy your code from last class period.*
```{r echo=FALSE}
dfiris <- read.csv("https://raw.githubusercontent.com/droglenc/NCData/master/Iris.csv")
```
```{r message=FALSE}
library(NCStats)
```
```{r eval=FALSE}
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/CEx")
dfiris <- read.csv("Iris.csv")
```
```{r}
str(dfiris)
headtail(dfiris)
```





----

## Examining Single Variables

* Examine the sepal length variable.
```{r}
dfiris$seplen
```

* Examine the species variable.
```{r}
dfiris$species
```

## Filtering Data Frames
* Isolate just the `setosa` species.
```{r}
set <- filterD(dfiris,species=="setosa")
str(set)
headtail(set)
```

* Isolate the irises with a sepal length greater than 5 cm.
```{r}
gt5 <- filterD(dfiris,seplen>5)
headtail(gt5)
```

* Remove the `virginica` species or retain the `setosa` and `versicolor` species.
```{r}
setver1 <- filterD(dfiris,species %in% c("setosa","versicolor"))
str(setver1)
headtail(setver1)
```

* Same as above.
```{r}
setver2 <- filterD(dfiris,species!="virginica")
str(setver2)
headtail(setver2)
```

* Isolate the `versicolor` species that had sepal lengths greater than 5 cm.
```{r}
ver_gt5 <- filterD(dfiris,species=="versicolor",seplen>5)
headtail(ver_gt5)
```

* Isolate all irises that had a sepal length between 5 and 9 cm.
```{r}
ver_bw5n9 <- filterD(dfiris,seplen>5,seplen<9)
headtail(ver_bw5n9)
```

----
