---
layout: page
title: "Bivariate EDA"
author: "Derek H. Ogle"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: no
    css: "/css/handouts.css"
  pdf_document:
    fig_width: 3.5
    fig_height: 3.5
  word_document:
    fig_width: 3.5
    fig_height: 3.5
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/class/HOs")
source("../../../rhelpers/rhelpers.R")
fnm <- "BivariateEDA"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message='FALSE'}
fnm <- "BivariateEDA"
source("../../../rhelpers/knitr_setup.R")
```

## Background
Measurements of the levels of arsenic in the drinking water, cooking water, and toenail samples, as well as related covariates, were measured on 21 individuals with private wells in a New Hampshire community.  The variables below were recorded in the [https://github.com/droglenc/NCData/blob/master/Arsenic.csv](Arsenic.csv) file located on the R Resources web page.

* `age`: Age (yrs) of person
* `sex`: Sex of person
* `usedrink`: Household well used for drinking -- A="$<\frac{1}{4}$", B="$\approx\frac{1}{4}$",  C="$\approx\frac{1}{2}$",  D="$\approx\frac{3}{4}$",  E="$>\frac{3}{4}$"
* `usecook`: Household well used for cooking -- A="$<\frac{1}{4}$", B="$\approx\frac{1}{4}$",  C="$\approx\frac{1}{2}$",  D="$\approx\frac{3}{4}$",  E="$>\frac{3}{4}$"
* `arswater`: Arsenic in water (ppm)
* `arsnails`: Arsenic in toenails (ppm)

In this handout, we will consider the questions below.

* What type of variable is each variable in the data set?
* Describe the relationship between the level of arsenic in the toenails and the age of the person.
* Describe the relationship between the "amount" that the well is used for drinking and the sex of the person.
* What percentage of females used the well for drinking water more than three-quarters of the time?
* What percentage of all persons in the study were female and used the well for drinking water about one-half of the time?
* What percentage of the sample was female?

## Getting The Data
```{r message=FALSE, warning=FALSE}
library(NCStats)
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/class/HOs")
Ars <- read.csv("Arsenic.csv")
str(Ars)
view(Ars)
```

## Bivariate EDA -- Quantitative
```{r fig.show='asis'}
plot(arsnails~age,data=Ars,pch=19)
cor(Ars$arsnails,Ars$age)
```

## Bivariate EDA -- Categorical
```{r }
( freq.tbl <- xtabs(~sex+usedrink,data=Ars) )
percTable(freq.tbl,margin=1,digits=1)
percTable(freq.tbl,margin=2,digits=1)
percTable(freq.tbl,digits=1)
```

```{r echo=FALSE, results="hide", message=FALSE, warnings=FALSE}
purl2(paste0(fnm,".Rmd"))
```
