---
layout: page
title: "t-Tests"
author: "Derek H. Ogle"
css: "/css/handouts.css"
output:
  html_document:
    fig_width: 3.5
    fig_height: 3.5
    self_contained: no
  pdf_document:
    fig_width: 3.5
    fig_height: 3.5
  word_document:
    fig_width: 3.5
    fig_height: 3.5
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/class/HOs")
source("../../../rhelpers/rhelpers.R")
fnm <- "tTests"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message='FALSE'}
fnm <- "tTests"
source("../../../rhelpers/knitr_setup.R")
```

## First Commands
```{r message=FALSE, warning=FALSE}
library(NCStats)
setwd("C:/aaaWork/Web/GitHub/NCMTH107/resources/class/HOs")
library(car)   # for leveneTest
```

## t Distribution Calculations
An example of computing the p-value if $H_{A}:\mu>70$, t=2.67, and df=18.
```{r par1a=TRUE}
( distrib(2.67,distrib="t",df=18,lower.tail=FALSE) )
```

An example of finding $t^{*}$ if $H_{A}:\mu\neq70$, $\alpha$=0.05, and df=28.
```{r par1a=TRUE}
( distrib(0.025,distrib="t",df=28,type="q",lower.tail=FALSE) )
```

## One-Sample t-Test
[Blem and Blem (1995)](http://www.jstor.org/stable/1564989?seq=1#page_scan_tab_contents) examined the reproductive characteristics of [Eastern Cottonmouth snakes (*Agkistrodon piscivorus*)](https://en.wikipedia.org/wiki/Agkistrodon_piscivorus), a once widely distributed snake whose numbers have decreased recently due to encroachment by humans.  In one part of their study they determined that the population being examined must have an average litter size greater than 5.8 snakes for the population to grow.  A random sample of snake litters from this population was taken and the number of snakes in each litter was recorded in in [Cottonmouth.csv](https://github.com/droglenc/NCData/raw/master/Cottonmouth.csv).  Test, at a very conservative level, if the average litter size is large enough for this population to grow.
```{r par1a=TRUE}
cm  <- read.csv("Cottonmouth.csv")
str(cm)
# if n was <40 then I would have done -- hist(~num,data=cm,xlab="Number in Litter")
( cm.t <- t.test(cm$num,mu=5.8,alt="greater",conf.level=0.99) )
plot(cm.t)
```

## Two-Sample t-Test
[Sholl *et al.* (2000)](http://www.sciencedirect.com/science/article/pii/S0272494499901469) performed an experiment to test the effect of sex (male, female) on spatial orientation ability.  In one part of their study, the researchers took 30 males and 30 females to an unfamiliar wooded park and asked each individual to point to the south.  The absolute pointing error (positive degrees off from due south, `abserr`) was recorded in [SexDirection.csv](https://raw.githubusercontent.com/droglenc/NCData/master/SexDirection.csv) .  Test if men have a better sense of direction than women, at the 1% level?
```{r}
sdir <- read.csv("SexDirection.csv")
str(sdir)
```

```{r fig.width=2*figsz, fig.height=figsz, par2=TRUE, fig.show='asis'}
Summarize(abserr~sex,data=sdir,digits=1)
hist(abserr~sex,data=sdir,xlab="Absolute Pointing Error")
leveneTest(abserr~sex,data=sdir)
```

```{r par1a=TRUE}
( t2 <- t.test(abserr~sex,data=sdir,var.equal=TRUE,alt="greater",conf.level=0.99) )
plot(t2)
```

```{r echo=FALSE, results="hide", message=FALSE, warnings=FALSE}
purl2(paste0(fnm,".Rmd"))
```
