---
layout: page
title: "Univariate EDA"
author: "Derek H. Ogle"
output:
  html_document:
    fig_width: 3.5
    fig_height: 3.5
    self_contained: no
  pdf_document:
    fig_width: 3.5
    fig_height: 3.5
  word_document:
    fig_width: 3.5
    fig_height: 3.5
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/lecture/HOs"); source("../../rhelpers/rhelpers.R")
fnm <- "UnivariateEDA"
modHTML(fnm)
```

```{r echo=FALSE, results='hide', message=FALSE}
fnm <- "UnivariateEDA"
source("../../rhelpers/knitr_setup.R")
```

## Background
Measurements of the levels of arsenic in the drinking water, cooking water, and toenail samples, as well as related covariates, were measured on 21 individuals with private wells in a New Hampshire community.  The variables below were recorded in the [https://github.com/droglenc/NCData/blob/master/Arsenic.csv](Arsenic.csv) file located on the R Resources web page.

* `age`: Age (yrs) of person
* `sex`: Sex of person
* `usedrink`: Household well used for drinking -- A="$<\frac{1}{4}$", B="$\approx\frac{1}{4}$",  C="$\approx\frac{1}{2}$",  D="$\approx\frac{3}{4}$",  E="$>\frac{3}{4}$"
* `usecook`: Household well used for cooking -- A="$<\frac{1}{4}$", B="$\approx\frac{1}{4}$",  C="$\approx\frac{1}{2}$",  D="$\approx\frac{3}{4}$",  E="$>\frac{3}{4}$"
* `arswater`: Arsenic in water (ppm)
* `arsnails`: Arsenic in toenails (ppm)

## Getting the Data
```{r message=FALSE, warning=FALSE}
library(NCStats)
setwd("C:/aaaWork/Web/GitHub/NCMTH107/lecture/HOs")
Ars <- read.csv("Arsenic.csv")
str(Ars)
view(Ars)
```

## Univariate EDA -- Quantitative
```{r}
Summarize(~age,data=Ars,digits=2)
hist(~age,data=Ars,main="",xlab="Age (yrs)")
```

## Univariate EDA -- Quantitative (Separated by Groups)
```{r fig.width=2*figsz, fig.height=figsz, par2=TRUE}
Summarize(age~sex,data=Ars,digits=2)
hist(age~sex,data=Ars,xlab="Age (yrs)",col="gray90")
```

## Univariate EDA -- Categorical
```{r}
( tbl.drink <- xtabs(~usedrink,data=Ars) )
percTable(tbl.drink,digits=1)
barplot(tbl.drink,xlab="Rating of Use for Drinking",ylab="Frequency",col="gray90")
```

```{r echo=FALSE, results="hide", message=FALSE, warnings=FALSE}
purl2(paste0(fnm,".Rmd"))
```
