---
title: "Answer Key"
subtitle: "Goodness-of-Fit Test Class Example"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3.5
    fig_width: 3.5
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file and moves to CE directory
source("1_CE_Keys/zzz_modHTML_CEKeys.R")
modHTML_CEKeys("KEY_GOFTest_CExmpl")
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
```

----

## Green Parakeet Crosses I

<ol>
<li>&alpha; = 0.1.</li>
<li>H<sub>0</sub>: "The distribution of parakeets will follow the 9 green to 3 blue to 3 yellow to 1 white theoretical distribution" and H<sub>A</sub>: "The distribution of parakeets will NOT follow the 9 green to 3 blue to 3 yellow to 1 white theoretical distribution".</li>
</ol>

<div class="alert alert-warning">
<ul>
  <li>The hypotheses for a goodness-of-fit test will usually be "wordy" like this.</li>
  <li>The H<sub>0</sub> will always have the format of the "distribution of INDIVIDUALS into the LEVELS OF THE RESPONSE VARIABLE will follow the THEORETICAL DISTRIBUTION." Thus, the H<sub>0</sub> above simply has the capitalized items replaced with their specific meanings in this example. That is, the individuals are parakeets, the levels of the response variable are the colors (i.e. phenotypes), and the theoretical distribution is the 9:3:3:1 genetic theory.</li>
  <li>The H<sub>A</sub> will be exactly like H<sub>0</sub> except that it will read "NOT follow."</li>
</ul>
</div>

<ol start="3">
  <li>A Goodness-of-Fit test will be used because</li>
  <ol type="i">
    <li>one group/population considered (this cross of parakeets) and</li>
    <li>a categorical variable (phenotypic color) was recorded.</li>
  </ol>
</ol>

<div class="alert alert-warning">
<ul>
  <li>These are the two characteristics that separate a Goodness-of-Fit test from other tests we will see this semester.</li>
  <li>Please include some sort of description that shows that the characteristics are met and that you are not just listing the characteristics -- in other words, include the parts in parentheses.</li>
</ul>
</div>

<ol start="4">
  <li>This study is experimental because the researcher performed the crosses.</li>
  <li>The assumptions are met because the expected table has more than five individuals in all six cells</li>
</ol>  

| Green  |  Blue  | Yellow | White
|:------:|:------:|:------:|:-----:
| 84.375 | 28.125 | 28.125 | 9.375

<br>

<div class="alert alert-warning">
<ul>
  <li>Each expected value is found by multiplying the sample size (n=150) by the expected proportion in each group. Thus, the expected values for "Green" is $150\frac{9}{16}$, for blue and yellow is $150\frac{3}{16}$, and for white is $150\frac{1}{16}$.</li>
  <li>Expected values are expected counts of individuals, but we traditionally keep at least one decimal place so that our calculations below will be more precise.</li>
</ul>
</div>

<ol start="6">
  <li>The statistic for a Goodness-of-Fit test is the observed frequency table.</li>
</ol>

| Green | Blue | Yellow | White | Total 
|:-----:|:----:|:------:|:-----:|:----:
|   95  |  27  |   22   |   6   |  150

<br>

<div class="alert alert-warning">
  <ul>
  <li>I usually add the marginal totals to my observed tables in preparation for making the expected tables.</li>
  </ul>
</div>

<ol start="7">
  <li>The &chi;<sup>2</sup> test statistic is 3.932, as computed from the observed and expected tables as shown below. This test statistic has 3 df.</li>
</ol>

\[ \chi^{2} = \sum\frac{(obs-exp)^{2}}{exp} = \]
\[ \frac{(95-84.375)^{2}}{84.375} + \frac{(27-28.125)^{2}}{28.125} + \frac{(22-28.125)^{2}}{28.125} + \frac{(6-9.375)^{2}}{9.375} = \]
\[ 1.338 + 0.045 + 1.334 + 1.215 = \]
\[ 3.932 \]

<div class="alert alert-warning">
  <ul>
  <li>I urge you to write out each of the parts of the test statistic and the intermediate computation for each part. This will help you solve problems if something goes sideways.</li>
  <li>The df is the number of groups (4) minus 1.</li>
  </ul>
</div>

<ol start="8">
  <li>p-value`r kPvalue(distrib(3.932,distrib="chisq",df=3,lower.tail=FALSE),include.p=FALSE,latex=FALSE)` as computed with `distrib(3.932,distrib="chisq",df=3,lower.tail=FALSE)`.</li>
</ol>

<div class="alert alert-warning">
  <ul>
  <li>When computing the p-value on a chi-square distribution it will always be a "right-of" and, thus, will always include `lower.tail=FALSE`.</li>
  </ul>
</div>

<ol start="9">
<li>Do not reject H<sub>0</sub> because the p-value &gt; &alpha;.</li>
<li>It appears that the phenotypic colors of the parakeets follows the 9:3:3:1 ratio suggested by the genetic theory. Thus, the genetic theory is upheld by these data.</li>
</ol>

<div class="alert alert-warning">
  <ul>
    <li>There is not usually a Step 11 for a Chi-square test. We will not ever do a Step 11 in this class.</li>
  </ul>
</div>
