---
title: "Answer Key"
subtitle: "Univ EDA - Quantitative II CE"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    self_contained: false
---

```{r echo=FALSE, eval=FALSE}
## Renders an appropriate HTML file for the webpage
fnm <- "UEDAQuant2_CE_Keys"
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/CE/Keys")
source("../../../rhelpers/rhelpers.R")
modHTML(fnm)
## And moves it to make it available to students
file.rename(from=paste0(fnm,".html"),to=paste0("../",fnm,".html"))
dir.create(paste0("../",fnm,"_files"),showWarnings=FALSE)
file.copy(from=paste0(fnm,"_files"),to=paste0("../"),recursive=TRUE)
```

```{r echo=FALSE, results='hide', message=FALSE}
library(NCStats)
source("../../../rhelpers/knitr_setup.R")
library(captioner)
tbls <- captioner(prefix="Table")
tbls(name="CWD1","Summary statistics for the diameter (cm) of coarse woody debris in Allequash Lake, WI.")
tbls(name="CWD2","Summary statistics for the diameter (cm) of coarse woody debris in Allequash Lake, WI separated by exposure category.")
tbls(name="Cafe1","Summary statistics of the number of coffees sold in the college cafe.")
tbls(name="Cafe2","Summary statistics of the number of coffees sold in the college cafe separated by day of the week.")
tbls(name="Cafe3","Summary statistics of the number of sodas sold in the college cafe.")
tbls(name="Cafe4","Summary statistics of the number of sodas sold in the college cafe separated by day of the week.")

figs <- captioner(prefix="Figure")
figs(name="CWD1","Histogram of the diameter (cm) of coarse woody debris in Allequash Lake, WI.")
figs(name="CWD2","Histogram of the diameter (cm) of coarse woody debris in Allequash Lake, WI separated by exposure category.")
figs(name="Cafe1","Histogram of the number of coffees sold in the college cafe.")
figs(name="Cafe3","Histogram of the number of sodas sold in the college cafe.")
```

----

## Shape and Outliers I

1. (Strongly) Right skewed, possible outlier at maximum.
1. Left skewed, no outliers.
1. Right skewed, outlier at maximum.
1. Approximately symmetric, no outliers.
1. Approximately symmetric, no outliers.
1. (Srongly) Left skewed, no outliers.
1. Left skewed, outlier at minimum.
1. Bimodal (symmetric), no outliers.

----

## Shape and Outliers II

1. Approximately symmetric (possibly slightly right-skewed).
1. No outliers.
1. I would use the mean and standard deviation because the distribution is approximately symmetric with no outliers.

----

## Shape and Outliers III

1. Strongly left-skewed.
1. Outliers between 30-40 and between 50-60 mm.
1. I would use the median and IQR because the distribution is strongly skewed and outliers are present.

----

## Mean vs Median
1. Generally, (i) right-skewed, (ii) no outliers, and (iii) mean is less than median.
1. Generally, (i) left-skewed, (ii) no outliers, and (iii) mean is greater than median.
1. The mean is greater than the median for right-skewed distributions and less than the median for left-skewed distributions.
1. The mean is equal to the median for perfectly symmetric distributions.
1. Generally, (i) symmetric, (ii) outlier at minimum, and (iii) mean is less than median.
1. Generally, (i) symmetric, (ii) outlier at maximum, and (iii) mean is greater than median.
1. Outliers attract the mean but have little effect on the median. For example, if the outlier is at the maximum value then the mean will be greater than the median.
1. See below.
    1. Mean > Median
    1. Mean < Median (only slightly)
    1. Mean > Median
    1. Mean = Median (approximately)
    1. Mean = Median (approximately)
    1. Mean < Median
    1. Mean < Median
    1. Mean = Median (approximately)
1. The mean uses the actual values whereas the median is only concerned with relative positions.

----

## Commute Time
The distribution of commute times is approximately symmetric with no outliers (Figure 1). The center as measured by the mean is 23.75 mins and the disperision as measured by the standard deviation is 3.54 mins (Table 1). The mean and standard deviation were used because the distribution was symmetric and no outliers were present.

----

## Dungeness Crabs

The distribution of post-molt carapace lengths for Dungeness Crabas is strongly left-skewed with two outliers between 38.8 and 60 mm (Figure 2). The center as measured by the median is 147.4 mm and the dispersion as measured by the IQR is from a Q1 of 138.0 to a Q3 of 153.4 mm (Table 2). The median and IQR were used because the distribution was strongly skewed and outliers were present.

----

## Coarse Woody Debris
<ol>
<li>The diameter of CWD is largely right-skewed with a separate group of four individuals at higher diameters (`r figs("CWD1",display="cite")`).  With this definition of shape, there are no outliers.  The center as measured by the median is 18 cm and the IQR is from a Q1 of 16.25 to a Q3 of 20.75 cm (`r tbls("CWD1",display="cite")`).  I chose to use the median and IQR because of the right-skewed shape.</li>
```{r CWDHist1, echo=FALSE, fig.width=3, fig.height=3, par1=TRUE}
d <- read.csv("cwd.csv")
hist(~diameter,data=d,xlab="Diameter of CWD")
```

`r figs("CWD1")`

`r tbls("CWD1")`

```{r CWDSum1, echo=FALSE, background="white"}
Summarize(~diameter,data=d,digits=2)
```

<li>The diameter of CWD is largely right-skewed for both low- and medium-exposure sites (`r figs("CWD2",display="cite")`).  There are no obvious outliers.  [<i>Note that it is difficult to describe shape and outliers in this case because the sample sizes in the two groups are so small.</i>]  The medians are 19 and 18, respectively, for the low- and medium-exposure sites (`r tbls("CWD2",display="cite")`).  The IQR is from a Q1 of 16 to a Q3 of 24 for the low-exposure sites and from a Q1 of 17 to a Q3 of 20 for the medium-exposure sites (`r tbls("CWD2",display="cite")`).  The median and IQR were used because of the skewed shape for both distributions.</li>

```{r CWDHist2, echo=FALSE, fig.width=6, fig.height=3, par1=TRUE}
hist(diameter~exposure,data=d,xlab="Diameter of CWD")
```

`r figs("CWD2")`

`r tbls("CWD2")`

```{r CWDSum2, echo=FALSE, background="white"}
Summarize(diameter~exposure,data=d,digits=2)
```
</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
<<CWDHist1>>
<<CWDSum1>>
<<CWDHist2>>
<<CWDSum2>>
```

----

## Cafe Sales
<ol>
<li>The distribution of coffees sold was right-skewed with a hint of bimodality (a small peak near 10 coffees sold and a taller peak near 30 coffees sold), but no obvious outliers (`r figs("Cafe1",display="cite")`).  The center as measured by the median was 23 coffees sold and the IQR was from a Q1 of 12 to a Q3 of 30 coffees sold (`r tbls("Cafe1",display="cite")`).  I chose to use the median and IQR to measure center and dispersion because of the skewed shape.</li>

```{r CafeHist1, echo=FALSE, fig.width=3, fig.height=3, par1=TRUE}
d <- read.csv("cafedata.csv")
d$dow <- factor(d$dow,levels=c("Mon","Tue","Wed","Thu","Fri"))
hist(~coffee.sold,data=d,xlab="Number of Coffees Sold",ylim=c(0,10))
```

`r figs("Cafe1")`

`r tbls("Cafe1")`

```{r CafeSum1, echo=FALSE}
Summarize(~coffee.sold,data=d,digits=2)
```

<li>In general, slightly more coffees were sold in the middle of the week (`r tbls("Cafe2",display="cite")`).  There was slightly less variability on Mondays (`r tbls("Cafe2",display="cite")`). [<i>I did not make histograms for each different day of the week because the sample sizes for each day were small (<=10).</i>]</li>

`r tbls("Cafe2")`

```{r CafeSum2, echo=FALSE}
Summarize(coffee.sold~dow,data=d,digits=2)
```

<li>The distribution of sodas sold was primarily symmetric with no obvious outliers (`r figs("Cafe3",display="cite")`).  The center as measured by the mean was 29.6 sodas sold and the dispersion as measured by the standard deviation was 11.9 sodas sold (`r tbls("Cafe3",display="cite")`).  I chose to use the mean and standard deviation to measure center and dispersion because of the symmetric shape and lack of outliers.</li>

```{r CafeHist3, echo=FALSE, fig.width=3, fig.height=3, par1=TRUE}
hist(~sodas.sold,data=d,xlab="Number of Sodas Sold",ylim=c(0,10))
```

`r figs("Cafe3")`

`r tbls("Cafe3")`

```{r CafeSum3, echo=FALSE}
Summarize(~sodas.sold,data=d,digits=2)
```

<li>In general, slightly more coffees were sold in the middle of the week (`r tbls("Cafe4",display="cite")`).  There was slightly less variability on Mondays (`r tbls("Cafe4",display="cite")`).</li>

`r tbls("Cafe4")`

```{r CafeSum4, echo=FALSE}
Summarize(sodas.sold~dow,data=d,digits=2)
```

<li>There was likely a smaller "population" of potential customers on Mondays and Fridays then during the middle of the week.</li>
</ol>

R Appendix.
```{r eval=FALSE, prompt=FALSE}
<<CafeHist1>>
<<CafeSum1>>
<<CafeSum2>>
<<CafeHist3>>
<<CafeSum3>>
<<CafeSum4>>
```
