---
title: "Review Exercise Key"
subtitle: "Bivariate EDA - Quantitative"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/zRevExAns/")
source("../../rhelpers/rhelpers.R")
modHTML("BEDAQuant_Defns")
```
```{r echo=FALSE}
library(NCStats)
library(captioner)
figs <- captioner(prefix="Figure")
```

----

## Order Scatterplots I

* The order from lowest (most negative) to highest (most positive) value of R is B, C, D, A.
* The order from weakest to strongest is D, C, A, B.

----

## Order Scatterplots II

* The order from lowest (most negative) to highest (most positive) value of R is B, A, C, D.
* The order from weakest to strongest is C, A, D, B.

----

## Guess r

* The scatterplots that most likely correspond to the given values of r are: 0.9 is B, -0.5 is D, -0.9 is H, 0.6 is C, and 0 is G.

----

## Hemlock and Deer

```{r echo=FALSE, results='hide'}
df <- data.frame(saplings=c(0.95,2.89,2.97,3.94,4.74,5.10,6.64,7.13),
                 browse=c(0.31,0.35,0.49,0.50,0.61,0.63,0.86,0.90))
figs(name="HemlockPlot","Scatterplot of wolf territory size versus deer density.")
```

The scatterplot between mean number of hemlock samplings and the deer browsing index is shown in `r figs("HemlockPlot",display="cite")`. The correlation coefficients is `r formatC(corr(saplings~browse,data=df),format="f",digits=2)`.

```{r HemlockPlot, echo=FALSE}
par(mar=c(3,3,0.5,0.5),mgp=c(1.9,0.5,0),tcl=-0.2)
plot(saplings~browse,data=df,pch=21,bg="gray70",
     ylab="Mean Number of Hemlock Saplings",xlab="Deer Browsing Index")
```

`r figs("HemlockPlot")`

### R Appendix
```{r eval=FALSE}
df <- read.csv("data/Hemlock.csv")
plot(saplings~browse,data=df,pch=21,bg="gray70",
     ylab="Mean Number of Hemlock Saplings",xlab="Deer Browsing Index")
corr(saplings~browse,data=df)
```

----
