---
title: "Review Exercise Key"
subtitle: "Data Production - Experimental Design"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/zRevExAns/")
source("../../rhelpers/rhelpers.R")
modHTML("DataProduction_ExpDChemRxn")
```

1. The response variable is the yield of the new product.
1. The two factors are temperature and stirring rate.
1. There are two levels of temperature and three levels of stirring rates.
1. There are six treatments in this experiment.
1. There are five vessels allocated to each treatment.
1. An individual or replicate is a vessel.
1. I would allocate vessels by assigning each a unique number from between 1 and 30 and then randomly ordering the numbers between 1 and 30. Vessles corresponding to the first five numbers would go in the first treatment, the next five in the second treatment, and so on. The random numbers would be drawn in R with
```{r}
vessels <- sample(30)
vessels[1:5]  # 50C, 60rpm
vessels[6:10] # 50C, 90rpm
# etc. see below
```

This allocation could be visualized in the following table (note that the "X" columns are simply counters for individuals allocated to that treatment).
```{r echo=FALSE}
data.frame(Treatment=1:6,
           Temp=rep(c("50C","60C"),each=3),
           Stir=rep(c("60 rpm","90 rpm","120 prm"),2),
           rbind(vessels[1:5],vessels[6:10],vessels[11:15],
                 vessels[16:20],vessels[21:25],vessels[26:30]))
```

----
