---
title: "Review Exercise Key"
subtitle: "Data Production - Experimental Design"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/zRevExAns/")
source("../../rhelpers/rhelpers.R")
modHTML("DataProduction_ExpDBeets")
```

1. The two factors are plowing depth and amount of fertilizer.
1. There are three levels of plowing depth and two levels of fertilizer amount.
1. There are six total treatments.
1. There would be six plots in each treatment.
1. A plot or field is the individual or replicate in this experiment.
1. The response variable is the "harvest" of sugar beets.
1. I would allocate plots by assigning each plot a unique number from between 1 and 36 and then randomly ordering the numbers between 1 and 36. Plots corresponding to the first six numbers would go in the first treatment, the next six in the second treatment, and so on. The random numbers would be drawn in R with
```{r}
rplots <- sample(36)
rplots[1:6]  # 10cm, no fert
rplots[7:12] # 10cm, 3kg fert
# etc. see below
```

This allocation could be visualized in the following table (note that the "X" columns are simply counters for individuals allocated to that treatment).
```{r echo=FALSE}
data.frame(Treatment=1:6,
           Depth=rep(c("10 cm","17 cm","25 cm"),each=2),
           Temperature=rep(c("No fert","3kg fert"),3),
           rbind(rplots[1:6],rplots[7:12],rplots[13:18],
                 rplots[19:24],rplots[25:30],rplots[31:36]))
```

----
