---
title: "Review Exercise Key"
subtitle: "Data Production - Experimental Design"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/zRevExAns/")
source("../../rhelpers/rhelpers.R")
modHTML("DataProduction_ExpDSeals")
```

1. The health of the seals is the response variable.
1. The two factors are water temperature and seal origin (i.e., wild or domestic).
1. There are two levels of water temperature and two levels of seal origin.
1. There are four total treatments.
1. There are ten seals per treatment.
1. A seal is an individual/replicate.
1. I would randomly select ten of the wild seals for the <47F treatment (with the other ten put in the >47F treatment) and then do the same for the domestic seals. Thus, I would generate two random selections of ten numbers from twenty numbers.
```{r}
wseals <- sample(20)
wseals[1:10]  # wild <47F
wseals[11:20] # wild >47F
dseals <- sample(20)
dseals[1:10]  # domestic <47F
dseals[11:20] # domestic >47F
```

This allocation could be visualized in the following table (note that the "X" columns are simply counters for individuals allocated to that treatment).
```{r echo=FALSE}
data.frame(Treatment=1:4,
           Seal=rep(c("Wild","Domestic"),each=2),
           Temperature=rep(c("<47",">47"),2),
           rbind(wseals[1:10],wseals[11:20],dseals[1:10],dseals[11:20]))
```



----
