---
title: "Review Exercise Key"
subtitle: "Data Production - Experimental Design"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/zRevExAns/")
source("../../rhelpers/rhelpers.R")
modHTML("DataProduction_ExpDTrans")
```

1. The response variable is whether the rabbit survives or not.
1. The two factors are tranquilization and acclimatization.
1. There are two levels for tranquilization (tranquilized, not tranquilized) and two levels for acclimatization (allowed to acclimate, not allowed to acclimate).
1. There are four treatments.
1. A rabbit is a replicate or individual in this experiment.
1. There are 16 rabbits in each treatment.
1. I would allocate rabbits by assigning each rabbit a unique number from between 1 and 64 and then randomly ordering the numbers between 1 and 64. Rabbits corresponding to the first sixteen numbers would go in the first treatment, the next sixteen in the second treatment, and so on. The random numbers would be drawn in R with
```{r}
rabbits <- sample(64)
rabbits[1:16]  # tranquilized, allowed
rabbits[17:32] # tranquilized, not allowed
# etc. see below
```

This allocation could be visualized in the following table (note that the "X" columns are simply counters for individuals allocated to that treatment).
```{r echo=FALSE}
data.frame(Treatment=1:4,
           Tranquilized=rep(c("Yes","No"),each=2),
           Acclimated=rep(c("Yes","No"),2),
           rbind(rabbits[1:16],rabbits[17:32],rabbits[33:48],rabbits[49:64]))
```

----
