---
title: "Review Exercise Key"
subtitle: "Data Production - Experimental Design"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/zRevExAns/")
source("../../rhelpers/rhelpers.R")
modHTML("DataProduction_ExpDDeerDiet")
```

1. The response variable is the health of the deer.
1. The factor is location and what the deer ate (*combined together as one factor as there is not a complete crossing of levels for two factors*).
1. There are four levels for the one factor.
1. There are four treatments.
1. There are 16 deer per treatment.
1. A deer is the individual or replicate in this experiment.
1. I would allocate deer by assigning each deer a unique number from between 1 and 64 and then randomly ordering the numbers between 1 and 64. Deer corresponding to the first sixteen numbers would go in the first treatment, the next sixteen in the second treatment, and so on. The random numbers would be drawn in R with
```{r}
deer <- sample(64)
deer[1:16]  # farm w/ strict diet
deer[17:32] # island in prairies
# etc. see below
```

This allocation could be visualized in the following table (note that the "X" columns are simply counters for individuals allocated to that treatment).
```{r echo=FALSE}
data.frame(Treatment=1:4,
           Group=c("Farm, strict","Island, prairie","Island, hardwood","Island, strict"),
           rbind(deer[1:16],deer[17:32],deer[33:48],deer[49:64]))
```

----
