---
title: "Review Exercise Key"
subtitle: "Data Production - Experimental Design"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/zRevExAns/")
source("../../rhelpers/rhelpers.R")
modHTML("DataProduction_ExpDBP")
```

1. The two factors are calcium and exercise.
1. There are two levels of calcium and two levels of exercise.
1. There are four treatments.
1. The response variable is the blood pressure reading.
1. There are eight male subjects per treatment.
1. A male subject is the individual or replicate in this experiment.
1. I would allocate mean by assigning each a unique number from between 1 and 32 and then randomly ordering the numbers between 1 and 32. Men corresponding to the first eight numbers would go in the first treatment, the next eight in the second treatment, and so on. The random numbers would be drawn in R with
```{r}
rmales <- sample(32)
rmales[1:8]  # Calcium, exercise
rmales[9:16] # Calcium, no exercise
# etc. see below
```

This allocation could be visualized in the following table (note that the "X" columns are simply counters for individuals allocated to that treatment).
```{r echo=FALSE}
data.frame(Treatment=1:4,
           Pill=rep(c("Calcium","Placebo"),each=2),
           Exercise=rep(c("Yes","No"),2),
           rbind(rmales[1:8],rmales[9:16],rmales[17:24],rmales[25:32]))
```

----
