---
title: "Review Exercise Key"
subtitle: "Normal Distributions"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/zRevExAns/")
source("../../rhelpers/rhelpers.R")
modHTML("NormalDist_Calc2")
```
```{r echo=FALSE, message=FALSE}
library(NCStats)
```

## College Graduation

```{r NormGrad, echo=FALSE, results='hide', fig.show='hide'}
( q1 <- distrib(21,mean=22.1,sd=1.1) )
( q2 <- distrib(24,mean=22.1,sd=1.1,lower.tail=FALSE) )
( q3a <- distrib(0.025,mean=22.1,sd=1.1,type="q") )
( q3b <- distrib(0.975,mean=22.1,sd=1.1,type="q") )
( q4 <- distrib(0.9,mean=22.1,sd=1.1,type="q",lower.tail=FALSE) )
```

1. Approximately `r formatC(100*q1,format="f",digits=1)`% of the students have graduated by age 21.
1. Approximately `r formatC(100*q2,format="f",digits=1)`% of the students graduated after age 24.
1. The most common 95% of graduates are between `r formatC(q3a,format="f",digits=1)` and  is `r formatC(q3b,format="f",digits=1)` years old.
1. The age at which 90% of the students have graduated is `r formatC(q4,format="f",digits=1)`

### R Appendix
```{r fig.show='hide', results='hide'}
<<NormGrad>>
```

----

## Black Bear Size

```{r NormBears, echo=FALSE, results='hide', fig.show='hide'}
library(NCStats)
( q1a <- distrib(45,mean=60,sd=10) )
( q1ab <- distrib(65,mean=60,sd=10) )
q1ab-q1a
( q2 <- distrib(0.2,mean=20.9,sd=4.8,type="q",lower.tail=FALSE) )
( q3a <- distrib(0.25,mean=13.7,sd=1.9,type="q") )
( q3b <- distrib(0.75,mean=13.7,sd=1.9,type="q") )
```

1. Approximately `r formatC(100*(q1ab-q1a),format="f",digits=1)`% of the bears have a body length between 45" and 65".
1. The largest 20% of neck girths are greater than `r formatC(q2,format="f",digits=1)`".
1. The most common 50% of head lengths are between `r formatC(q3a,format="f",digits=2)` and `r formatC(q3b,format="f",digits=2)`.
1. This question cannot be answered because weight is not know to be normally distributed.

### R Appendix
```{r fig.show='hide', results='hide'}
<<NormBears>>
```

----

## Shrew Brain Weights

```{r NormShrew, echo=FALSE, results='hide', fig.show='hide'}
( q1 <- distrib(0.09,mean=0.14,sd=0.04) )
( q2 <- distrib(0.17,mean=0.14,sd=0.04) )
q2-q1
( q3 <- distrib(0.3,mean=0.14,sd=0.04,type="q",lower.tail=FALSE) )
```

1. Approximately `r formatC(100*q1,format="f",digits=1)`% of short-tailed shrews have a brain weight of less than 0.09 g.
1. Approximately `r formatC(100*(q2-q1),format="f",digits=1)`% of short-tailed shrews have a brain weight between 0.09 and 0.17 g.
1. The largest 30% of short-tailed shrew brain weights are larger than `r formatC(q3,format="f",digits=2)` g.

### R Appendix
```{r fig.show='hide', results='hide'}
<<NormShrew>>
```

----

## Bus Arrival Times

```{r NormBART, echo=FALSE, results='hide', fig.show='hide'}
( q1 <- distrib(5,mean=0,sd=3,lower.tail=FALSE) )
( q2 <- distrib(-4,mean=0,sd=3) )
( q3 <- distrib(4,mean=0,sd=3) )
q3-q2
( q4 <- distrib(0.25,mean=0,sd=3,type="q",lower.tail=FALSE) )
( q5a <- distrib(0.2,mean=0,sd=3,type="q") )
( q5b <- distrib(0.8,mean=0,sd=3,type="q") )
```

1. Approximately `r formatC(100*q1,format="f",digits=1)`% of arrivals are more than 5 minutes late.
1. Approximately `r formatC(100*q2,format="f",digits=1)`% of arrivals are more than 4 minutes early.
1. Approximately `r formatC(100*(q3-q2),format="f",digits=1)`% of arrivals are between 4 minutes early and 4 minutes late.
1. The latest 25% of arrival times are more than `r formatC(q4,format="f",digits=1)` minutes late.
1. The most common 60% of arrival times are between `r formatC(q5a,format="f",digits=1)` and `r formatC(q5b,format="f",digits=1)` minutes.
1. The arrival time is a continuous quantitative variable.


### R Appendix
```{r fig.show='hide', results='hide'}
<<NormBART>>
```

----

## Roe Deer Fawns

```{r NormRoeDeer, echo=FALSE, results='hide', fig.show='hide'}
( q1 <- distrib(2,mean=2.235,sd=0.460) )
( q2 <- distrib(3,mean=2.235,sd=0.460,lower.tail=FALSE) )
( q3a <- distrib(1,mean=2.235,sd=0.460) )
( q3ab <- distrib(3,mean=2.235,sd=0.460) )
q3ab-q3a
( q4 <- distrib(0.076,mean=2.235,sd=0.460,type="q") )
( q5 <- distrib(0.042,mean=2.235,sd=0.460,type="q",lower.tail=FALSE) )
( q6a <- distrib(0.065,mean=2.235,sd=0.460,type="q") )
( q6b <- distrib(0.935,mean=2.235,sd=0.460,type="q") )
```

1. Approximately `r formatC(100*q1,format="f",digits=1)`% of does have less than two fawns.
1. Approximately `r formatC(100*q2,format="f",digits=1)`% of does have more than three fawns.
1. Approximately `r formatC(100*(q3ab-q3a),format="f",digits=1)`% of does have between one and three fawns.
1. The number of fawns such that 7.6% of does have fewer fawns is `r formatC(q4,format="f",digits=1)`.
1. The number of fawns such that 4.2% of does have more fawns is `r formatC(q5,format="f",digits=1)`.
1. The most common 87% of number of fawns per does are between `r formatC(q6a,format="f",digits=1)` and `r formatC(q6b,format="f",digits=1)`.

### R Appendix
```{r fig.show='hide', results='hide'}
<<NormRoeDeer>>
```

----

## Commercial Crayfishing

```{r NormCrayfish, echo=FALSE, results='hide', fig.show='hide'}
( q1ab <- distrib(110,mean=93,sd=8) )
( q1a <- distrib(90,mean=93,sd=8) )
50000*(q1ab-q1a)
50000*(1-q1ab)
```

1. I could send the Swedish market approximately `r formatC(50000*(q1ab-q1a),format="f",digits=0)` crayfish.
1. I could send the alternative market approximately `r formatC(50000*(1-q1ab),format="f",digits=0)` crayfish.


### R Appendix
```{r fig.show='hide', results='hide'}
<<NormCrayfish>>
```

----
