---
title: "Review Exercise Key"
subtitle: "Probability"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/zRevExAns/")
source("../../rhelpers/rhelpers.R")
modHTML("Probability")
```
```{r echo=FALSE, message=FALSE}
library(NCStats)
```

## Coin Purse

1. The probability of randomly selecting a nickel from this purse is `r formatC(17/32,format="f",digits=3)` (i.e., $\frac{17}{32}$).
1. The probability of randomly selecting a dime from this purse is `r formatC(15/32,format="f",digits=3)` (i.e., $\frac{15}{32}$).
1. The probability of randomly selecting a dime if two nickels and three dimes had been removed is `r formatC(12/27,format="f",digits=3)` (i.e., $\frac{12}{27}$).

## Greenhouse Plants

1. The probability of randomly selected a tomato plant is `r formatC(10/30,format="f",digits=3)` (i.e., $\frac{10}{30}$).
1. The probability of randomly selected a cauliflower plant is `r formatC(8/30,format="f",digits=3)` (i.e., $\frac{8}{30}$).
1. The probability of randomly selected a pea plant assuming that all tomato plants were gone is `r formatC(12/20,format="f",digits=3)` (i.e., $\frac{12}{20}$).

## Lengths of Pine Needles
```{r ProbLPN, echo=FALSE, results='hide', fig.show='hide'}
( q1ab <- distrib(80,mean=75,sd=8) )
( q1a <- distrib(70,mean=75,sd=8) )
q1ab-q1a
( q2 <- distrib(90,mean=75,sd=8,lower.tail=FALSE) )
( q3 <- distrib(50,mean=75,sd=8) )
```

1. The probability that a randomly selected needle is between 70 and 80 mm long is
`r formatC(q1ab-q1a,digits=3,format="f")`.
1. The probability that a randomly selected needle is longer than 90 mm is `r formatC(q2,digits=3,format="f")`.
1. The probability that a randomly selected needle is less than 50 mm is `r formatC(q3,digits=4,format="f")`.

### R Appendix
```{r fig.show='hide', results='hide'}
<<ProbLPN>>
```

----
