---
title: "Review Exercise Key"
subtitle: "Sampling Distributions"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/zRevExAns/")
source("../../rhelpers/rhelpers.R")
modHTML("SamplingDist")
```
```{r echo=FALSE}
library(NCStats)
library(captioner)
figs <- captioner(prefix="Figure")
```

----

## Sampling Distribution Characteristics I

1. The shape will be normal because the population is normal (or $n\geq30$).
1. The center will be equal to 100 because $\bar{x}$ is an unbiased estimator of $\mu$=100.
1. This is the same as the SE which is $\frac{\sigma}{\sqrt{n}}$=$\frac{20}{\sqrt{50}}$=2.83.
1. This is the same as the previous question, i.e., 2.83.

----

## Sampling Distribution Characteristics II

  1. The shape will be normal because $n\geq30$.
  1. The center will be equal to 500 because $\bar{x}$ is an unbiased estimator of $\mu$=500.
  1. This is the same as the SE which is $\frac{\sigma}{\sqrt{n}}$=$\frac{60}{\sqrt{100}}$=6.
  1. This is the same as the previous question, i.e., 6.

----

## Sampling Distribution Characteristics III

1. The shape will be normal because $n\geq15$ and the population is only slightly skewed.
1. The center will be equal to 500 because $\bar{x}$ is an unbiased estimator of $\mu$=500.
1. This is the same as the SE which is $\frac{\sigma}{\sqrt{n}}$=$\frac{60}{\sqrt{20}}$=13.42.
1. This is the same as the previous question, i.e., 13.42.

----

## Accuracy and Precision

1. The 9, 10, 11, and 9 means are more accurate.
1. The 6, 14, 8, and 12 means are more accurate.
1. The 7, 7, 9, and 8 means are more precise.
1. The 2,8,12, and 18 means are accurate, but imprecise.
1. The means 9,10,11, and 10 are accurate and precise.
1. The means 1,7,8, and 19 are inaccurate and imprecise.

----

## Raccoons Consuming Corn
```{r NormRocky, echo=FALSE, include=FALSE}
( q2 <-  distrib(10,mean=8,sd=2/sqrt(25),lower.tail=FALSE) )
( q3 <- distrib(8.5,mean=8,sd=2/sqrt(60),lower.tail=FALSE) )

```

1. This question is about an individual and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of  $\bar{x}\sim N(8,\frac{2}{\sqrt{25}})$ because $n=25 \geq 15$ and the population distribution is not strongly skewed.  Thus, the probability is `r kPvalue(q2,include.p=FALSE,latex=FALSE)`.
1. This question is ultimately about a statistic as the total of 510 should be converted an $\bar{x}$ of $\frac{510}{60}=8.5$ kg.  The sampling distribution of $\bar{x}\sim N(8,\frac{2}{\sqrt{60}})$ or $\bar{x}\sim N(8,0.2582)$ because $n=60 \geq 30$.  Thus, the probability is `r formatC(q3,digits=4,format="f")`.

### R Appendix
```{r eval=FALSE}
<<NormRocky>>
```

----

## NFL Running Backs 
```{r NormNFL, echo=FALSE, include=FALSE}
( q2 <- distrib(100,mean=82,sd=26/sqrt(16),lower.tail=FALSE) )
q3ab <- distrib(90,mean=82,sd=26/sqrt(16))
q3a <- distrib(70,mean=82,sd=26/sqrt(16))
q3ab-q3a
( q4 <- distrib(70,mean=82,sd=26/sqrt(32),lower.tail=FALSE) )
( q6 <- distrib(0.05,mean=82,sd=26/sqrt(16),type="q",lower.tail=FALSE) )

```

1. This question is about an individual (a game) and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(82,\frac{26}{\sqrt{16}})$ because $n=16 \geq 15$ and the population distribution is not strongly skewed.  Thus, the probability is `r formatC(q2,digits=4,format="f")`.
1. This question is about a statistic and requires the same sampling distribution as the previous question --  $\bar{x}\sim N(82,\frac{26}{\sqrt{16}})$..  Thus, the probability is `r formatC(q3ab-q3a,digits=4,format="f")`.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(82,\frac{26}{\sqrt{32}})$ because $n=32 \geq 30$.  Thus, the probability is `r formatC(q4,digits=4,format="f")`.
1. This question is about an individual (a game) and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(82,\frac{26}{\sqrt{16}})$ because $n=16 \geq 15$ and the population distribution is not strongly skewed.  Thus, the mean yards gained is `r formatC(q6,digits=1,format="f")`.

### R Appendix
```{r eval=FALSE}
<<NormNFL>>
```

----

## Stock Returns
```{r NormStocks, echo=FALSE, include=FALSE}
( q1 <- distrib(0,mean=4.2,sd=4.9/sqrt(5),lower.tail=FALSE) )
( q2 <- distrib(0,mean=4.2,sd=4.9,lower.tail=FALSE) )
( q3 <- distrib(2,mean=4.2,sd=4.9/sqrt(10)) )
( q4 <- distrib(0.10,mean=4.2,sd=4.9,type="q",lower.tail=FALSE) )
( q5 <- distrib(0.10,mean=4.2,sd=4.9/sqrt(10),type="q",lower.tail=FALSE) )
```

1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(4.2,\frac{4.9}{\sqrt{5}})$ because the population distribution is approximately normal.  Thus, the probability is `r formatC(q1,digits=4,format="f")`.
1. This question is about an individual (a stock) and requires the population distribution. Because the population distribution is approximately normally distributed, the probability can be computed.  Thus, the probability is `r formatC(q2,digits=4,format="f")`.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(4.2,\frac{4.9}{\sqrt{10}})$ because the population distribution is approximately normal.  Thus, the probability is `r formatC(q3,digits=4,format="f")`.
1. This question is about an individual (a stock) and requires the population distribution. Because the population distribution is approximately normally distributed, the probability can be computed.  Thus, the mean return is `r formatC(q4,digits=1,format="f")`.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(4.2,\frac{4.9}{\sqrt{10}})$ because the population distribution is approximately normal.  Thus, the mean return is `r formatC(q5,digits=1,format="f")`.

### R Appendix
```{r eval=FALSE}
<<NormStocks>>
```

----

## HMF in Honey
```{r NormHoney, echo=FALSE, include=FALSE}
( q3 <- distrib(10,mean=9.5,sd=13.5/sqrt(50)) )
( q4 <- distrib(0.20,mean=9.5,sd=13.5/sqrt(50),type="q") )
```

1. This question is about an individual (one kg of honey) and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution, however, is not normal because $n < 30$ and the population distribution is strongly skewed.  Thus, this question CANNOT BE COMPUTED.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(9.5,\frac{13.5}{\sqrt{50}})$ because $n \geq 30$.  Thus, the probability is `r formatC(q3,digits=4,format="f")`.
1. This question is about a statistic and requires the same sampling distribution as in the previous question -- $\bar{x}\sim N(9.5,\frac{13.5}{\sqrt{50}})$.  Thus, the average amount is `r formatC(distrib(0.20,mean=9.5,sd=13.5/sqrt(50),type="q",plot=FALSE),digits=1,format="f")`.

### R Appendix
```{r eval=FALSE}
<<NormHoney>>
```

----

## Size of Farms in England
```{r NormFarms, echo=FALSE, include=FALSE}
( q2a <- distrib(0.45,mean=65.13,sd=108.71/sqrt(60),type="q") )
( q2b <- distrib(0.55,mean=65.13,sd=108.71/sqrt(60),type="q") )
( q3 <- distrib(50,mean=65.13,sd=108.71/sqrt(60)) )
```

1. This question is about an individual (a farm) and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(65.13,\frac{108.71}{\sqrt{60}})$ because $n \geq 30$.  Thus, the most common 10\% of means is between `r formatC(q2a,digits=2,format="f")` and `r formatC(q2b,digits=2,format="f")`.
1. This question is about a statistic and requires the same sampling distribution as in the previous question -- $\bar{x}\sim N(65.13,\frac{108.71}{\sqrt{60}})$.  Thus, the probability is `r formatC(q3,digits=4,format="f")`.
1. This question is about an individual (a farm) and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.

### R Appendix
```{r eval=FALSE}
<<NormFarms>>
```

----

## Painted Turtle Hatchling Size
```{r NormTurtle, echo=FALSE, include=FALSE}
( q2 <- distrib(4.5,mean=4.46,sd=0.13/sqrt(20),lower.tail=FALSE) )
( q3 <- distrib(4.5,mean=4.46,sd=0.13/sqrt(50),lower.tail=FALSE) )
( q4 <- distrib(0.20,mean=4.46,sd=0.13/sqrt(50),type="q") )
```

1. This question is about an individual (a turtle) and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(4.46,\frac{0.13}{\sqrt{20}})$ because $n \geq 15$ and the population is only slightly skewed.  Thus, the probability that a sample of 20 turtles will have an average number of days until hatching that is greater than 4.5 days is `r formatC(q2,digits=4,format="f")`.
1. This question is about a statistic and requires the sampling distribution.  The sampling distribution of $\bar{x}\sim N(4.46,\frac{0.13}{\sqrt{50}})$ because $n \geq 30$.  Thus, the probability that a sample of 50 turtles will have an average number of days until hatching that is greater than 4.5 days is `r formatC(q3,digits=4,format="f")`.
1. This question is about a statistic and requires the same sampling distribution as in the previous question -- $\bar{x}\sim N(4.46,\frac{0.13}{\sqrt{50}})$.  Thus, the mean number of days until hatching such that 20\% of samples of 50 turtles have a smaller mean is `r formatC(q4,digits=2,format="f")`.
1. This question is about an individual (a turtle) and requires the population distribution. It CANNOT BE COMPUTED because the population distribution is not normally distributed.

### R Appendix
```{r eval=FALSE}
<<NormTurtle>>
```

----

## Moose Hunting
```{r NormMoose, echo=FALSE, include=FALSE}
( q4 <- distrib(48,mean=40,sd=15/sqrt(25),lower.tail=FALSE) )
```

1. An individual is a Minnesota moose hunter.
1. Time spent hunting is a continuous quantitative variable.
1. This question is about a quantitative variable and an individual; thus, the population distribution would be used.  This question cannot be answered because the population distribution is NOT normal (it is only known that it is symmetric).
1. This question is about a quantitative variable and a statistic; thus, the sampling distribution of $\bar{x}$ would be used.  Because the population distribution is only symmetric, the sample size must be greater than 15, which it is in this question.  Thus, $\bar{x}\sim N(40,\frac{15}{\sqrt{25}})$ or $\bar{x}\sim N(40,3)$.  Thus, the probability is `r formatC(q4,digits=4,format="f")`.

### R Appendix
```{r eval=FALSE}
<<NormMoose>>
```

----
