---
title: "Review Exercise Key"
subtitle: "Univariate EDA"
author: Derek H. Ogle
layout: page
css: "/css/modules.css"
output:
  html_document:
    fig_height: 3
    fig_width: 3.5
    self_contained: no
---

```{r echo=FALSE, eval=FALSE}
# Renders an appropriate HTML file for the webpage
setwd("C:/aaaWork/Web/GitHub/NCMTH107/modules/zRevExAns/")
source("../../rhelpers/rhelpers.R")
modHTML("UEDAQuant2_EDAs")
```
```{r echo=FALSE}
library(NCStats)
library(captioner)
figs <- captioner(prefix="Figure")
tabs <- captioner(prefix="Table")
```

## Time Spent Playing Video Games

The time spent playing video games is very strongly right-skewed with an extreme outlier at 30 hours and another outlier at approximately 15 hours. The center is at a median of 0 hours and the dispersion is from a Q1 of 0 hours to a Q3 of 1.2 hours. The median and IQR were used because of the skew and presence of outliers. [*Results are derived from the figure and table presented in the question.*]

----

## Bank Transaction Times

The bank transaction times are strongly right-skewed with no obvious outliers. The center is at a median of 93.0 hours and the dispersion is from a Q1 of 60.3 hours to a Q3 of 145.2 hours. The median and IQR were used because of the skew and presence of outliers. [*Results are derived from the figure and table presented in the question.*]

----

## Creatine Phosphate Concentrations
```{r}
cp <- c(121,82,100,151,68,58,95,145,64,119,104,110,113,118,203,62,83,67,201,
101,163,84,57,139,60,78,94,93,92,110,25,123,70,48,95,42)
cpsum <- Summarize(~cp,digits=1)
figs(name="GCP","Histogram of the creatine phosphokinase concentrations in 36 male volunteers.")
tabs(name="GCP","Summary statistics of the creatine phosphokinase concentrations in 36 male volunteers.")
```

The distribution of creatine phosphate concentrations is slightly right-skewed with no obvious outliers (`r figs("GCP",display="cite")`). [*Note that I did not call the individuals above 200 outliers because they were only one bar removed, there were two individuals, and they "felt" like an extension of the right-tail.*] The center is at a mean of `r formatC(cpsum["mean"],format="f",digits=1)` and the dispersion is a standard deviation of `r formatC(cpsum["sd"],format="f",digits=1)`. The mean and standard deviation were used because the distribution was not strongly skewed and outliers were not present.

```{r CPHist, echo=FALSE, comment=""}
hist(~cp,xlab="Creatine Phosphate Concentration")
```

`r figs("CP")`

`r tabs("CP")`
```{r CPSum, echo=FALSE, comment=""}
cpsum
```


# R Code Appendix
```
library(NCStats)
df <- read.csv("data/CreatinePhosphate.csv")
hist(~cp,data=df,xlab="Creatine Phosphate Concentration")
Summarize(~cp,data=df)
```

